---
title: "Projekt"
author: "Stefan Graf"
date: "28 3 2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading environment / libraries

# Todo:

# Data handling
# Zeiten kontrollieren ob bei Originaldatei in den Tracks (anhand Dateiname) teilweise GPS Messungen mit viel späterem Zeitpunkt zwischendrin vorkommen



# Markings:
#### Quality check: --> are quality checks of chunks executed before
########### DELETE ############--> are chunks to delete
## Title --> are titles for sections
### explanations of the following chunk
##### SUAQ TRACKS 2018 & 2019 ##### --> Code sections to understand code better

```{r import, echo= 'F',results='hide'}
# CLASSICAL PACKAGE IMPORTER (Stefan Graf & Tobias Frey)
packages_checker <- function(packages_list){
  for (package in packages_list){
    if (!require(package, character.only = T)) {
      install.packages(package)
    }
    library(package, character.only = T)
  }
}

packages_checker(
  c('tmaptools',# read_gpx function
    'plotKML',# read GPX to dataframe
    'lubridate',
    'zoo',
    'dbplyr',
    'plyr',
    'dplyr',
    'tidyr',
    'tidyverse',
    'sf', # Simple feature --> super spatial data handling structure
    'raster', # For raster data
    'ggspatial',
    'rgdal',
    'data.table',
    'RColorBrewer', # color schemas from color brewer web
    'ggraph', # visualization
    'igraph',
    'tmap',
    'scales',
    'trajr',
    'plotly', # zoomable, pannable ggplot
    'mapview',
    'ggmap', # easier interactive ggmaps 
    'leaflet', # Javascript library (originally) for plotting on a maptile interactive map
    'tmap',
    'stringr',
    'remotes', # geom_convexhull function
    'magrittr'
    ))
rm(list = ls())
getwd()

### CUSTOM FUNCTIONS ###
clockS = function(t){hour(t)*3600+minute(t)*60+second(t)}

read_GPX_waypoints <- function(flnm){
    file <- read_GPX(flnm)
    if(typeof(file)!="list"){file <- file$waypoints}
    file<- file %>%  # add file name as an attribute
        mutate(filename = flnm)
  return(out)
}
st_read_GPX_waypoints <- function(flnm){
    out <- tryCatch(
        { message("Try to read file")
            file <- data.frame()
            start <- Sys.time()
            file <- st_read(flnm,layer = "waypoints") # better than read_GPX
            #if(class(file)[1]=="list"){file <- file$waypoints} # only use if read_GPX is used
            end <- Sys.time()
            processduration <- end-start
            file <- file %>% #Skip the first 16 rows because see below
              mutate(fileN=substr(flnm, 33,(nchar(flnm)+1)-5))
            
            return(file)
        },
        error=function(cond) {
            message(paste("File seems not to exist", flnm))
            message("Here's the original error message:")
            message(cond)
            # Choose a return value in case of error
            return(data.frame()) # use dataframe as return instead of NA --> See here why, without it throws an error. https://stackoverflow.com/questions/48512461/error-in-bind-rows-x-id-argument-1-must-have-names-using-map-df-in-purrr
        },
        warning=function(cond) {
            message(paste("Loading file didn't work", flnm))
            message("Here's the original warning message:")
            message(cond)
            # Choose a return value in case of warning
            return(data.frame())
        },
        finally={
            message(paste("Processed URL: ", flnm," Duration: ",processduration))
            message("Finished")
        }
    )    
    return(out)
}



####### Flatten nested lists #######
tl <- function(e) { if (is.null(e)) return(NULL); ret <- typeof(e); if (ret == 'list' && !is.null(names(e))) ret <- list(type='namedlist') else ret <- list(type=ret,len=length(e)); ret; };
mkcsv <- function(v) paste0(collapse=',',v);
keyListToStr <- function(keyList) paste0(collapse='','/',sapply(keyList,function(key) if (is.null(key)) '*' else paste0(collapse=',',key)));

extractLevelColumns <- function(
    nodes, ## current level node selection
    ..., ## additional arguments to data.frame()
    keyList=list(), ## current key path under main list
    sep=NULL, ## optional string separator on which to join multi-element vectors; if NULL, will leave as separate columns
    mkname=function(keyList,maxLen) paste0(collapse='.',if (is.null(sep) && maxLen == 1L) keyList[-length(keyList)] else keyList) ## name builder from current keyList and character vector max length across node level; default to dot-separated keys, and remove last index component for scalars
) {
    cat(sprintf('extractLevelColumns(): %s\n',keyListToStr(keyList)));
    if (length(nodes) == 0L) return(list()); ## handle corner case of empty main list
    tlList <- lapply(nodes,tl);
    typeList <- do.call(c,lapply(tlList,`[[`,'type'));
    if (length(unique(typeList)) != 1L) stop(sprintf('error: inconsistent types (%s) at %s.',mkcsv(typeList),keyListToStr(keyList)));
    type <- typeList[1L];
    if (type == 'namedlist') { ## hash; recurse
        allKeys <- unique(do.call(c,lapply(nodes,names)));
        ret <- do.call(c,lapply(allKeys,function(key) extractLevelColumns(lapply(nodes,`[[`,key),...,keyList=c(keyList,key),sep=sep,mkname=mkname)));
    } else if (type == 'list') { ## array; recurse
        lenList <- do.call(c,lapply(tlList,`[[`,'len'));
        maxLen <- max(lenList,na.rm=T);
        allIndexes <- seq_len(maxLen);
        ret <- do.call(c,lapply(allIndexes,function(index) extractLevelColumns(lapply(nodes,function(node) if (length(node) < index) NULL else node[[index]]),...,keyList=c(keyList,index),sep=sep,mkname=mkname))); ## must be careful to translate out-of-bounds to NULL; happens automatically with string keys, but not with integer indexes
    } else if (type%in%c('raw','logical','integer','double','complex','character')) { ## atomic leaf node; build column
        lenList <- do.call(c,lapply(tlList,`[[`,'len'));
        maxLen <- max(lenList,na.rm=T);
        if (is.null(sep)) {
            ret <- lapply(seq_len(maxLen),function(i) setNames(data.frame(sapply(nodes,function(node) if (length(node) < i) NA else node[[i]]),...),mkname(c(keyList,i),maxLen)));
        } else {
            ## keep original type if maxLen is 1, IOW don't stringify
            ret <- list(setNames(data.frame(sapply(nodes,function(node) if (length(node) == 0L) NA else if (maxLen == 1L) node else paste(collapse=sep,node)),...),mkname(keyList,maxLen)));
        }; ## end if
    } else stop(sprintf('error: unsupported type %s at %s.',type,keyListToStr(keyList)));
    if (is.null(ret)) ret <- list(); ## handle corner case of exclusively empty sublists
    ret;
}; ## end extractLevelColumns()
## simple interface function
flattenList <- function(mainList,...) do.call(cbind,extractLevelColumns(mainList,...));


```

#### SUAQ research site
## Preprocessing
1. Getting data from excel/csv files
2. Getting data from gpx files, clean and order it
3. Merge Files
```{r GPS Processing}
## 0. Loading metadatasets
### 0.1 Orangutans names and information
SUAQ_orangutans<- read_csv2("data/20200806_Orangutan_ID.csv")
#### delete empty collumns of orangutan list
SUAQ_orangutans <- SUAQ_orangutans[,(colSums(is.na(SUAQ_orangutans))<nrow(SUAQ_orangutans))]
#### make list lowercase especially names
SUAQ_orangutans %<>% mutate(`Orangutan name`= tolower(`Orangutan name`))


## 1. Getting data from excel/csv files
##### SUAQ TRACKS 2018 & 2019 #####
SUAQ_waypoints_17_18 <- read_csv2("data/20200415_Master_GPS_Suaq_2017_Mar18_SH.csv")

#### CRS Code: EPSG 32647
#### Create SF Points
SUAQ_waypoints_17_18_sf <-  st_as_sf(SUAQ_waypoints_17_18, 
                          coords = c("Y", "X"), 
                          crs = 4326)

coordinates_tmp <- st_coordinates(SUAQ_waypoints_17_18_sf)
colnames(coordinates_tmp) <- c("E","N")
SUAQ_waypoints_17_18_sf <- cbind(SUAQ_waypoints_17_18_sf,coordinates_tmp)




#### SUAQ TRACKS DATA 2020 ####
## 2. Getting data from gpx files, clean and order it
### 2.1 Loading data from GPX files

#### create a list of all gpx files
suaq_gpx_list<- list.files(path = "./data/20200717_All_stedit/gpx/",pattern = "*.gpx",full.names = T)

####  import gpx files from files list, only retrieving waypoints (no automatically saved trackpoints) and adding the filename as attribute
start_time <- Sys.time()
SUAQ_waypoints_11_20<- suaq_gpx_list%>% 
  map_df(~st_read_GPX_waypoints(.))
end_time <- Sys.time()
print(start_time-end_time)

# Two follows throw an error --> 20200225_Lois_Ahmad_FN, 20200225_Otto_Armas_FN are empty --> deleted from data files (not from delivery files)


### 2.2 Tidying the attributes and values
#### 2.2.1 extract the 5 attributes from filename: filedate, orangutanname, infant name, followername, followtype. "Adress" in the first field is left in the list to identify the start of a new list.
SUAQ_waypoints_11_20 <-
  SUAQ_waypoints_11_20 %>% mutate(
    filedate = str_extract(fileN, "\\d{8}"),
    orangutanname = substr(
      str_extract(fileN, "\\d{8}_[A-Z][a-z]+"),
      10,
      nchar(str_extract(fileN, "\\d{8}_[A-Z][a-z]+"))
    ),
    infantname = substr(
      str_extract(fileN, "_[A-Z][a-z]+(\\([A-Z][a-z]+\\)|[A-Z][a-z]+)"),
      2 + nchar(orangutanname),
      nchar(
        str_extract(fileN, "_[A-Z][a-z]+(\\([A-Z][a-z]+\\)|[A-Z][a-z]+)")
      )
    ),
    followername = substr(
      str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}"),
      2,
      nchar(str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}")) - 3
    ),
    followtype = substr(
      str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}"),
      nchar(str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}")) - 1,
      nchar(str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}"))
    )
)


### 2.2.2 get rid of empty collumns
SUAQ_waypoints_11_20 <-SUAQ_waypoints_11_20 %>% slice(1:nrow(SUAQ_waypoints_11_20)) # BUG: some kind of bug related to the rbind of rows of sf data frames, when applying the function read_gpx with map_df is causing a problem. The sf df are not index-able and dropping the geometry is also not possible
st_geometry(SUAQ_waypoints_11_20) <- SUAQ_waypoints_11_20$geometry

SUAQ_waypoints_11_20 <- SUAQ_waypoints_11_20[,(colSums(is.na(SUAQ_waypoints_11_20))<nrow(SUAQ_waypoints_11_20))]


### 2.2.3 creating automatic Datetime (from gps collumn time or collumn cmt) and manual Datetime (from input user GPS Device and from name of Data file)
SUAQ_waypoints_11_20 <-
  SUAQ_waypoints_11_20 %>%
  mutate(
    cmt = dmy_hms(cmt, tz = "Asia/Pontianak"),
    automaticDatetime = ymd_hms(time, tz = "Asia/Pontianak"),
    manualDatetime = as.POSIXct(paste0(ymd(filedate)," ", str_extract(name, "^\\d{4}")), format = "%Y-%m-%d %H%M",tz="Asia/Pontianak") # matching string start following 4 digits. Nameing datetime but changeing afterwards (to a real datetime). Just for ordering purposes
  )


#### adding automatic datetime if 
SUAQ_waypoints_11_20[is.na(SUAQ_waypoints_11_20$automaticDatetime),] %<>% mutate(automaticDatetime = cmt) # piping forth and back. In documentation its written x %<>% foo %>% bar,  its the same as x <- x %>% foo %>% bar. In this case thats not correct!


SUAQ_waypoints_11_20 %<>%mutate(
    manualDate = date(manualDatetime),
    manualTime = hms::as_hms(manualDatetime),
    automaticDate = date(automaticDatetime),
    automaticTime = hms::as_hms(automaticDatetime),
    name=str_extract(name, "[^\\d{4} ].*") # not matching the digits in the beginning plus whitespace but everything after
  )



### 2.2.4 Extracting information from Name field: PtType--> Type of point: Longcall, Nest etc. See GPS processing information sheet. 

#### To lowercase name collumn
SUAQ_waypoints_11_20<- SUAQ_waypoints_11_20 %>% mutate(name=tolower(name)) # reduces complexity by around 277 combinations to total 897

#### See what are all possible combinations of the name fields (all levels)
SUAQ_wp_namecollumn_levels<- SUAQ_waypoints_11_20 %>% group_by(.$name,.$sym) %>% summarise(count=n())

#### Making naveaid green symbols same as green circles. Probably meant the same.
# baseline for this categorization is the excel "Dateneckdaten" there are the main information on how many special cases there are and how they are handled.

#Comment:
SUAQ_waypoints_11_20 <-
  SUAQ_waypoints_11_20 %>% mutate(
    PtType = ifelse((sym == "Circle, Green" |sym == "Navaid, Green") &
                      (grepl("bangun", name)|
                         grepl("arrive", name)|
                         grepl("jumpa", name)|
                         grepl("ketemu", name)|
                         grepl("found", name)|
                         grepl("jmp", name)|
                         grepl("jmpa", name)),
                    "found", # two falsly classified with navaid symbol
                    "unknown"),
    PtType = ifelse((sym == "Circle, Green" |
                       sym == "Navaid, Green") &
                      (grepl("sp", name) | # all with criteria Green + sp are morning nests not something else parsed with sp.
                       grepl("pagi", name)),
                    "mornnest",
                    PtType),
    PtType = ifelse((sym == "Circle, Green" |
                       sym == "Navaid, Green") &
                      (grepl("lost", name)|
                         grepl("lepas", name)|
                         grepl("dilepas", name)|
                         grepl("hilang", name)),
                    "lost", # two falsly classified with navaid symbol
                    PtType),
    PtType = ifelse((sym == "Circle, Green" |
                       sym == "Navaid, Green") &
                      (grepl("sm", name) | 
                         grepl("malam", name)),
                    "nightnest",
                    PtType),
    PtType = ifelse((sym == "Circle, Green" |
                       sym == "Navaid, Green") &
                      (
                        grepl("ss", name) |
                        grepl("ss1", name) | 
                        grepl("ss2", name) |  
                        grepl("siang", name)
                      ),
                    "daynest",
                    # maximum 3 daynests
                    PtType
    ),
#Task: --> change every pttype variable above to a different attribute. See which are categorized in multiple groups!
    
    PtType = ifelse((sym == "Block, Blue" ) & grepl("lch", name), # Metainformation for some lch and lcg are not considered because there are only written down for very less points(10-20 points). Information about call direction is given in the activity table
                    "long call heard",
                    PtType),
    PtType = ifelse((sym == "Block, Blue" ) & grepl("lcg", name),
                    "long call given",
                    PtType),
    PtType = ifelse((sym == "Block, Red" |
                       sym == "Circle, Red") & grepl("dna", name),
                    "DNA sample taken",
                    PtType),
    PtType = ifelse((sym == "Circle, Blue" |
                   sym == "Navaid, Blue"),
                    # until now these symbols are all right categorized 
                    # two falsly classified with navaid symbol
                    "party",
                    PtType),
    "Food Type / Individual / Direction" = ifelse((sym == "Circle, Blue" |
                   sym == "Navaid, Blue")&
                    grepl(paste(as.vector(SUAQ_orangutans$`Orangutan name`)
, collapse="|"), name),
                    # until now these symbols are all right categorized 
                    # two falsly classified with navaid symbol
                    str_extract(SUAQ_waypoints_11_20$name,gsub("\\?","\\\\?",paste(as.vector(SUAQ_orangutans$`Orangutan name`),collapse="|"))), # retrieving elements where its matchs an orangutan name in the list and getting the last element of the name string description. (Problem is "s?" in orangutanname list because its used in regex of the grepl function and matches more because the questionmark is not escaped.)
                    NA),
    PtType = ifelse(grepl("exp", name),
                    "experiments", # two falsly classified with navaid symbol
                    PtType),
    PtType = ifelse((sym == "Flag" | sym == "Flag, Green"),
                    "tree", # two falsly classified with navaid symbol
                    PtType),
    PtType = ifelse((sym == "Mine")&grepl("tool", name),
                    "tree", # two falsly classified with navaid symbol
                    PtType),
    PtType = ifelse((sym == "Flag, Red"),
                    "range", # two falsly classified with navaid symbol
                    PtType)
)

paste(as.vector(SUAQ_orangutans$`Orangutan name`),collapse="|")
str_match_all("party ellie eden","alain|ellie|alice|amor|aqra|eden|balu")

do <- flattenList(tmp)


my_result <- tmp %>%
  # extract first dataframe from each nested list
  map('[[', 1) %>% 
  # bind rows together
  bind_rows()

tmp<- str_match_all(SUAQ_waypoints_11_20$name,gsub("\\?","\\\\?",paste(as.vector(SUAQ_orangutans$`Orangutan name`),collapse="|")))
agsub("\\?","\\\\?",paste(as.vector(SUAQ_orangutans$`Orangutan name`)
,collapse="|"))
?str_match_all

tmp<- as.data.frame(tmp)

paste0("[",paste(as.character(SUAQ_orangutans$`Orangutan name`)
,collapse="]|['"),"]")

indx<- paste(as.vector(SUAQ_orangutans$`Orangutan name`)
, collapse="|")
text.df1<- SUAQ_waypoints_11_20[grep(indx, SUAQ_waypoints_11_20$name),]
text.df1$text <- str_extract(text.df1$name, indx)
text.df1

indx <- paste(countries, collapse="|")
text.df1 <- text.df[grep(indx, text.df$text),]
text.df1$text <- str_extract(text.df1$text, indx)
text.df1


st_geometry(SUAQ_wp_namecollumn_levels) <- NULL
write_delim(SUAQ_wp_namecollumn_levels ,delim = ";",path = "output/SUAQ_wp_namecollumn_levels.tsv")

tmp <- SUAQ_waypoints_11_20 %>% filter(grepl("mulai ikut",name))
grepl("dna", "name lch")

tmp<- SUAQ_waypoints_11_20 %>% filter(.$PtType=="nightnest")
tmp<- levels(as.factor(SUAQ_waypoints_11_20$name))
tmp1<- as.data.frame(tmp)

tmp_view<- waypoints %>% count(sym) # number of observation per symbol level
tmp<- waypoints %>% filter(is.na(sym))


  points_red[[i]]$PtType <- "unknown"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Navaid, Green" &
                               (grepl("PAGI", points_red[[i]]$Text) |
                                 grepl("SP", points_red[[i]]$Text)))] <- "MORNNEST"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Navaid, Green" &
                                 (grepl("SORE", points_red[[i]]$Text) |
                                    grepl("MALAM", points_red[[i]]$Text) |
                                    grepl("SM", points_red[[i]]$Text)))] <- "NIGHTNEST"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Navaid, Green" &
                                 (grepl("SIANG", points_red[[i]]$Text) |
                                    grepl("SS", points_red[[i]]$Text)|))] <- "DAYNEST"
  
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Block, Blue" &
                                grepl("LCH", points_red[[i]]$Text))] <- "LCH"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Block, Blue" &
                                 grepl("LCG", points_red[[i]]$Text))] <- "LCG"
  
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Flag, Red")] <- "RANGE"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Flag, Green")] <- "TREE"




```
```{r R-specific preparation}
# CRS Code: EPSG 32647
# Create SF Points
suaq_orangutan_waypoints <- waypoints %>% drop_na(lat,lon)
suaq_orangutan_follows_2020_sf <-  st_as_sf(suaq_orangutan_waypoints, 
                          coords = c("lat", "lon"), 
                          crs = 4326)

coordinates_tmp <- st_coordinates(suaq_orangutan_follows_2020_sf)
colnames(coordinates_tmp) <- c("E","N")
suaq_orangutan_follows_2020_sf <- cbind(suaq_orangutan_follows_2020_sf,coordinates_tmp)


##### DATA WRANGLING ######
# Create timelag
suaq_orangutan_follows_2020_sf<- suaq_orangutan_follows_2020_sf %>% 
  group_by(fileN)%>% 
  mutate(timelag=as.numeric(difftime(lead(as.POSIXlt(CreationTime)),as.POSIXlt(CreationTime),units = "secs"))) # change datetime to time, see analysis below

suaq_neg_timelag <- suaq_orangutan_follows_2020_sf %>% filter(timelag<0)

##### Set a GPS Tracking Laufnummer oder Lösung des Zeitproblems ##### 
# Begründung: In den Tracks kommen teilweise Punkte vor welche zu einer viel späteren Zeit gemacht wurden. Oder wurde Zeit falsch gespeichert (bzw. lief umwandlung in Indonesische Zeitzone falsch)
#suaq_orangutan_follows_2020_sf<- suaq_orangutan_follows_2020_sf %>%  arrange(desc(datetime))

#SUAQ_data<- SUAQ_data %>%  arrange(desc(.$CreationTime))

# Laufnummer für GPS Punkt erstellen




## Create DateTime from time column (time is not datetime)
suaq_orangutan_follows_2020_sf <- suaq_orangutan_follows_2020_sf %>%
  mutate(datetime=with_tz(ymd_hms(CreationTime),"Asia/Pontianak"))
suaq_orangutan_follows_2020_sf <- suaq_orangutan_follows_2020_sf %>%
  mutate(date=date(datetime), truetime=hms::as_hms(datetime))

suaq_orangutan_follows_2020_sf %>% ggplot(aes(truetime))+geom_histogram()
suaq_orangutan_follows_2020_sf %>% ggplot(aes(hms::as_hms(CreationTime)))+geom_histogram() # time before conversion

follows_points_in_night<- suaq_orangutan_follows_2020_sf %>% filter(truetime<lubridate::hms("04:00:00")) # probably one outlier and one track in the night. Or wrong time settings

```
## Calculate timelag 
- Corrected time --> datetime, 
```{r Timelag suaq}
# Timelag histogram
summary(suaq_orangutan_follows_2020_sf$timelag)
suaq_orangutan_follows_2020_sf %>% ggplot(aes(suaq_orangutan_follows_2020_sf$timelag))+
  geom_histogram(binwidth = )
suaq_orangutan_follows_2020_sf <- suaq_orangutan_follows_2020_sf %>% filter(timelag<0)

# dates have the same distribution
ketambe_orangutan_tracks_sf %>% ggplot(aes(date_tmp))+geom_histogram()
ketambe_orangutan_tracks_sf %>% ggplot(aes(dateandtime))+geom_histogram()

# change the calculation of timelag to time_tmp
```

### Analysis
## First visuals SUAQ
```{r Visualization Suaq}
##### VISUALIZE ######
# Grouping and visualize it
plot1<- suaq_orangutan_follows_2020_sf %>% 
  group_by(Follow..) %>%
  ggplot(aes(E,N))+
  geom_point(aes(color=suaq_orangutan_follows_2020_sf$Focal),size=0.3)+
  geom_path(aes(color=suaq_orangutan_follows_2020_sf$Focal), alpha = 0.2,size=0.3)
ggplotly(plot1)
# Grouping --> research tracklengths / number of points per track / time intervall tracks
suaq_num_point_follows<- suaq_orangutan_follows_2020_sf %>% 
  group_by(fileN) %>%
  tally()

suaq_num_point_follows_histogram<- suaq_num_point_follows %>% ggplot(aes(n))+
  geom_histogram()+
  xlab("Lenght of follow [num of waypoints]")+
  ylab("num of waypoints")
suaq_num_point_follows_histogram

# Zoom-Panable map of tracks
ggplotly(plot1)

# plot specific track
pal<- brewer.pal(30, "BrBG")

plot_tmp<- suaq_orangutan_follows_2020_sf %>% 
  filter(fileN=='20190113_Lisa(Leon)_Saidi_NN.') %>%
  ggplot(aes(E,N, label=paste(ID,time)))+
  geom_point()+
  geom_path()+
  geom_text(aes(label=paste(ID,truetime)),position = position_nudge(y = -0.0003),size=3)
ggplotly(plot_tmp)


plot_tmp<- suaq_orangutan_follows_2020_sf %>% 
  filter(fileN=='20190113_Lisa(Leon)_Saidi_NN.') %>% 
  arrange(desc(datetime)) %>% 
  ggplot(aes(E,N, label=paste(ID,time)))+
  geom_point()+
  geom_path()+a
  geom_text(aes(label=paste(ID,truetime)),position = position_nudge(y = -0.0003),size=3)
ggplotly(plot_tmp)


# Map with tiles (leaflet)



# create convex hulls on points of each orangutan
suaq_orangutan_chulls_2019 <- suaq_orangutan_follows_2020_sf %>%
  group_by(orangutanname) %>% 
  summarise(geometry = st_combine(geometry) ) %>%
  st_convex_hull()

plot_tmp<- ggplot(suaq_orangutan_chulls_2019) + geom_sf(data = suaq_orangutan_chulls_2019, aes(colour = orangutanname,alpha=1),fill=NA)
ggplotly(plot_tmp)

## next idea homerange per month or year (need more years!)

# suaq_orangutan_chulls_month <- suaq_orangutan_follows_2020_sf %>%
#   group_by(orangutanname) %>%
#   summarise(geometry = st_combine(geometry)) %>%
#   st_convex_hull()
# rownames(suaq_orangutan_chulls_month) <- suaq_orangutan_chulls_month$orangutanname
# 
# suaq_orangutan_chulls_january <- suaq_orangutan_follows_2020_sf %>%
#   filter(between(datetime, as.Date("2019-01-01"),as.Date("2019-01-31"))) %>% 
#   group_by(orangutanname) %>%
#   summarise(geometry = st_combine(geometry)) %>%
#   st_convex_hull()
# colnames(suaq_orangutan_chulls_january)[2]
# class(suaq_orangutan_chulls_january$geometry[1])
# 
# suaq_orangutan_chulls_month<-bind_cols(suaq_orangutan_chulls_month, suaq_orangutan_chulls_january,by = c("orangutanname" = "orangutanname"))
# ?full_join
# 
# suaq_orangutan_chulls %>% ggplot()+
#     geom_polygon(suaq_orangutan_chulls$geometry)
# plot_tmp<- ggplot(suaq_orangutan_chulls) + geom_sf(data = suaq_orangutan_chulls, aes(colour = orangutanname,alpha=1),fill=NA)
# ggplotly(plot_tmp)




```
