---
title: "Projekt"
author: "Stefan Graf"
date: "28 3 2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading environment / libraries

# Todo:

# Data handling
# Zeiten kontrollieren ob bei Originaldatei in den Tracks (anhand Dateiname) teilweise GPS Messungen mit viel späterem Zeitpunkt zwischendrin vorkommen



# Markings:
#### Quality check: --> are quality checks of chunks executed before
########### DELETE ############--> are chunks to delete
## Title --> are titles for sections
### explanations of the following chunk
##### SUAQ TRACKS 2018 & 2019 ##### --> Code sections to understand code better

```{r import, echo= 'F',results='hide'}
# CLASSICAL PACKAGE IMPORTER (Stefan Graf & Tobias Frey)
packages_checker <- function(packages_list){
  for (package in packages_list){
    if (!require(package, character.only = T)) {
      install.packages(package)
    }
    library(package, character.only = T)
  }
}

packages_checker(
  c('tmaptools',# read_gpx function
    'plotKML',# read GPX to dataframe
    'lubridate',
    'zoo',
    'dbplyr',
    'plyr',
    'dplyr',
    'tidyr',
    'tidyverse',
    'sf', # Simple feature --> super spatial data handling structure
    'raster', # For raster data
    'ggspatial',
    'rgdal',
    'data.table',
    'RColorBrewer', # color schemas from color brewer web
    'ggraph', # visualization
    'igraph',
    'tmap',
    'scales',
    'trajr',
    'plotly', # zoomable, pannable ggplot
    'mapview',
    'ggmap', # easier interactive ggmaps 
    'leaflet', # Javascript library (originally) for plotting on a maptile interactive map
    'tmap',
    'stringr',
    'remotes', # geom_convexhull function
    'magrittr'
    ))
rm(list = ls())
getwd()

### CUSTOM FUNCTIONS ###
clockS = function(t){hour(t)*3600+minute(t)*60+second(t)}

read_GPX_waypoints <- function(flnm){
    file <- read_GPX(flnm)
    if(typeof(file)!="list"){file <- file$waypoints}
    file<- file %>%  # add file name as an attribute
        mutate(filename = flnm)
  return(out)
}
st_read_GPX_waypoints <- function(flnm){
    out <- tryCatch(
        { message("Try to read file")
            file <- data.frame()
            start <- Sys.time()
            file <- st_read(flnm,layer = "waypoints") # better than read_GPX
            #if(class(file)[1]=="list"){file <- file$waypoints} # only use if read_GPX is used
            end <- Sys.time()
            processduration <- end-start
            file <- file %>% #Skip the first 16 rows because see below
              mutate(fileN=substr(flnm, 33,(nchar(flnm)+1)-5))
            
            return(file)
        },
        error=function(cond) {
            message(paste("File seems not to exist", flnm))
            message("Here's the original error message:")
            message(cond)
            # Choose a return value in case of error
            return(data.frame()) # use dataframe as return instead of NA --> See here why, without it throws an error. https://stackoverflow.com/questions/48512461/error-in-bind-rows-x-id-argument-1-must-have-names-using-map-df-in-purrr
        },
        warning=function(cond) {
            message(paste("Loading file didn't work", flnm))
            message("Here's the original warning message:")
            message(cond)
            # Choose a return value in case of warning
            return(data.frame())
        },
        finally={
            message(paste("Processed URL: ", flnm," Duration: ",processduration))
            message("Finished")
        }
    )    
    return(out)
}


```

#### SUAQ research site
## Preprocessing
1. Getting data from excel/csv files
2. Getting data from gpx files, clean and order it
3. Merge Files
```{r GPS Processing}
## 1. Getting data from excel/csv files
##### SUAQ TRACKS 2018 & 2019 #####
SUAQ_waypoints_17_18 <- read_csv2("data/20200415_Master_GPS_Suaq_2017_Mar18_SH.csv")

#### CRS Code: EPSG 32647
#### Create SF Points
SUAQ_waypoints_17_18_sf <-  st_as_sf(SUAQ_waypoints_17_18, 
                          coords = c("Y", "X"), 
                          crs = 4326)

coordinates_tmp <- st_coordinates(SUAQ_waypoints_17_18_sf)
colnames(coordinates_tmp) <- c("E","N")
SUAQ_waypoints_17_18_sf <- cbind(SUAQ_waypoints_17_18_sf,coordinates_tmp)


#### SUAQ TRACKS DATA 2020 ####
## 2. Getting data from gpx files, clean and order it
### 2.1 Loading data from GPX files

#### create a list of all gpx files
suaq_gpx_list<- list.files(path = "./data/20200717_All_stedit/gpx/",pattern = "*.gpx",full.names = T)

####  import gpx files from files list, only retrieving waypoints (no automatically saved trackpoints) and adding the filename as attribute
start_time <- Sys.time()
SUAQ_waypoints_11_20<- suaq_gpx_list%>% 
  map_df(~st_read_GPX_waypoints(.))
end_time <- Sys.time()
print(start_time-end_time)

# Two follows throw an error --> 20200225_Lois_Ahmad_FN, 20200225_Otto_Armas_FN are empty --> deleted from data files (not from delivery files)


### 2.2 Tidying the attributes and values
#### 2.2.1 extract the 5 attributes from filename: filedate, orangutanname, infant name, followername, followtype. "Adress" in the first field is left in the list to identify the start of a new list.
SUAQ_waypoints_11_20 <-
  SUAQ_waypoints_11_20 %>% mutate(
    filedate = str_extract(fileN, "\\d{8}"),
    orangutanname = substr(
      str_extract(fileN, "\\d{8}_[A-Z][a-z]+"),
      10,
      nchar(str_extract(fileN, "\\d{8}_[A-Z][a-z]+"))
    ),
    infantname = substr(
      str_extract(fileN, "_[A-Z][a-z]+(\\([A-Z][a-z]+\\)|[A-Z][a-z]+)"),
      2 + nchar(orangutanname),
      nchar(
        str_extract(fileN, "_[A-Z][a-z]+(\\([A-Z][a-z]+\\)|[A-Z][a-z]+)")
      )
    ),
    followername = substr(
      str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}"),
      2,
      nchar(str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}")) - 3
    ),
    followtype = substr(
      str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}"),
      nchar(str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}")) - 1,
      nchar(str_extract(fileN, "_[A-Z][a-z_]+_[A-Z]{2}"))
    )
)


### 2.2.2 get rid of empty collumns
SUAQ_waypoints_11_20 <-SUAQ_waypoints_11_20 %>% slice(1:nrow(SUAQ_waypoints_11_20)) # BUG: some kind of bug related to the rbind of rows of sf data frames, when applying the function read_gpx with map_df is causing a problem. The sf df are not index-able and dropping the geometry is also not possible
st_geometry(SUAQ_waypoints_11_20) <- SUAQ_waypoints_11_20$geometry

SUAQ_waypoints_11_20 <- SUAQ_waypoints_11_20[,(colSums(is.na(SUAQ_waypoints_11_20))<nrow(SUAQ_waypoints_11_20))]


### 2.2.3 creating automatic Datetime (from gps collumn time or collumn cmt) and manual Datetime (from input user GPS Device and from name of Data file)
SUAQ_waypoints_11_20 <-
  SUAQ_waypoints_11_20 %>%
  mutate(
    cmt = dmy_hms(cmt, tz = "Asia/Pontianak"),
    automaticDatetime = ymd_hms(time, tz = "Asia/Pontianak"),
    manualDatetime = as.POSIXct(paste0(ymd(filedate)," ", str_extract(name, "^\\d{4}")), format = "%Y-%m-%d %H%M",tz="Asia/Pontianak") # matching string start following 4 digits. Nameing datetime but changeing afterwards (to a real datetime). Just for ordering purposes
  )


#### adding automatic datetime if 
SUAQ_waypoints_11_20[is.na(SUAQ_waypoints_11_20$automaticDatetime),] %<>% mutate(automaticDatetime = cmt) # piping forth and back. In documentation its written x %<>% foo %>% bar,  its the same as x <- x %>% foo %>% bar. In this case thats not correct!


SUAQ_waypoints_11_20 %<>%mutate(
    manualDate = date(manualDatetime),
    manualTime = hms::as_hms(manualDatetime),
    automaticDate = date(automaticDatetime),
    automaticTime = hms::as_hms(automaticDatetime),
    name=str_extract(name, "[^\\d{4} ].*") # not matching the digits in the beginning plus whitespace but everything after
  )



### 2.2.4 Extracting information from Name field: PtType--> Type of point: Longcall, Nest etc. See GPS processing information sheet. 

#### To lowercase name collumn
SUAQ_waypoints_11_20<- SUAQ_waypoints_11_20 %>% mutate(name=tolower(name)) # reduces complexity by around 277 combinations to total 897

#### See what are all possible combinations of the name fields (all levels)
SUAQ_wp_namecollumn_levels<- SUAQ_waypoints_11_20 %>% group_by(.$name,.$sym) %>% summarise(count=n())

#### Making naveaid green symbols same as green circles. Probably meant the same.
# SUAQ_waypoints_11_20[SUAQ_waypoints_11_20$sym=="Navaid, Green",] %<>% mutate(sym = "Circle, Green")  # all navaid green values controlled should be nest or starting ending point


tmp<- SUAQ_waypoints_11_20 %>% filter(.$name=='EXPL')

SUAQ_waypoints_11_20 <-
  SUAQ_waypoints_11_20 %>% mutate(
    PtType = ifelse((sym == "Circle, Green" |
                       sym == "Navaid, Green") &
                      (grepl("SP", name) | grepl("PAGI", name)),
                    "mornnest",
                    "unknown"),
    PtType = ifelse((sym == "Circle, Green" |
                       sym == "Navaid, Green") &
                      (grepl("SM", name) | grepl("MALAM", name)),
                    "nightnest",
                    PtType),
    PtType = ifelse((sym == "Circle, Green" |
                       sym == "Navaid, Green") &
                      (
                        grepl("SS", name) |
                        grepl("SS1", name) | 
                        grepl("SS2", name) |  
                        grepl("SIANG", name)
                      ),
                    "daynest",
                    # maximum 3 daynests
                    PtType
    ),
    PtType = ifelse((sym == "Block, Blue" |
                       sym == "Circle, Blue" | sym == "Circle, Green" |
                       sym == "Navaid, Green") &
                      (grepl("SM", name) | grepl("MALAM", name)),
                    "nightnest",
                    PtType),
    PtType = ifelse(grepl("lch", name),
                    "long call heard",
                    PtType),
    PtType = ifelse(grepl("lcg", name),
                    "long call given",
                    PtType)
)

write_csv(SUAQ_wp_namecollumn_levels,path = "")
tmp <- SUAQ_waypoints_11_20 %>% filter(grepl("lch", name))
grepl("lch", "name lch")

tmp<- SUAQ_waypoints_11_20 %>% filter(.$PtType=="nightnest")
tmp<- levels(as.factor(SUAQ_waypoints_11_20$name))
tmp1<- as.data.frame(tmp)

tmp_view<- waypoints %>% count(sym) # number of observation per symbol level
tmp<- waypoints %>% filter(is.na(sym))


  points_red[[i]]$PtType <- "unknown"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Navaid, Green" &
                               (grepl("PAGI", points_red[[i]]$Text) |
                                 grepl("SP", points_red[[i]]$Text)))] <- "MORNNEST"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Navaid, Green" &
                                 (grepl("SORE", points_red[[i]]$Text) |
                                    grepl("MALAM", points_red[[i]]$Text) |
                                    grepl("SM", points_red[[i]]$Text)))] <- "NIGHTNEST"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Navaid, Green" &
                                 (grepl("SIANG", points_red[[i]]$Text) |
                                    grepl("SS", points_red[[i]]$Text)|))] <- "DAYNEST"
  
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Block, Blue" &
                                grepl("LCH", points_red[[i]]$Text))] <- "LCH"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Block, Blue" &
                                 grepl("LCG", points_red[[i]]$Text))] <- "LCG"
  
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Flag, Red")] <- "RANGE"
  points_red[[i]]$PtType[which(points_red[[i]]$Symbol == "Flag, Green")] <- "TREE"




```
```{r R-specific preparation}
# CRS Code: EPSG 32647
# Create SF Points
suaq_orangutan_waypoints <- waypoints %>% drop_na(lat,lon)
suaq_orangutan_follows_2020_sf <-  st_as_sf(suaq_orangutan_waypoints, 
                          coords = c("lat", "lon"), 
                          crs = 4326)

coordinates_tmp <- st_coordinates(suaq_orangutan_follows_2020_sf)
colnames(coordinates_tmp) <- c("E","N")
suaq_orangutan_follows_2020_sf <- cbind(suaq_orangutan_follows_2020_sf,coordinates_tmp)


##### DATA WRANGLING ######
# Create timelag
suaq_orangutan_follows_2020_sf<- suaq_orangutan_follows_2020_sf %>% 
  group_by(fileN)%>% 
  mutate(timelag=as.numeric(difftime(lead(as.POSIXlt(CreationTime)),as.POSIXlt(CreationTime),units = "secs"))) # change datetime to time, see analysis below

suaq_neg_timelag <- suaq_orangutan_follows_2020_sf %>% filter(timelag<0)

##### Set a GPS Tracking Laufnummer oder Lösung des Zeitproblems ##### 
# Begründung: In den Tracks kommen teilweise Punkte vor welche zu einer viel späteren Zeit gemacht wurden. Oder wurde Zeit falsch gespeichert (bzw. lief umwandlung in Indonesische Zeitzone falsch)
#suaq_orangutan_follows_2020_sf<- suaq_orangutan_follows_2020_sf %>%  arrange(desc(datetime))

#SUAQ_data<- SUAQ_data %>%  arrange(desc(.$CreationTime))

# Laufnummer für GPS Punkt erstellen




## Create DateTime from time column (time is not datetime)
suaq_orangutan_follows_2020_sf <- suaq_orangutan_follows_2020_sf %>%
  mutate(datetime=with_tz(ymd_hms(CreationTime),"Asia/Pontianak"))
suaq_orangutan_follows_2020_sf <- suaq_orangutan_follows_2020_sf %>%
  mutate(date=date(datetime), truetime=hms::as_hms(datetime))

suaq_orangutan_follows_2020_sf %>% ggplot(aes(truetime))+geom_histogram()
suaq_orangutan_follows_2020_sf %>% ggplot(aes(hms::as_hms(CreationTime)))+geom_histogram() # time before conversion

follows_points_in_night<- suaq_orangutan_follows_2020_sf %>% filter(truetime<lubridate::hms("04:00:00")) # probably one outlier and one track in the night. Or wrong time settings

```
## Calculate timelag 
- Corrected time --> datetime, 
```{r Timelag suaq}
# Timelag histogram
summary(suaq_orangutan_follows_2020_sf$timelag)
suaq_orangutan_follows_2020_sf %>% ggplot(aes(suaq_orangutan_follows_2020_sf$timelag))+
  geom_histogram(binwidth = )
suaq_orangutan_follows_2020_sf <- suaq_orangutan_follows_2020_sf %>% filter(timelag<0)

# dates have the same distribution
ketambe_orangutan_tracks_sf %>% ggplot(aes(date_tmp))+geom_histogram()
ketambe_orangutan_tracks_sf %>% ggplot(aes(dateandtime))+geom_histogram()

# change the calculation of timelag to time_tmp
```

### Analysis
## First visuals SUAQ
```{r Visualization Suaq}
##### VISUALIZE ######
# Grouping and visualize it
plot1<- suaq_orangutan_follows_2020_sf %>% 
  group_by(Follow..) %>%
  ggplot(aes(E,N))+
  geom_point(aes(color=suaq_orangutan_follows_2020_sf$Focal),size=0.3)+
  geom_path(aes(color=suaq_orangutan_follows_2020_sf$Focal), alpha = 0.2,size=0.3)
ggplotly(plot1)
# Grouping --> research tracklengths / number of points per track / time intervall tracks
suaq_num_point_follows<- suaq_orangutan_follows_2020_sf %>% 
  group_by(fileN) %>%
  tally()

suaq_num_point_follows_histogram<- suaq_num_point_follows %>% ggplot(aes(n))+
  geom_histogram()+
  xlab("Lenght of follow [num of waypoints]")+
  ylab("num of waypoints")
suaq_num_point_follows_histogram

# Zoom-Panable map of tracks
ggplotly(plot1)

# plot specific track
pal<- brewer.pal(30, "BrBG")

plot_tmp<- suaq_orangutan_follows_2020_sf %>% 
  filter(fileN=='20190113_Lisa(Leon)_Saidi_NN.') %>%
  ggplot(aes(E,N, label=paste(ID,time)))+
  geom_point()+
  geom_path()+
  geom_text(aes(label=paste(ID,truetime)),position = position_nudge(y = -0.0003),size=3)
ggplotly(plot_tmp)


plot_tmp<- suaq_orangutan_follows_2020_sf %>% 
  filter(fileN=='20190113_Lisa(Leon)_Saidi_NN.') %>% 
  arrange(desc(datetime)) %>% 
  ggplot(aes(E,N, label=paste(ID,time)))+
  geom_point()+
  geom_path()+a
  geom_text(aes(label=paste(ID,truetime)),position = position_nudge(y = -0.0003),size=3)
ggplotly(plot_tmp)


# Map with tiles (leaflet)



# create convex hulls on points of each orangutan
suaq_orangutan_chulls_2019 <- suaq_orangutan_follows_2020_sf %>%
  group_by(orangutanname) %>% 
  summarise(geometry = st_combine(geometry) ) %>%
  st_convex_hull()

plot_tmp<- ggplot(suaq_orangutan_chulls_2019) + geom_sf(data = suaq_orangutan_chulls_2019, aes(colour = orangutanname,alpha=1),fill=NA)
ggplotly(plot_tmp)

## next idea homerange per month or year (need more years!)

# suaq_orangutan_chulls_month <- suaq_orangutan_follows_2020_sf %>%
#   group_by(orangutanname) %>%
#   summarise(geometry = st_combine(geometry)) %>%
#   st_convex_hull()
# rownames(suaq_orangutan_chulls_month) <- suaq_orangutan_chulls_month$orangutanname
# 
# suaq_orangutan_chulls_january <- suaq_orangutan_follows_2020_sf %>%
#   filter(between(datetime, as.Date("2019-01-01"),as.Date("2019-01-31"))) %>% 
#   group_by(orangutanname) %>%
#   summarise(geometry = st_combine(geometry)) %>%
#   st_convex_hull()
# colnames(suaq_orangutan_chulls_january)[2]
# class(suaq_orangutan_chulls_january$geometry[1])
# 
# suaq_orangutan_chulls_month<-bind_cols(suaq_orangutan_chulls_month, suaq_orangutan_chulls_january,by = c("orangutanname" = "orangutanname"))
# ?full_join
# 
# suaq_orangutan_chulls %>% ggplot()+
#     geom_polygon(suaq_orangutan_chulls$geometry)
# plot_tmp<- ggplot(suaq_orangutan_chulls) + geom_sf(data = suaq_orangutan_chulls, aes(colour = orangutanname,alpha=1),fill=NA)
# ggplotly(plot_tmp)




```
