---
title: "SUAQr_homeranges"
author: "Stefan Graf"
date: "2/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Markings:
#### Quality check: --> are quality checks of chunks executed before
########### DELETE ############--> are chunks to delete
## Title --> are titles for sections
### explanations of the following chunk
##### SUAQ TRACKS 2018 & 2019 ##### --> Code sections to understand code better

```{r import, echo= 'F',results='hide'}
# CLASSICAL PACKAGE IMPORTER (Stefan Graf & Tobias Frey)
packages_checker <- function(packages_list){
  for (package in packages_list){
    if (!require(package, character.only = T)) {
      install.packages(package)
    }
    library(package, character.only = T)
  }
}

packages_checker(
  c('tlocoh',
    'tidyr',
    'tmap',
    'tmaptools',# read_gpx function
    'hrbrthemes', # for colors --> viridis
    'rgdal',
    'ctmm',
    'lubridate',
    'plyr',
    'dplyr',
    'tidyr',
    'tidyverse',
    'sf', # Simple feature --> super spatial data handling structure
    'raster', # For raster data
    'rgdal',
    'plotly', # zoomable, pannable ggplot
    'move',# Brownian Bridge Movement Model
    'adehabitatHR',
    'lme4',
    'nlme',
    'ggplot2',
    'ggpubr',
    'GGally',
    'ggfortify', #check linear regression assumptions
    'GGally', # usefull pairs function to create many scatterplots
    'maptools',# used in move package
    'viridis',
    'lmerTest',
    'sp',
    'ks' # used in KDE functions
    ))
#install.packages("tlocoh", dependencies=TRUE, repos=c("http://R-Forge.R-project.org", "http://cran.cnr.berkeley.edu"), type="source")
#require(tlocoh)

rm(list = ls())
getwd()



### CUSTOM FUNCTIONS ###

# Function for displaying scale labels in ggplot not as f.e. 4e+05
fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, scientific = TRUE)
     # quote the part before the exponent to keep all the digits
     l <- gsub("^(.*)e", "'\\1'e", l)
     # turn the 'e+' into plotmath format
     l <- gsub("e", "%*%10^", l)
     # return this as an expression
     parse(text=l)
}


clockS = function(t){hour(t)*3600+minute(t)*60+second(t)}
st_kde <- function(points,cellsize, bandwith, extent = NULL){
  require(MASS)
  require(raster)
  require(sf)
  if(is.null(extent)){
    extent_vec <- st_bbox(points)[c(1,3,2,4)]
  } else{
    extent_vec <- st_bbox(extent)[c(1,3,2,4)]
  }
  
  n_y <- ceiling((extent_vec[4]-extent_vec[3])/cellsize)
  n_x <- ceiling((extent_vec[2]-extent_vec[1])/cellsize)
  
  extent_vec[2] <- extent_vec[1]+(n_x*cellsize)-cellsize
  extent_vec[4] <- extent_vec[3]+(n_y*cellsize)-cellsize
  
  coords <- st_coordinates(points)
  matrix <- kde2d(coords[,1],coords[,2],h = bandwith,n = c(n_x,n_y),lims = extent_vec)
  raster(matrix)
}




#### Custom functions for analysis
clockS = function(t){hour(t)*3600+minute(t)*60+second(t)}


euclid <- function(x1,y1,x2,y2){
  return(sqrt((x1-x2)^2+(y1-y2)^2))
}

turning_angle <- function(x,y,lead_lag = 1){ 
  if(length(x) < 3){return(NA)}
  if(length(x) != length(y)){stop("x and y must be of the same length")}
  p1x <- lag(x,lead_lag)
  p1y <- lag(y,lead_lag)
  p2x <- x
  p2y <- y
  p3x <- lead(x,lead_lag)
  p3y <- lead(y,lead_lag)
  p12 <- euclid(p1x,p1y,p2x,p2y)
  p13 <- euclid(p1x,p1y,p3x,p3y)
  p23 <- euclid(p2x,p2y,p3x,p3y)
  rad <- acos((p12^2+p23^2-p13^2)/(2*p12*p23))
  grad <- (rad*180)/pi
  grad[p12 == 0 | p23 == 0] <- NA
  d <-  (p3x-p1x)*(p2y-p1y)-(p3y-p1y)*(p2x-p1x)
  d <- ifelse(d == 0,1,d)
  d[d>0] <- 1
  d[d<0] <- -1
  d[d==0] <- 1
  turning <- grad*d*-1+180
  return(turning)
}

# Source https://exploratory.io/note/kanaugust/1701090969905358
calculate_mode <- function(x) {
  uniqx <- unique(na.omit(x))
  uniqx[which.max(tabulate(match(x, uniqx)))]
}


# Statistics from here: https://github.com/aufrank/R-hacks/blob/master/mer-utils.R
vif.mer <- function (fit) {
    ## adapted from rms::vif

    v <- vcov(fit)
    nam <- names(fixef(fit))

    ## exclude intercepts
    ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
    if (ns > 0) {
        v <- v[-(1:ns), -(1:ns), drop = FALSE]
        nam <- nam[-(1:ns)]
    }

    d <- diag(v)^0.5
    v <- diag(solve(v/(d %o% d)))
    names(v) <- nam
    v
}

aggregate.sf = function(x, by, FUN, ..., do_union = TRUE, simplify = TRUE,
		join = st_intersects) {

	if (inherits(by, "sf") || inherits(by, "sfc")) {
		if (inherits(by, "sfc"))
			by = st_sf(by)
		i = join(st_geometry(by), st_geometry(x))
		st_geometry(x) = NULL
		# dispatch to stats::aggregate:
		a = aggregate(x[unlist(i), , drop = FALSE],
			list(rep(seq_len(nrow(by)), lengths(i))), FUN, ...)
		nrow_diff = nrow(by) - nrow(a)
		if(nrow_diff > 0) {
			a_na = a[rep(NA, nrow(by)),] # 'top-up' missing rows
			a_na[a$Group.1,] = a
			a = a_na
		}
		a$Group.1 = NULL # remove
		row.names(a) = row.names(by)
		st_set_geometry(a, st_geometry(by))
	} else {
		crs = st_crs(x)
		lst = lapply(split(st_geometry(x), by), function(y) do.call(c, y))
		geom = do.call(st_sfc, lst[!sapply(lst, is.null)])

		if (do_union)
			geom = st_union(st_set_precision(geom, st_precision(x)), by_feature = TRUE)

		st_geometry(x) = NULL
		x = aggregate(x, by, FUN, ..., simplify = simplify)
		st_geometry(x) = geom # coerces to sf
		st_crs(x) = crs

		# now set agr:
		geoms = which(vapply(x, function(vr) inherits(vr, "sfc"), TRUE))
		agr_names = names(x)[-geoms]
		agr = rep("aggregate", length(agr_names))
		names(agr) = agr_names
		# which ones are identity variables?
		n = if (!is.null(names(by)))
			names(by)
		else
			paste0("Group.", seq_along(by))
		agr[n] = "identity"
		st_agr(x) = agr

		x
	}
}

#### read in homeranges from shp list in data f.e. for ctmm results
st_read_homeranges <- function(flnm){
    out <- tryCatch(
        { message("Try to read file")
            file <- data.frame()
            start <- Sys.time()
            file <- st_read(flnm) # better than read_GPX
            file<- st_transform(file,crs = 23847)
            #if(class(file)[1]=="list"){file <- file$waypoints} # only use if read_GPX is used
            end <- Sys.time()
            processduration <- end-start
            file <- file %>% #Skip the first 16 rows because see below
              mutate(fileN=substr(flnm, regexpr("\\/[^\\/]*$", flnm)+1,nchar(flnm)-4)) %>% 
              separate(fileN,c("number","hrprcnt","focal2","skip","model","modeltype"),sep =" ")%>%
              separate(name,c("focal","hrprcnt","statisticalBoundary"),sep =" ") %>% 
              filter(statisticalBoundary=="est") %>% 
              dplyr::select(-skip,-number,-focal2)

            return(file)
        },
        error=function(cond) {
            message(paste("File seems not to exist", flnm))
            message("Here's the original error message:")
            message(cond)
            # Choose a return value in case of error
            return(data.frame()) # use dataframe as return instead of NA --> See here why, without it throws an error. https://stackoverflow.com/questions/48512461/error-in-bind-rows-x-id-argument-1-must-have-names-using-map-df-in-purrr
        },
        warning=function(cond) {
            message(paste("Loading file didn't work", flnm))
            message("Here's the original warning message:")
            message(cond)
            # Choose a return value in case of warning
            return(data.frame())
        },
        finally={
            message(paste("Processed URL: ", flnm," Duration: ",processduration))
            message("Finished")
        }
    )    
    return(out)
}

setwd("/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/")

```

#### SUAQ homerange analysis

```{r GPS Loader,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
### Load Project variables ### 
##### waypoints ##### "data/Processed_data/SUAQ_waypoints_11all20.csv"
##### ATTENTION: read_csv guesses a columns type as boolean if first 1200 entries are NA
# find most recent file
mostrecent <- list.files(path = "data/Processed_data/", full.names = FALSE, recursive = FALSE)
mostrecent <- as.data.frame(mostrecent)
mostrecent <- mostrecent %>% filter(grepl("SUAQ_waypoints_11all20",mostrecent)) %>% 
  mutate(date=as.numeric(str_extract(mostrecent,"\\d{8}"))) %>% arrange(desc(date)) %>% top_n(1,date)

SUAQ_waypoints_11all20_loc <-
  read_csv(paste0("data/Processed_data/",mostrecent$date[1],"_SUAQ_waypoints_11all20.csv"),guess_max = min(10000, 30000))
SUAQ_waypoints_11all20_loc <- SUAQ_waypoints_11all20_loc %>% mutate(monthYearDate=format(as.Date(manualDate), "%Y-%m"))
SUAQ_waypoints_11all20_loc <- SUAQ_waypoints_11all20_loc %>% mutate(automaticDatetime=with_tz(automaticDatetime,tz = "Asia/Pontianak"))
SUAQ_waypoints_11all20_loc <- SUAQ_waypoints_11all20_loc %>% mutate(manualDatetime=with_tz(manualDatetime,tz = "Asia/Pontianak"))

#### add party individuals as individual gps points.
SUAQ_waypoints_11all20_loc_party <- SUAQ_waypoints_11all20_loc %>% filter(PtType%in%c("party"))

SUAQ_waypoints_11all20_loc_party1 <- SUAQ_waypoints_11all20_loc_party %>% 
  dplyr::select(altitude,name_original,sym,fileN,gpsextraction,gpsextractionquality,filedate,followername,manualDatetime, manualTime, manualDate, automaticDatetime, automaticTime,automaticDate,PtType,E,N,AgeSex,follow,PtType_individual_1) %>% filter(!is.na(PtType_individual_1)) %>%  mutate(focal=PtType_individual_1) %>% dplyr::select(-PtType_individual_1) 

SUAQ_waypoints_11all20_loc_party2 <- SUAQ_waypoints_11all20_loc_party %>% 
  dplyr::select(altitude,name_original,sym,fileN,gpsextraction,gpsextractionquality,filedate,followername,manualDatetime, manualTime, manualDate, automaticDatetime, automaticTime,automaticDate,PtType,E,N,AgeSex,follow,PtType_individual_2) %>% filter(!is.na(PtType_individual_2)) %>%  mutate(focal=PtType_individual_2) %>% dplyr::select(-PtType_individual_2) 

#### Combine datasets if for one follow there is an individual at party1 and at party2 position (f.e. if maybe at the first party point another individual was present but at the second on the same follow not) its still gets one new follownumber.
SUAQ_waypoints_11all20_loc_party <- SUAQ_waypoints_11all20_loc_party1 %>% bind_rows(SUAQ_waypoints_11all20_loc_party2)

SUAQ_waypoints_11all20_loc_party <- SUAQ_waypoints_11all20_loc_party%>% 
  group_by(fileN,focal,manualDate,follow) %>% # if .$follow it can be modified later idk why with only "follow" it doesnt work
  mutate(follow = as.numeric(cur_group_id()+1110000)) ## adding FN above 11

SUAQ_waypoints_11all20_loc <- SUAQ_waypoints_11all20_loc %>% 
  bind_rows(SUAQ_waypoints_11all20_loc_party)

##### Weather and FAI #####
SUAQ_weather <- read_csv("data/Processed_data/SUAQ_weather.csv")
SUAQ_fai <- read_csv("data/Processed_data/SUAQ_fai.csv")

SUAQ_weather_monthly<-SUAQ_weather %>% 
  group_by(SUAQ_weather.year,SUAQ_weather.month) %>% 
  summarise_all(funs(mean))
SUAQ_weather_monthly<- SUAQ_weather_monthly%>% dplyr::select(-SUAQ_weather.weather_id,-SUAQ_weather.day,-SUAQ_weather.daymonthyear)

##### location network ##### 
SUAQ_locationnetwork_sf <- st_read("data/20201201_StudyBalimbingStudyArea_points.GPX")
SUAQ_locationnetwork_sf_loc <- st_transform(SUAQ_locationnetwork_sf,crs = 23847)
##### for locale coordinate system
coordinates_tmp <- st_coordinates(SUAQ_locationnetwork_sf_loc)
colnames(coordinates_tmp) <- c("E","N")
SUAQ_locationnetwork_sf_loc <- cbind(SUAQ_locationnetwork_sf_loc,coordinates_tmp)
SUAQ_locationnetwork_loc <- st_drop_geometry(SUAQ_locationnetwork_sf_loc)

SUAQ_pathnetwork<- as.data.frame(read_GPX(file="data/SUAQ_Peta_WithLines.GPX",layers = c("routes")))
colnames(SUAQ_pathnetwork)[15]<- "geometry"
SUAQ_pathnetwork_sf<- st_as_sf(SUAQ_pathnetwork,crs=4326)
SUAQ_pathnetwork_sf_loc <- st_transform(SUAQ_pathnetwork_sf,crs = 23847)

##### physical features for plots
SUAQ_river_sf_wgs<- st_read(dsn="data/external/GeodataIndonesia/Wasser/fluss_50k_perimetergross.shp")
SUAQ_river_sf_loc <- st_transform(SUAQ_river_sf_wgs,crs = 23847)

SUAQ_smallriver_sf_wgs<- st_read(dsn="data/external/GeodataIndonesia/Wasser/kleinflüsse_50k_perimetergross.shp")
SUAQ_smallriver_sf_loc <- st_transform(SUAQ_smallriver_sf_wgs,crs = 23847)

SUAQ_contour_sf_wgs<- st_read(dsn="data/external/GeodataIndonesia/Hoehenllinien/hohenlinien_50k_perimetergross.shp")
SUAQ_contour_sf_loc <- st_transform(SUAQ_contour_sf_wgs,crs = 23847)

SUAQ_station_sf_wgs<- st_read(dsn="data/external/researchstation.shp")
SUAQ_station_sf_loc <- st_transform(SUAQ_station_sf_wgs,crs = 23847)

##### Selection only pointtypes which are taken on the orangutan way? is this true for all? ##### 
SUAQ_rangepoints_11all20_loc <- SUAQ_waypoints_11all20_loc %>% filter(PtType %in% c("daynest","found","lost","mornnest","nightnest","range","tree")) # get rid of lcg lch points and

##### Change trackpoint_id and total num of GPS per follow
###### trackpoint ID 
SUAQ_rangepoints_11all20_loc <- SUAQ_rangepoints_11all20_loc %>% group_by(follow) %>% dplyr::select(-trackpoint_id) %>% mutate(trackpoint_id = row_number())
###### Definitive number of gps per follow
result_totnumfollows <- SUAQ_rangepoints_11all20_loc %>% 
  group_by(follow) %>% summarise(follow_totNumOfGPSpoints=n()) # 1323 Follows
SUAQ_rangepoints_11all20_loc <- SUAQ_rangepoints_11all20_loc %>%dplyr::select(-follow_totNumOfGPSpoints) %>% left_join(result_totnumfollows,by=c("follow"="follow"))
###### orangutan number of points ID 
result_numoffollows<- SUAQ_rangepoints_11all20_loc %>% group_by(follow,focal) %>% summarize(count=n()) %>% group_by(focal) %>% summarize(count=n())
specs_orangutan_tot_num_of_gps <-SUAQ_rangepoints_11all20_loc %>% 
  group_by(focal) %>%
  summarise(n()) %>% left_join(result_numoffollows,by=c('focal'='focal'))
colnames(specs_orangutan_tot_num_of_gps) <- c("focal",
                                           "orangutan_tot_num_of_gps","orangutan_num_of_follows")
tmp<- specs_orangutan_tot_num_of_gps
SUAQ_rangepoints_11all20_loc <-SUAQ_rangepoints_11all20_loc %>% dplyr::select(-orangutan_num_of_follows,-orangutan_tot_num_of_gps) %>% 
  left_join(tmp,by = c("focal" = "focal"))

##### Recalculate edge lengths etc. based on new point selection for rangepoints dataframe
SUAQ_rangepoints_11all20_loc<- SUAQ_rangepoints_11all20_loc %>% 
  group_by(follow)%>% arrange(manualTime,.by_group = TRUE) %>% 
  mutate(timelag=as.numeric(difftime(as.POSIXlt(manualTime),
                                     as.POSIXlt(lag(manualTime)),units 
                                     ="secs")),
         # calculate distance with approach number 1 --> st_distance but to do that we need to calculate lead first because the lead function doesnt work within st distance (not same datatype (df))
         timelag_lead=as.numeric(difftime(as.POSIXlt(lead(manualTime)),
                                          as.POSIXlt(manualTime),units ="secs")),
         manual_distTolast=sqrt((E-lag(E))^2+(N-lag(N))^2),
         manual_distToNext=sqrt((E-lead(E))^2+(N-lead(N))^2),
         speed_last = manual_distTolast/timelag,
         speed_next = manual_distToNext/timelag_lead,
         turningAngle=turning_angle(E,N),
         altitudediff_next=altitude-lead(altitude),
         )





##### WGS84 file
SUAQ_rangepoints_11all20_loc_sf <-  st_as_sf(SUAQ_rangepoints_11all20_loc, 
                          coords = c("E","N"), 
                          crs = 23847) # can also do that in the new sricpts
SUAQ_rangepoints_11all20_loc_wgs <- st_transform(SUAQ_rangepoints_11all20_loc_sf,4326)
coordinates_tmp <- st_coordinates(SUAQ_rangepoints_11all20_loc_wgs)
colnames(coordinates_tmp) <- c("long","lat")
SUAQ_rangepoints_11all20_loc_wgs <- cbind(SUAQ_rangepoints_11all20_loc_wgs,coordinates_tmp)
SUAQ_rangepoints_11all20_loc_wgs<- st_drop_geometry(SUAQ_rangepoints_11all20_loc_wgs)


##### Selection only females / most tracked females ##### 
SUAQ_rangepoints_11all20_loc_females <-
  SUAQ_rangepoints_11all20_loc %>%   filter(!(SUAQ_orangutans.Sex == "male")) %>% ungroup()

SUAQ_rangepoints_11all20_loc_above50GPS <- SUAQ_rangepoints_11all20_loc %>% filter(orangutan_tot_num_of_gps>250) %>% 
  filter(!(ClassFocal%in%c("character(0)","infant","juvenile")))
levels(as.factor(SUAQ_rangepoints_11all20_loc$ClassFocal))
SUAQ_rangepoints_11all20_loc_females_above50GPS <- SUAQ_rangepoints_11all20_loc %>% filter(!(SUAQ_orangutans.Sex == "male")) %>% filter(orangutan_tot_num_of_gps>50)

SUAQ_rangepoints_11all20_loc_males_above50GPS <- SUAQ_rangepoints_11all20_loc %>% filter((SUAQ_orangutans.Sex == "male")) %>% filter(orangutan_tot_num_of_gps>50)

SUAQ_rangepoints_11all20_loc_females_10 <-
  SUAQ_rangepoints_11all20_loc %>%   filter(!(SUAQ_orangutans.Sex == "male")) %>% filter(
    focal %in% c(
      "lisa",
      "friska",
      "ellie",
      "cissy",
      "lilly",
      "yulia",
      "raffi",
      "sarabi",
      "trident",
      "tiara"
    )
  )%>% ungroup()

#### Read in most recent bandwiths for KDE if not newly calculated
mostrecent <- list.files("/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/output/Homeranges/KDE/bandwiths/", full.names = FALSE, recursive = FALSE)
mostrecent <- as.data.frame(mostrecent)
colnames(mostrecent) <- c("date")
mostrecent <- mostrecent %>% filter(date!="_archive") %>%  arrange(desc(date)) %>% top_n(1,date)
imdir <-paste("/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/output/Homeranges/KDE/bandwiths/")
bandwiths <- read_csv(paste0(imdir,paste(mostrecent$date[1], sep="")))

#### Homerange calculations parameter
### Extent
# use always same extent
extentmargin <- 500
axisshifty <- 125 # shift of axis in plots y axis
axisshiftx <- -280 # shift of axis in plots x axis
legendoffsetx <- 0
legendoffsety <- 0
xmin <- min(SUAQ_rangepoints_11all20_loc_females_10$E)-extentmargin 
xmax <- max(SUAQ_rangepoints_11all20_loc_females_10$E)+extentmargin
ymin <- min(SUAQ_rangepoints_11all20_loc_females_10$N)-extentmargin 
ymax <- max(SUAQ_rangepoints_11all20_loc_females_10$N)+extentmargin
step <- 200
xmin_xmax_round_steps<- c(step*(ceiling(xmin/step)):(round(xmax/step)))
xmin_xmax_round_steps <- xmin_xmax_round_steps[2:(length(xmin_xmax_round_steps)-2)] #manual adjustment of tick breaks how many does it need
ymin_ymax_round_steps<- c(step*(ceiling(ymin/step)):(round(ymax/step)))
ymin_ymax_round_steps <- ymin_ymax_round_steps[2:(length(ymin_ymax_round_steps)-1)] #manual adjustment of tick breaks how many does it need

#### research periods data as list
# Split the data into 4 sampling periods. Calculate the respective HR.
SUAQ_rangepoints_11all20_loc_females_10 <- SUAQ_rangepoints_11all20_loc_females_10 %>% mutate(year=year(manualDatetime))

# make 5 lists for all 4 research periods and one for all data
period_list <- split(SUAQ_rangepoints_11all20_loc_females_10, f = SUAQ_rangepoints_11all20_loc_females_10$researchperiod)

# add all data as the 5th list (period)
period_list[["total research period"]] <- SUAQ_rangepoints_11all20_loc_females_10
period_list <- period_list[c("total research period","before 2012","2012 until 2016","2016 until August 2018","Since August 2018")]

# make lists for every year
year_list <- split(SUAQ_rangepoints_11all20_loc_females_10, f = SUAQ_rangepoints_11all20_loc_females_10$year)

#### Selected animals as list/vector
selected_females <- SUAQ_rangepoints_11all20_loc_females_10 %>% group_by(focal) %>% summarise()
selected_females <- as_vector(selected_females)


```

Homerange sizes:

# Homerange calculations
## MCP
```{r MCP - HR per-researchperiods and all ,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/MCP/periods/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/MCP/periods/",format(Sys.Date(), "%Y%m%d"), "/", sep="")


mcp_list <- data.frame()

for (samplinginterval in c(1:length(period_list))){
  print(paste("mcp for period: ",samplinginterval))
  ##### Determine the smoothing parameters
  selected_females <- period_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
  selected_females <- as_vector(selected_females)
  
  iterator <- 1
  for (i in selected_females) { # do the following for every animal in the list
    print(paste("mcp for focal: ",i))
    n_follows <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
        group_by(follow) %>% summarise() %>% nrow() # just as a variable for later results dataframe --> number of follows

    xy_selected <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
      dplyr::select(E,N)
    n_gps <- xy_selected %>% nrow() # just as a variable for later results dataframe --> number of GPS points
    
    coordinates(xy_selected) <- c("E","N")
    proj4string(xy_selected) <- CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" )
    
    # check if enough gps points are still here otherwise make NA's
    xy_selected_toolessdata <- TRUE
    if(length(xy_selected)<30){xy_selected_toolessdata <- FALSE}
    
    iterator <- iterator+1
    
      if(xy_selected_toolessdata){# only execute when enough data
      hr100<- mcp(xy_selected,percent = 100)
      hr99 <- mcp(xy_selected,percent = 99)
      hr95 <- mcp(xy_selected,percent = 95)
      hr50 <- mcp(xy_selected,percent = 50)
      hr100_area <- mcp.area(xy_selected, percent = 100,unout = "km2")
      hr99_area <- mcp.area(xy_selected, percent = 99,unout = "km2")
      hr95_area <- mcp.area(xy_selected, percent = 95,unout = "km2")
      hr50_area <- mcp.area(xy_selected, percent = 50,unout = "km2")
      hr100_sf<- st_as_sf(hr100)
      hr99_sf<- st_as_sf(hr99)
      hr95_sf<- st_as_sf(hr95)
      hr50_sf<- st_as_sf(hr50)
      
      png(paste(imdir,i, "_MCP_UD_",names(period_list)[samplinginterval], ".png", sep=""), width=900, height=700, res=120, units="px")
      plot(xy_selected,pch = 16,cex=0.3,fill="darkolivegreen", xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1)))
      plot(hr50, lty=3, lwd=2.5, border="coral1", add=T,axes=F)
      plot(hr95, lty=1, lwd=2, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=0.8, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkred", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, " : ",names(period_list)[samplinginterval], sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()
      }
      

    ########## export results ############
    if (xy_selected_toolessdata) {
      resultList_50 <-
        data.frame(
          focal = i,
          homerange = hr50_area$a,
          hrpercentage = "50%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(period_list)[samplinginterval],
          geometry= hr50_sf$geometry
        )
      resultList_95 <-
        data.frame(
          focal = i,
          homerange = hr95_area$a,
          hrpercentage = "95%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(period_list)[samplinginterval],
          geometry= hr95_sf$geometry
        )
      resultList_99 <-
        data.frame(
          focal = i,
          homerange = hr99_area$a,
          hrpercentage = "99%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(period_list)[samplinginterval],
          geometry= hr99_sf$geometry
        )
      resultList_100 <-
        data.frame(
          focal = i,
          homerange = hr100_area$a,
          hrpercentage = "100%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(period_list)[samplinginterval],
          geometry= hr100_sf$geometry
        )
    }
    if (!xy_selected_toolessdata) {
      resultList_50 <-
        data.frame(
          focal = i,
          homerange = NA,
          hrpercentage = "50%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(period_list)[samplinginterval],
          geometry= NA
        )
      resultList_95 <-
        data.frame(
          focal = i,
          homerange = NA,
          hrpercentage = "95%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(period_list)[samplinginterval],
          geometry= NA
        )
      resultList_99 <-
        data.frame(
          focal = i,
          homerange = NA,
          hrpercentage = "99%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(period_list)[samplinginterval],
          geometry= NA
        )
      resultList_100 <-
        data.frame(
          focal = i,
          homerange = NA,
          hrpercentage = "100%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(period_list)[samplinginterval],
          geometry= NA
        )
    }
    mcp_list <- mcp_list %>% rbind(resultList_50,resultList_95,resultList_99,resultList_100)
    print(paste(i,": Calculation finished", sep=""))
  }
}

st_write(mcp_list, dsn = paste(
  paste(
    imdir,
    "mcp_values_allperiods",
    sep = ""
  ),
  "shp",
  sep = "."
))
#system("say alle minimum konvex Polygone berechnet für die verschiedenen Untersuchungsperioden")


```

```{r MCP - HR per-year ,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/MCP/years/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/MCP/years/",format(Sys.Date(), "%Y%m%d"), "/", sep="")


mcp_list <- data.frame() # results

for (samplinginterval in c(1:length(year_list))){
  print(paste("mcp for period: ",samplinginterval))
  ##### Determine the smoothing parameters
  selected_females <- year_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
  selected_females <- as_vector(selected_females)
  
  iterator <- 1
  for (i in selected_females) { # do the following for every animal in the list
    print(paste("mcp for focal: ",i))
    n_follows <- year_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
        group_by(follow) %>% summarise() %>% nrow() # just as a variable for later results dataframe --> number of follows

    xy_selected <- year_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
      dplyr::select(E,N)
    n_gps <- xy_selected %>% nrow() # just as a variable for later results dataframe --> number of GPS points
    
    coordinates(xy_selected) <- c("E","N")
    proj4string(xy_selected) <- CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" )
    
    # check if enough gps points are still here otherwise make NA's
    xy_selected_toolessdata <- TRUE
    if(length(xy_selected)<30){xy_selected_toolessdata <- FALSE}
    
    iterator <- iterator+1
    
      if(xy_selected_toolessdata){# only execute when enough data
      hr100<- mcp(xy_selected,percent = 100)
      hr99 <- mcp(xy_selected,percent = 99) 
      hr95 <- mcp(xy_selected,percent = 95)
      hr50 <- mcp(xy_selected,percent = 50)
      hr100_area <- mcp.area(xy_selected, percent = 100,unout = "km2")
      hr99_area <- mcp.area(xy_selected, percent = 99,unout = "km2")
      hr95_area <- mcp.area(xy_selected, percent = 95,unout = "km2")
      hr50_area <- mcp.area(xy_selected, percent = 50,unout = "km2")
  
      
      png(paste(imdir,i, "_MCP_UD_",names(year_list)[samplinginterval], ".png", sep=""), width=900, height=700, res=120, units="px")
      plot(xy_selected,pch = 16,cex=0.3,fill="darkolivegreen", xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1)))
      plot(hr50, lty=3, lwd=2.5, border="coral1", add=T,axes=F)
      plot(hr95, lty=1, lwd=2, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=2, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, " : ",names(year_list)[samplinginterval], sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()
      }
      

    ########## export results ############
    if (xy_selected_toolessdata) {
      resultList_50 <-
        data.frame(
          focal = i,
          homerange = hr50_area$a,
          hrpercentage = "50%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(year_list)[samplinginterval],
          geometry= hr50_sf$geometry
        )
      resultList_95 <-
        data.frame(
          focal = i,
          homerange = hr95_area$a,
          hrpercentage = "95%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(year_list)[samplinginterval],
          geometry= hr95_sf$geometry
        )
      resultList_99 <-
        data.frame(
          focal = i,
          homerange = hr99_area$a,
          hrpercentage = "99%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(year_list)[samplinginterval],
          geometry= hr99_sf$geometry
        )
      resultList_100 <-
        data.frame(
          focal = i,
          homerange = hr100_area$a,
          hrpercentage = "100%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(year_list)[samplinginterval],
          geometry= hr100_sf$geometry
        )
    }
    if (!xy_selected_toolessdata) {
      resultList_50 <-
        data.frame(
          focal = i,
          homerange = NA,
          hrpercentage = "50%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(year_list)[samplinginterval],
          geometry= NA
        )
      resultList_95 <-
        data.frame(
          focal = i,
          homerange = NA,
          hrpercentage = "95%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(year_list)[samplinginterval],
          geometry= NA
        )
      resultList_99 <-
        data.frame(
          focal = i,
          homerange = NA,
          hrpercentage = "99%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(year_list)[samplinginterval],
          geometry= NA
        )
      resultList_100 <-
        data.frame(
          focal = i,
          homerange = NA,
          hrpercentage = "100%",
          n_gps = n_gps,
          n_follows = n_follows,
          period = names(year_list)[samplinginterval],
          geometry= NA
        )
    }
    mcp_list <- mcp_list %>% rbind(resultList_50,resultList_95,resultList_99,resultList_100)
    print(paste(i,": Calculation finished", sep=""))
  }
}

st_write(mcp_list, dsn = paste(
  paste(
    imdir,
    "mcp_values_years",
    sep = ""
  ),
  "shp",
  sep = "."
))
system("say alle minimum konvex Polygone berechnet für jedes Jahr")


```

```{r MCP -  HR sizes with bootstrap ,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
### Homerange Stability
##### Using MCP --> The latter plot probably shows more the usage then the stability because mcp.area takes the points away based on their spatial location
mcp.area(SUAQ_rangepoints_11all20_loc_females_more50GPS.sp, percent = seq(30, 90, by = 5))

##### FEMALES: Using MCP bootstrapping, randomly selecting points
SUAQ_move_selected <- SUAQ_rangepoints_11all20_loc_females_above50GPS[order(SUAQ_rangepoints_11all20_loc_females_above50GPS$focal,SUAQ_rangepoints_11all20_loc_females_above50GPS$manualDatetime , decreasing = FALSE ),] %>% ungroup()

SUAQ_move_selected <- move(x=SUAQ_move_selected$E, y=SUAQ_move_selected$N, 
              time=as.POSIXct(SUAQ_move_selected$manualDatetime, format="%Y-%m-%d %H:%M:%OS", tz="Asia/Pontianak"),
              proj=CRS("+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs"), 
              data=SUAQ_move_selected, animal=SUAQ_move_selected$focal, sensor="gps")
hrBootstrap(x=SUAQ_move_selected, rep=25,level=100,
            levelMax=100, unin="m",unout='km2')

##### Exkurs MALES: Using MCP bootstrapping, randomly selecting points
SUAQ_move_selected <- SUAQ_rangepoints_11all20_loc_above50GPS[order(SUAQ_rangepoints_11all20_loc_above50GPS$focal,SUAQ_rangepoints_11all20_loc_above50GPS$manualDatetime , decreasing = FALSE ),] %>% ungroup()

SUAQ_move_selected <- move(x=SUAQ_move_selected$E, y=SUAQ_move_selected$N, 
              time=as.POSIXct(SUAQ_move_selected$manualDatetime, format="%Y-%m-%d %H:%M:%OS", tz="Asia/Pontianak"),
              proj=CRS("+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs"), 
              data=SUAQ_move_selected, animal=SUAQ_move_selected$focal, sensor="gps")
all<- hrBootstrap(x=SUAQ_move_selected, rep=25,level=100,plot=FALSE,
            levelMax=100, unin="m",unout='km2')
all_flatten<- data.frame(do.call(rbind, all))
all_flatten$name<- rownames(all_flatten)
all_flatten <- all_flatten %>% rowwise() %>%  mutate(focal=strsplit(name,"\\.")[[1]][1],points=as.numeric(strsplit(name,"\\.")[[1]][2])) %>% 
  dplyr::select(-name) %>% left_join(SUAQ_orangutans[,c(1,3)],by=c("focal"="orangutanname")) 

all_flatten %>% ggplot()+geom_path(aes(x=points,X50.,colour=focal,linetype=Sex)) + scale_x_log10()+
  geom_path(aes(x=points,X50.,colour=focal,linetype=Sex))+
  geom_path(aes(x=points,X0.,colour=focal,linetype=Sex,alpha=0.5))+
  geom_path(aes(x=points,X100.,colour=focal,linetype=Sex,alpha=0.5))
```

## KDE
```{r KDE - parametrization,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
##### Determine the smoothing parameters
bandwiths <-  data.frame(focal=as_vector(selected_females),
                 hpi=vector(mode="numeric", length=10),hbcv=vector(mode="numeric", length=10),hscv=vector(mode="numeric", length=10),
                 stringsAsFactors=FALSE)
iterator <- 1
for (i in selected_females){
  print(paste("calculating bandwiths for focal: ",i))
  xy_selected <- SUAQ_rangepoints_11all20_loc_females_10 %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  colnames(xy_selected) <- c("x","y") #adjust column names
  hpi <- sqrt(prod(sqrt(diag(Hpi.diag(xy_selected[,1:2], nstage=2))))) # using diagonal bandwith not full bandwith? why?
  hbcv <- sqrt(prod(sqrt(diag(Hbcv.diag(xy_selected[,1:2], whichbcv=1)))))
  hscv <- sqrt(prod(sqrt(diag(Hscv.diag(xy_selected[,1:2], nstage=2)))))
  bandwiths[iterator,"focal"] <- i
  bandwiths[iterator,"hpi"] <- as.numeric(hpi)
  bandwiths[iterator,"hbcv"] <- as.numeric(hbcv)
  bandwiths[iterator,"hscv"] <- as.numeric(hscv)

  iterator <- iterator+1
}
rownames(bandwiths) <- c(1:10)

write_csv(bandwiths, path = paste(
  paste(
    "/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/output/Homeranges/KDE/bandwiths/",
    format(Sys.Date(), "%Y%m%d"),
    "_KDE_bandwiths",
    sep = ""
  ),
  "csv",
  sep = "."
))
```

```{r KDE - HR sizes with bootstrap sampling amount,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
#bandwiths <- read_csv(paste0("data/Processed_data/",str_remove_all(Sys.Date(), "-"),"_KDE_bandwiths.csv"))

dir.create(paste(
    "output/Homeranges/KDE/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))


### KDE-stability:
for(iter in c(2:30)){
amountofdata <- seq(0.5,10,by=0.5)/10
kde95_bootstrap <- data.frame(focal=as_vector(selected_females),
                 HSCV_area_5=vector(mode="numeric", length=10),             
                 HSCV_area_10=vector(mode="numeric", length=10),
                 HSCV_area_15=vector(mode="numeric", length=10),
                 HSCV_area_20=vector(mode="numeric", length=10),
                 HSCV_area_25=vector(mode="numeric", length=10),
                 HSCV_area_30=vector(mode="numeric", length=10),
                 HSCV_area_35=vector(mode="numeric", length=10),
                 HSCV_area_40=vector(mode="numeric", length=10),
                 HSCV_area_45=vector(mode="numeric", length=10),
                 HSCV_area_50=vector(mode="numeric", length=10),
                 HSCV_area_55=vector(mode="numeric", length=10),
                 HSCV_area_60=vector(mode="numeric", length=10),
                 HSCV_area_65=vector(mode="numeric", length=10),
                 HSCV_area_70=vector(mode="numeric", length=10),
                 HSCV_area_75=vector(mode="numeric", length=10),
                 HSCV_area_80=vector(mode="numeric", length=10),
                 HSCV_area_85=vector(mode="numeric", length=10),
                 HSCV_area_90=vector(mode="numeric", length=10),
                 HSCV_area_95=vector(mode="numeric", length=10),
                 HSCV_area_100=vector(mode="numeric", length=10),
                 stringsAsFactors=FALSE)
kde50_bootstrap <- data.frame(focal=as_vector(selected_females),
                 HSCV_area_5=vector(mode="numeric", length=10),             
                 HSCV_area_10=vector(mode="numeric", length=10),
                 HSCV_area_15=vector(mode="numeric", length=10),
                 HSCV_area_20=vector(mode="numeric", length=10),
                 HSCV_area_25=vector(mode="numeric", length=10),
                 HSCV_area_30=vector(mode="numeric", length=10),
                 HSCV_area_35=vector(mode="numeric", length=10),
                 HSCV_area_40=vector(mode="numeric", length=10),
                 HSCV_area_45=vector(mode="numeric", length=10),
                 HSCV_area_50=vector(mode="numeric", length=10),
                 HSCV_area_55=vector(mode="numeric", length=10),
                 HSCV_area_60=vector(mode="numeric", length=10),
                 HSCV_area_65=vector(mode="numeric", length=10),
                 HSCV_area_70=vector(mode="numeric", length=10),
                 HSCV_area_75=vector(mode="numeric", length=10),
                 HSCV_area_80=vector(mode="numeric", length=10),
                 HSCV_area_85=vector(mode="numeric", length=10),
                 HSCV_area_90=vector(mode="numeric", length=10),
                 HSCV_area_95=vector(mode="numeric", length=10),
                 HSCV_area_100=vector(mode="numeric", length=10),
                 stringsAsFactors=FALSE)


for(per in amountofdata){
  iterator <- 1
for (i in selected_females){ # compute KDE home range for every individual
  xy_selected <- SUAQ_rangepoints_11all20_loc_females_10 %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  xy_selected<- sample_n(xy_selected, round(nrow(xy_selected)*per))

  colnames(xy_selected) <- c("x","y") #adjust column names
  xy_selected <- SpatialPoints(xy_selected,proj4string = CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" )) # Converts SpatialPointDataFrame to SpatialPoint (only x & y)

  
  resultList <- list()
  methListTitle <- list("SCV")
  methList <- list("HSCV")
  
  
  extVal <- 0.4 #0.3 How much bigger the UD should be calculated than the research are (minimum bounding box). The higher the more Computation
  gridVal <- 800 #700 How much cells or pixels should be considered for UD calculation. The higher the more Computation
  
  for (j in seq(1, length(methList))) { # compute one home range for each bandwidth method and individual
    if (j == 1) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hscv"])) 
    }
    else {
      print("Error")
    }
    
    hr95 <- getverticeshr(UD, percent=95,unin = c("m"),unout = c("km2"))
    hr50 <- getverticeshr(UD, percent=50,unin = c("m"),unout = c("km2"))
    
    ########## export results #########
    resultList <- list(hr50=hr50, hr95=hr95, UD=UD)
    assign(methList[[j]], resultList) #resultList is renamed according to the ith element of methList
  }
  animRes <- list(HSCV=HSCV)
  assign(selected_females[[iterator]], animRes) # is renamed according to the ith element of animList
  print(paste(selected_females[[iterator]], ": Calculation finished with Subset of ",as.character(100*per),"% of Points", sep=""))
  
  iterator <- iterator+1
}
  kde_list <- list(lisa=lisa,
      friska=friska,
      ellie=ellie,
      cissy=cissy,
      lilly=lilly,
      yulia=yulia,
      raffi=raffi,
      sarabi=sarabi,
      trident=trident,
      tiara=tiara)
  itera <- 1
for(n in as_vector(selected_females)){
  kde50_bootstrap[itera,paste("HSCV_area","_",as.character(100*per),sep = "")] <-  kde_list[[n]]$HSCV$hr50$area
  kde95_bootstrap[itera,paste("HSCV_area","_",as.character(100*per),sep = "")] <-  kde_list[[n]]$HSCV$hr95$area
  itera <- itera+1
}
}
colnames(kde95_bootstrap) <- c("focal",as.character(seq(0.5,10,0.5)*10))
kde95_bootstrap_long_iter<- gather(kde95_bootstrap,as.character(seq(0.5,10,0.5)*10),key="subset %",value = "HR [km2]")
kde95_bootstrap_long_iter <- kde95_bootstrap_long_iter %>% mutate(`subset %`=as.numeric(`subset %`))
kde95_bootstrap_long_iter <- kde95_bootstrap_long_iter %>% mutate(iteration=iter)
kde95_bootstrap_long <- kde95_bootstrap_long %>% rbind(kde95_bootstrap_long_iter)

colnames(kde50_bootstrap) <- c("focal",as.character(seq(0.5,10,0.5)*10))
kde50_bootstrap_long_iter<- gather(kde50_bootstrap,as.character(seq(0.5,10,0.5)*10),key="subset %",value = "HR [km2]")
kde50_bootstrap_long_iter <- kde50_bootstrap_long_iter %>% mutate(`subset %`=as.numeric(`subset %`))
kde50_bootstrap_long_iter <- kde50_bootstrap_long_iter %>% mutate(iteration=iter)
kde50_bootstrap_long <- kde50_bootstrap_long %>% rbind(kde50_bootstrap_long_iter)
}




colnames(kde95_bootstrap) <- c("focal",as.character(seq(0.5,10,0.5)*10))
kde95_bootstrap_long<- gather(kde95_bootstrap,as.character(seq(0.5,10,0.5)*10),key="subset %",value = "HR [km2]")
kde95_bootstrap_long <- kde95_bootstrap_long %>% mutate(`subset %`=as.numeric(`subset %`))
kde95_bootstrap_long <- kde95_bootstrap_long %>% mutate(iteration=1)

colnames(kde50_bootstrap) <- c("focal",as.character(seq(0.5,10,0.5)*10))
kde50_bootstrap_long<- gather(kde50_bootstrap,as.character(seq(0.5,10,0.5)*10),key="subset %",value = "HR [km2]")
kde50_bootstrap_long <- kde50_bootstrap_long %>% mutate(`subset %`=as.numeric(`subset %`))
kde50_bootstrap_long <- kde50_bootstrap_long %>% mutate(iteration=1)



kde50_bootstrap_long %>%  ggplot(aes(`subset %`,`HR [km2]`,colour=focal)) + geom_line()
rownames(kde) <- c(1:10)
kde95_bootstrap_long %>%  ggplot(aes(`subset %`,`HR [km2]`,colour=focal)) + geom_point(size=0.3)+geom_jitter(size=0.3)
rownames(kde) <- c(1:10)
```

```{r KDE - HR sizes with bootstrap temporal amount,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
# Split the data into 4 sampling periods. Calculate the respective HR.
monthyear_list <- split(SUAQ_rangepoints_11all20_loc_females_10, f = SUAQ_rangepoints_11all20_loc_females_10$monthYearDate)

dir.create(paste(
    "output/Homeranges/KDE/temporalbootstrap/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/KDE/temporalbootstrap/",format(Sys.Date(), "%Y%m%d"), "/", sep="")

kde_results <- data.frame() # results
for (samplinginterval in c(1:length(monthyear_list))){
print(paste("kdes for month: ",samplinginterval))
##### Determine the smoothing parameters
selected_females <- data.frame()
for(l in c(1:samplinginterval)){selected_females <-selected_females %>% rbind(monthyear_list[[l]])}
selected_females <- selected_females%>% group_by(focal) %>% summarise()
selected_females <- as_vector(selected_females)

bandwiths <-  data.frame(focal=as_vector(selected_females),
                 hpi=vector(mode="numeric", length=length(selected_females)),hbcv=vector(mode="numeric", length=length(selected_females)),hscv=vector(mode="numeric", length=length(selected_females)),
                 stringsAsFactors=FALSE)
iterator <- 1
for (i in selected_females){
  print(paste("kdes for focal: ",i))
  xy_selected <- data.frame()
  for(l in c(1:samplinginterval)){xy_selected <-xy_selected %>% rbind(monthyear_list[[l]])}
  xy_selected <- SUAQ_rangepoints_11all20_loc_females_10 %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  colnames(xy_selected) <- c("x","y") #adjust column names
  hpi <- sqrt(prod(sqrt(diag(Hpi.diag(xy_selected[,1:2], nstage=2))))) # using diagonal bandwith not full bandwith? why?
  hbcv <- sqrt(prod(sqrt(diag(Hbcv.diag(xy_selected[,1:2], whichbcv=1)))))
  hscv <- sqrt(prod(sqrt(diag(Hscv.diag(xy_selected[,1:2], nstage=2)))))
  bandwiths[iterator,"focal"] <- i
  bandwiths[iterator,"hpi"] <- as.numeric(hpi)
  bandwiths[iterator,"hbcv"] <- as.numeric(hbcv)
  bandwiths[iterator,"hscv"] <- as.numeric(hscv)

  iterator <- iterator+1
}
rownames(bandwiths) <- c(1:length(selected_females))

### KDE: Calculate UD, HR
iterator <- 1
results_focal <- data.frame()
for (i in selected_females){ # compute KDE home range for every individual
  xy_selected <- data.frame()
  for(l in c(1:samplinginterval)){xy_selected <-xy_selected %>% rbind(monthyear_list[[l]])}
  n_follows <- xy_selected %>% filter(focal%in%c(i)) %>% group_by(follow) %>% summarise() %>% nrow()
  xy_selected <- xy_selected %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  n_gps <- xy_selected %>% nrow()
  
  colnames(xy_selected) <- c("x","y") #adjust column names
  resultList <- list()
  methListTitle <- list("REF", "PI", "BCV", "SCV")
  methList <- list("HREF", "HPI", "HBCV", "HSCV")
  
  xy_selected <- SpatialPoints(xy_selected,proj4string = CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" )) # Converts SpatialPointDataFrame to SpatialPoint (only x & y)
  
  extVal <- 15 #0.3 How much bigger the UD should be calculated than the research are (minimum bounding box). The higher the more Computation
  gridVal <- 700 #700 How much cells or pixels should be considered for UD calculation. The higher the more Computation

  for (j in seq(1, length(methList))) { # compute one home range for each bandwidth method and individual
    if (j == 1) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h="href"))
    }
    else if (j == 2) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hpi"]))
    }
    else if (j == 3) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hbcv"]))
    }
    else if (j == 4) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hscv"])) 
    }
    else {
      print("Error")
    }
    hr99 <- getverticeshr(UD, percent=99,unin = c("m"),unout = c("km2"))
    hr95 <- getverticeshr(UD, percent=95,unin = c("m"),unout = c("km2"))
    hr50 <- getverticeshr(UD, percent=50,unin = c("m"),unout = c("km2"))
    hr99_sf <- st_as_sf(hr99)
    hr95_sf <- st_as_sf(hr95)
    hr50_sf <- st_as_sf(hr50)
    
    ########## plot results ##########

    png(paste(imdir, selected_females[[iterator]],"_",gsub(" ", "", names(monthyear_list)[samplinginterval], fixed = TRUE), "_KDE_VUD_", methList[[j]], ".png", sep=""), width=1800, height=1400, res=240, units="px")
    image(getvolumeUD(UD), xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1)))
    plot(hr50, lty=3, lwd=1.5, border="azure4", add=T,axes=F)
    plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
    plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
    plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
    plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
    plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
    axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
    axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
    #text(xmin-500, (ymin+ymax)/2, "y [m]", adj=c(NA,-4), srt=90)
    title(paste(selected_females[[iterator]], ": KDE ", "(", methListTitle[[j]], ")", sep=""), line=-1.5)
    legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
    dev.off()
    
    results_selectedmethod <- data.frame(bandwithmethod=c(methList[[j]],methList[[j]],methList[[j]]),
            hrpercentage=c("50%","95%","99%"),
            area=c(hr50_sf$area,hr95_sf$area,hr99_sf$area),
            n_gps=c(n_gps,n_gps,n_gps),
            n_follows=c(n_follows,n_follows,n_follows))

    results_selectedfocal<- results_selectedmethod %>% mutate(focal=selected_females[[iterator]])
    
    results_focal <- rbind(results_focal,results_selectedfocal)
  }

  

  print(paste(selected_females[[iterator]], ": Calculation finished", sep=""))
  iterator <- iterator+1
}
results_focal <- results_focal %>% mutate(period=names(monthyear_list)[samplinginterval])
kde_results <- kde_results %>% rbind(results_focal)


}
st_write(kde_results, dsn = paste(
  paste(imdir,
    "kde_temporal_bootstrap_monthly",
    sep = ""
  ),
  "csv",
  sep = "."
))
system("say alle kernel distanzen berechnet für die verschiedenen Untersuchungsperioden")

```

```{r KDE - HR per-researchperiods and all - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/KDE/periods/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/KDE/periods/",format(Sys.Date(), "%Y%m%d"), "/", sep="")

kde_results_perperiod <- data.frame() # results
for (samplinginterval in c(1:length(period_list))){
print(paste("kdes for period: ",samplinginterval))
##### Determine the smoothing parameters
selected_females <- period_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
selected_females <- as_vector(selected_females)

bandwiths <-  data.frame(focal=as_vector(selected_females),
                 hpi=vector(mode="numeric", length=length(selected_females)),hbcv=vector(mode="numeric", length=length(selected_females)),hscv=vector(mode="numeric", length=length(selected_females)),
                 stringsAsFactors=FALSE)
iterator <- 1
for (i in selected_females){
  print(paste("kdes for focal: ",i))
  xy_selected <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  colnames(xy_selected) <- c("x","y") #adjust column names
  hpi <- sqrt(prod(sqrt(diag(Hpi.diag(xy_selected[,1:2], nstage=2))))) # using diagonal bandwith not full bandwith? why?
  hbcv <- sqrt(prod(sqrt(diag(Hbcv.diag(xy_selected[,1:2], whichbcv=1)))))
  hscv <- sqrt(prod(sqrt(diag(Hscv.diag(xy_selected[,1:2], nstage=2)))))
  bandwiths[iterator,"focal"] <- i
  bandwiths[iterator,"hpi"] <- as.numeric(hpi)
  bandwiths[iterator,"hbcv"] <- as.numeric(hbcv)
  bandwiths[iterator,"hscv"] <- as.numeric(hscv)

  iterator <- iterator+1
}
rownames(bandwiths) <- c(1:length(selected_females))

### KDE: Calculate UD, HR
iterator <- 1
results_focal <- data.frame()
for (i in selected_females){ # compute KDE home range for every individual
  n_follows <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>% group_by(follow) %>% summarise() %>% nrow()
  xy_selected <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  n_gps <- xy_selected %>% nrow()
  
  colnames(xy_selected) <- c("x","y") #adjust column names
  resultList <- list()
  methListTitle <- list("REF", "PI", "BCV", "SCV")
  methList <- list("HREF", "HPI", "HBCV", "HSCV")
  
  xy_selected <- SpatialPoints(xy_selected,proj4string = CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" )) # Converts SpatialPointDataFrame to SpatialPoint (only x & y)
  
  extVal <- 15 #0.3 How much bigger the UD should be calculated than the research are (minimum bounding box). The higher the more Computation
  gridVal <- 2000 #700 How much cells or pixels should be considered for UD calculation. The higher the more Computation

  for (j in seq(1, length(methList))) { # compute one home range for each bandwidth method and individual
    if (j == 1) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h="href"))
    }
    else if (j == 2) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hpi"]))
    }
    else if (j == 3) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hbcv"]))
    }
    else if (j == 4) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hscv"])) 
    }
    else {
      print("Error")
    }
    hr99 <- getverticeshr(UD, percent=99,unin = c("m"),unout = c("km2"))
    hr95 <- getverticeshr(UD, percent=95,unin = c("m"),unout = c("km2"))
    hr50 <- getverticeshr(UD, percent=50,unin = c("m"),unout = c("km2"))
    
    hr99_sf <- st_as_sf(hr99)
    hr95_sf <- st_as_sf(hr95)
    hr50_sf <- st_as_sf(hr50)
    
    ########## plot results ##########
    png(paste(imdir, selected_females[[iterator]],"_",gsub(" ", "", names(period_list)[samplinginterval], fixed = TRUE), "_KDE_UD_", methList[[j]], ".png", sep=""), width=1800, height=1400, res=240, units="px")
    par(0,0,0,0)
    image(UD, xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
    plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
    plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
    plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
    plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
    plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
    plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
    axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
    axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
    #text(xmin-550, (ymin+ymax)/2, "y [m]", adj=c(NA,-4), srt=90)
    title(paste(selected_females[[iterator]], ": KDE ", "(", methListTitle[[j]], ")", sep=""), line=-1.5)
    legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
    dev.off()

    png(paste(imdir, selected_females[[iterator]],"_",gsub(" ", "", names(period_list)[samplinginterval], fixed = TRUE), "_KDE_VUD_", methList[[j]], ".png", sep=""), width=1800, height=1400, res=240, units="px")
    image(getvolumeUD(UD), xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1)))
    plot(hr50, lty=3, lwd=1.5, border="azure4", add=T,axes=F)
    plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
    plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
    plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
    plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
    plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
    axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
    axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
    #text(xmin-500, (ymin+ymax)/2, "y [m]", adj=c(NA,-4), srt=90)
    title(paste(selected_females[[iterator]], ": KDE ", "(", methListTitle[[j]], ")", sep=""), line=-1.5)
    legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
    dev.off()
    
    results_selectedmethod <- data.frame(bandwithmethod=c(methList[[j]],methList[[j]],methList[[j]]),
            hrpercentage=c("50%","95%","99%"),
            area=c(hr50_sf$area,hr95_sf$area,hr99_sf$area),
            n_gps=c(n_gps,n_gps,n_gps),
            n_follows=c(n_follows,n_follows,n_follows),
            geometry=c(hr50_sf$geometry,hr95_sf$geometry,hr99_sf$geometry))

    results_selectedfocal<- results_selectedmethod %>% mutate(focal=selected_females[[iterator]])
    
    results_focal <- rbind(results_focal,results_selectedfocal)
  }

  

  print(paste(selected_females[[iterator]], ": Calculation finished", sep=""))
  iterator <- iterator+1
}
results_focal <- results_focal %>% mutate(period=names(period_list)[samplinginterval])
kde_results_perperiod <- kde_results_perperiod %>% rbind(results_focal)


}
st_write(kde_results_perperiod, dsn = paste(
  paste(imdir,
    "kde_values_allperiods",
    sep = ""
  ),
  "shp",
  sep = "."
))
system("say alle kernel distanzen berechnet für die verschiedenen Untersuchungsperioden")


```

```{r KDE - HR per-year - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/KDE/years/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))

kde_results_perperyear <- data.frame() # results
for (samplinginterval in c(1:length(year_list))){
print(paste("kdes for period: ",samplinginterval))
##### Determine the smoothing parameters
selected_females <- year_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
selected_females <- as_vector(selected_females)

bandwiths <-  data.frame(focal=as_vector(selected_females),
                 hpi=vector(mode="numeric", length=length(selected_females)),hbcv=vector(mode="numeric", length=length(selected_females)),hscv=vector(mode="numeric", length=length(selected_females)),
                 stringsAsFactors=FALSE)
iterator <- 1
for (i in selected_females){
  print(paste("kdes for focal: ",i))
  xy_selected <- year_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  colnames(xy_selected) <- c("x","y") #adjust column names
  hpi <- sqrt(prod(sqrt(diag(Hpi.diag(xy_selected[,1:2], nstage=2))))) # using diagonal bandwith not full bandwith? why?
  hbcv <- sqrt(prod(sqrt(diag(Hbcv.diag(xy_selected[,1:2], whichbcv=1)))))
  hscv <- sqrt(prod(sqrt(diag(Hscv.diag(xy_selected[,1:2], nstage=2)))))
  bandwiths[iterator,"focal"] <- i
  bandwiths[iterator,"hpi"] <- as.numeric(hpi)
  bandwiths[iterator,"hbcv"] <- as.numeric(hbcv)
  bandwiths[iterator,"hscv"] <- as.numeric(hscv)

  iterator <- iterator+1
}
rownames(bandwiths) <- c(1:length(selected_females))

### KDE: Calculate UD, HR
iterator <- 1
results_focal <- data.frame()
for (i in selected_females){ # compute KDE home range for every individual
  n_follows <- year_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>% group_by(follow) %>% summarise() %>% nrow()
  xy_selected <- year_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  n_gps <- xy_selected %>% nrow()
  
  colnames(xy_selected) <- c("x","y") #adjust column names
  resultList <- list()
  methListTitle <- list("REF", "PI", "BCV", "SCV")
  methList <- list("HREF", "HPI", "HBCV", "HSCV")
  
  xy_selected <- SpatialPoints(xy_selected,proj4string = CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" )) # Converts SpatialPointDataFrame to SpatialPoint (only x & y)
  
  extVal <- 15 #0.3 How much bigger the UD should be calculated than the research are (minimum bounding box). The higher the more Computation
  gridVal <- 2000 #700 How much cells or pixels should be considered for UD calculation. The higher the more Computation

  for (j in seq(1, length(methList))) { # compute one home range for each bandwidth method and individual
    if (j == 1) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h="href"))
    }
    else if (j == 2) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hpi"]))
    }
    else if (j == 3) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hbcv"]))
    }
    else if (j == 4) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hscv"])) 
    }
    else {
      print("Error")
    }
    hr99 <- getverticeshr(UD, percent=99,unin = c("m"),unout = c("km2"))
    hr95 <- getverticeshr(UD, percent=95,unin = c("m"),unout = c("km2"))
    hr50 <- getverticeshr(UD, percent=50,unin = c("m"),unout = c("km2"))
    
    hr99_sf <- st_as_sf(hr99)
    hr95_sf <- st_as_sf(hr95)
    hr50_sf <- st_as_sf(hr50)
    
    ########## plot results ##########
    imdir <- paste("output/Homeranges/KDE/years/",format(Sys.Date(), "%Y%m%d"), "/", sep="")
    imdirUD <- paste("output/Homeranges/KDE/years/",format(Sys.Date(), "%Y%m%d"),"/", sep="")
    step <- 200
    xmin_xmax_round_steps<- c(step*(ceiling(xmin/step)):(round(xmax/step)))
    xmin_xmax_round_steps <- xmin_xmax_round_steps[2:(length(xmin_xmax_round_steps)-2)] #manual adjustment of tick breaks how many does it need
    ymin_ymax_round_steps<- c(step*(ceiling(ymin/step)):(round(ymax/step)))
    ymin_ymax_round_steps <- ymin_ymax_round_steps[2:(length(ymin_ymax_round_steps)-1)] #manual adjustment of tick breaks how many does it need

    png(paste(imdirUD, selected_females[[iterator]],"_",gsub(" ", "", names(year_list)[samplinginterval], fixed = TRUE), "_KDE_UD_", methList[[j]], ".png", sep=""), width=1800, height=1400, res=240, units="px")
    par(0,0,0,0)
    image(UD, xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
    plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
    plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
    plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
    plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
    plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
    plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
    axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
    axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
    #text(xmin-550, (ymin+ymax)/2, "y [m]", adj=c(NA,-4), srt=90)
    title(paste(selected_females[[iterator]], ": KDE ", "(", methListTitle[[j]], ")", sep=""), line=-1.5)
    legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
    dev.off()

    png(paste(imdir, selected_females[[iterator]],"_",gsub(" ", "", names(year_list)[samplinginterval], fixed = TRUE), "_KDE_VUD_", methList[[j]], ".png", sep=""), width=1800, height=1400, res=240, units="px")
    image(getvolumeUD(UD), xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1)))
    plot(hr50, lty=3, lwd=1.5, border="azure4", add=T,axes=F)
    plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
    plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
    plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
    plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
    plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
    axis(1, at=xmin_xmax_round_steps,pos=ymin+150) # x axis
    axis(2, at=ymin_ymax_round_steps, pos=xmin-500)
    #text(xmin-500, (ymin+ymax)/2, "y [m]", adj=c(NA,-4), srt=90)
    title(paste(selected_females[[iterator]], ": KDE ", "(", methListTitle[[j]], ")", sep=""), line=-1.5)
    legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
    dev.off()
    
    results_selectedmethod <- data.frame(bandwithmethod=c(methList[[j]],methList[[j]],methList[[j]]),
            hrpercentage=c("50%","95%","99%"),
            area=c(hr50_sf$area,hr95_sf$area,hr99_sf$area),
            n_gps=c(n_gps,n_gps,n_gps),
            n_follows=c(n_follows,n_follows,n_follows),
            geometry=c(hr50_sf$geometry,hr95_sf$geometry,hr99_sf$geometry))

    results_selectedfocal<- results_selectedmethod %>% mutate(focal=selected_females[[iterator]])
    
    results_focal <- rbind(results_focal,results_selectedfocal)
  }

  

  print(paste(selected_females[[iterator]], ": Calculation finished", sep=""))
  iterator <- iterator+1
}
results_focal <- results_focal %>% mutate(period=names(year_list)[samplinginterval])
kde_results_perperyear <- kde_results_perperyear %>% rbind(results_focal)


}
st_write(kde_results_perperyear, dsn = paste(
  paste(
    "output/Homeranges/KDE/years/",
    format(Sys.Date(), "%Y%m%d"),"/",
    "kde_values_years",
    sep = ""
  ),
  "shp",
  sep = "."
))
system("say alle kernel distanzen berechnet für jedes Jahr")


```

## LOCOH
```{r TESTING LOCOH - HR per-researchperiods and all - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/LOCOH/periods/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/LOCOH/periods/",format(Sys.Date(), "%Y%m%d"), "/", sep="")

locoh_list <- data.frame() # results
for (samplinginterval in c(1:length(period_list))){
  print(paste("LOCOH for period: ",samplinginterval))
  ##### Determine the smoothing parameters
  selected_females <- period_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
  selected_females <- as_vector(selected_females)
  
  iterator <- 1
  for (i in selected_females) { # do the following for every animal in the list
    print(paste("LOCOH for focal: ",i))
    n_follows <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
        group_by(follow) %>% summarise() %>% nrow()

    xy_selected <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
      dplyr::select(E,N,manualDatetime,infantname,followtype,Activity,SUAQ_followlog.Observer1,SUAQ_orangutans.dominancecat,SUAQ_orangutans.matriline,SUAQ_weather.rainfall_ml_evening,SUAQ_fai.fai,ageFromDOB)
    xy_selected <- xy_selected[order(xy_selected$manualDatetime , decreasing = FALSE ),] 
    n_gps <- xy_selected %>% nrow()

    xy_only <- xy_selected %>%
      dplyr::select(E,N) %>% mutate(x=as.double(E),y=as.double(N)) %>% as.data.frame() %>% dplyr::select(-E,-N) # dataframe was a tibble --> caused troubles in the locoh package

    df.lxy <- xyt.lxy(xy=xy_only, dt=xy_selected$manualDatetime,anv=xy_selected[,4:12], id=i, proj4string=CRS("+init=epsg:23847")) # create xy-object. Duplicates are autom. removed
    # plot(df.lxy)
    hist(df.lxy) #distribution of locations by date, step length and SI -> some gaps, several SI
    # lxy.plot.freq(df.lxy, deltat.by.date=T) #sampling frequency over time -> several SI
    # lxy.plot.freq(df.lxy, cp=T) #find bursts of points that need to be thinned out, 2 sections because of two main SI
    
    # find bursts
    df.lxy <- lxy.thin.bursts(df.lxy, thresh=0.98) ## adjust threshold
    df.lxy <- lxy.ptsh.add(df.lxy)# find appropriate time parameter s. with rule from manual. Depending on RQ --> f.e. researching daily movement we want high s because we want two points near each other not as nearest neighbour because its seperated by a day.
    
    # # find s    
    # lxy.plot.pt2ctr(df.lxy) # proposes a s-value around 0.4
    # lxy.plot.sfinder(df.lxy, delta.t=3600*c(12,24,36,48,96,192)) # proposes a s-value around 0.15 # see how the time gaps are distributed compared to s selection. How much of timegap should be included in nearest neighbour
    # lxy.plot.mtdr(df.lxy, k=10)
    # lxy.plot.tspan(df.lxy, k=10) 
    
    ########################
    svalue <- 0.01      ###### s-VALUE SELECTION
    kselected <- 12    ## kmax-SELECTION
    aval <- 7000
    ########################
    
    df.lxy <- lxy.nn.add(df.lxy, s=svalue, k=kselected)
    df.lhs_12k <- lxy.lhs(df.lxy, k=kselected, s=svalue)
    df.lhs_12k <- lhs.iso.add(df.lhs_12k, iso.levels = c(0.25, 0.5, 0.65, 0.75, 0.85, 0.95)) # creates isopleths for each k-hullset, sorted by density
    

    
    # a method
    df.lxy <- lxy.nn.add(df.lxy, s=svalue, a=auto.a(nnn=12, ptp=0.98))
    df.lxy <- lxy.nn.add(df.lxy, s=svalue, a=12000)
    lhs_amixed <- lxy.lhs(df.lxy, s=svalue, a=4:12*1000, iso.add=T)
    
    for (l in seq(1:length(kmax_vec))) {
      start_time <- Sys.time()
      kmax <- kmax_vec[[l]]
      df.lxy <- lxy.nn.add(df.lxy, s=0.02, k=kmax)
      
      
      # revisits
      df.lhs_12k_revisits <- lhs.visit.add(df.lhs_12k, ivg=3600*18)
      summary(df.lhs_12k_revisits)
      plot(df.lhs_12k_revisits, hpp=T, hpp.classify="nsv", ivg=3600*18, col.ramp="rainbow") # number of visits
      plot(df.lhs_12k_revisits, hpp=T, hpp.classify="mnlv", col.ramp="rainbow") # duration of visit
      hsp <- lhs.plot.scatter(df.lhs_12k_revisits, x="nsv", y="mnlv", col="spiral", bg="black")
      plot(df.lhs_12k_revisits, hpp=T, hsp=hsp, hpp.classify="hsp")
      
      # Directional movement
      df.lhs_12k_directional <- lhs.ellipses.add(df.lhs_12k)
      df.lhs_12k_directional <- lhs.iso.add(df.lhs_12k_directional, sort.metric="ecc")
      plot(df.lhs_12k_directional, iso=T, iso.sort.metric="ecc")
      #lhs.save(df.lhs, dir=paste("~/_MyData/Masterarbeit Movement Analysis/Resultate/thesis/calculations/LCH/", animal, sep="")
      # plot(df.lhs, iso=T, record=T, ufipt=F) #Show results of the different k-/a-/r-values [pg up / pg dn]]
      # lhs.plot.isoarea(df.lhs, legend=0)
      # lhs.plot.isoear(df.lhs, legend=0)
      plot(df.lhs_12k_directional, iso=T, k=12, allpts=T, cex.allpts=0.3, col.allpts="gray30", ufipt=F)
      
      
      
      
      
      
      
      if(TRUE){# only execute when enough data xy_selected_toolessdata
      UD <- BRB(traj, D=vv, Tmax=tmax, Lmin=lmin, hmin=hmin, b=T, same4all=T, grid=gridVal, extent=extVal, filtershort=T, type="UD")
      
      hr95 <- getverticeshr(UD, percent=95, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr50 <- getverticeshr(UD, percent=50, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr99 <- getverticeshr(UD, percent=99, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr95_sf <- st_as_sf(hr95)
      hr50_sf <- st_as_sf(hr50)
      hr99_sf <- st_as_sf(hr99)
      
      par(0,0,0,0)
      step <- 200
      
      ########## plot results ##########
      hminTxt <- gsub(".", "_", as.character(hminFactor[[l]]), fixed=T)
      xmin_xmax_round_steps<- c(step*(ceiling(xmin/step)):(round(xmax/step)))
      xmin_xmax_round_steps <- xmin_xmax_round_steps[2:(length(xmin_xmax_round_steps)-2)] #manual adjustment of tick breaks how many does it need
      ymin_ymax_round_steps<- c(step*(ceiling(ymin/step)):(round(ymax/step)))
      ymin_ymax_round_steps <- ymin_ymax_round_steps[2:(length(ymin_ymax_round_steps)-1)] #manual adjustment of tick breaks how many does it need    
      
  
      
      png(paste(imdir,i, "_BRB_UD_", hminTxt,"_",names(period_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(UD, xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()
      
      png(paste(imdir,i, "_BRB_VUD_", hminTxt,"_",names(period_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(getvolumeUD(UD), xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()}
      
      end_time <- Sys.time()
      duration <- end_time-start_time
      
      ########## export results ############
      if(xy_selected_toolessdata) {
        resultList_50 <-
          data.frame(
            focal = i,
            homerange = hr50$area,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr50_sf$geometry
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = hr95$area,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr95_sf$geometry
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = hr99$area,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr99_sf$geometry
          )
      }
      if (!xy_selected_toolessdata) {
                resultList_50 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
      }
      locoh_list <- locoh_list %>% rbind(resultList_50,resultList_95,resultList_99)
      
      print(paste(hminTxt, ": Calculation finished", sep=""))
    }
  }
}

st_write(locoh_list, dsn = paste(
  paste(
    imdir,
    "locoh_values_allperiods",
    sep = ""
  ),
  "shp",
  sep = "."
))
system("say alle locoh verteilungen berechnet für die verschiedenen Untersuchungsperioden")

lhs.plot.isoarea(cissy.lhs.amixed)
lhs.plot.isoear(cissy.lhs.amixed)
toni.lhs.k15 <- lhs.visit.add(toni.lhs.k15, ivg=3600*12)

```

## BRB
```{r BRB - HR sizes with bootstrap sampling - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/BRB/bootstrap/sampling/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/BRB/bootstrap/sampling/",format(Sys.Date(), "%Y%m%d"), "/", sep="")

#### retrieve sampling intervals
mainSI <- rep(30,length(selected_females))

#### amount of data
amountofdata <- seq(0.5,10,by=0.5)/10
brb_list_bootstrap <- data.frame() # results
follows <- SUAQ_rangepoints_11all20_loc_females_10 %>% group_by(follow) %>% summarise()

for(per in amountofdata){
  selectedfollows<- sample_n(follows, round(nrow(follows)*per))
  xy_selected_perc <- SUAQ_rangepoints_11all20_loc_females_10 %>% filter(follow %in% pull(selectedfollows,follow))
  #### Selected animals as list/vector
  selected_females <- xy_selected_perc %>% group_by(focal) %>% summarise()
  selected_females <- as_vector(selected_females)
  iterator <- 1
  for (i in selected_females) { # do the following for every animal in the list
      print(paste("brb for focal: ",i))
      n_follows <- xy_selected_perc %>% filter(focal%in%c(i)) %>%
          group_by(follow) %>% summarise() %>% nrow()
  
      xy_selected <- xy_selected_perc %>% filter(focal%in%c(i)) %>%
        dplyr::select(E,N,timelag_lead,manualDatetime)
      n_gps <- xy_selected %>% nrow()
  
      colnames(xy_selected) <- c("x","y","timelag_lead","timestamp") #adjust column names
      xy_selected <- xy_selected %>%  mutate(timelag_lead = round(timelag_lead,digits=0))
    
      # only select points where timegaps are between 20-40min --> To do pretty bad selection we would need a function which selects points automatically by subsampling if mean timegap is too high or by interpolating if too small. But how much does it really affect the result of HR --> because we mainly compare HR sizes.
      #xy_selected <- xy_selected %>% filter((timelag_lead/60)>20) #select which SI to remove (xy_selected$timelag_lead/60 < (mainSI[[iterator]]-10))
      #xy_selected <- xy_selected %>% filter((timelag_lead/60)<40)
      #xy_selected <- xy_selected[!(is.na(xy_selected$y)),]
      
      # check if enough gps points are still here otherwise make NA's
      xy_selected_toolessdata <- TRUE
      if(nrow(xy_selected)<30){xy_selected_toolessdata <- FALSE}
      
      # samplingI <- vector()
      # samplingI <- xy_selected$timelag_lead/60
      # samplingI <- samplingI[2:length(samplingI)] # First value of timelag_lead is NA
      # plot(table(samplingI), type="h", xlim=c(20,40),col=4, ylab="frequency", xlab="sampling interval [min]")
      
      xy_selected <- xy_selected[order(xy_selected$timestamp , decreasing = FALSE ),] 
      moveObj <- move(x=xy_selected$x, y=xy_selected$y, time=xy_selected$timestamp, proj=CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" ),animal=i)
      ##########################################################################################
      relUn <- 12 # relocation uncertainty [m]
      ##########################################################################################
      velocity <- unname(quantile(speed(moveObj), 0.99)) * 60 # unit: [m/min]
      hmin <- relUn + (0.5 * velocity * 30) # meters, smoothing
      iterator <- iterator+1
      
      
      
      traj <- as.ltraj(xy=xy_selected[,c("x", "y")], date=xy_selected$timestamp, id=i)
    
      ##########################################################################################
      tmax <- 70*60 # seconds, maximum time span for a segment
      lmin <- 12 # meters, minimal distance of a segment in order to count as movement
      tau <- 1*60 #seconds, approx. duration of one interpolated interval -> defines the number of intervals per step 
      extVal <- 15
      gridVal <- 2000
      hminRef <- hmin # computed through "BRB_leo_hmin computation.R"
      hminFactor <- c(0.6, 0.8, 1,1.2,1.4,  1.6, 1.8)
      ##########################################################################################
    
      vv <- BRB.D(traj, Tmax=tmax, Lmin=lmin) #estimate the diffusion coefficient
      for (l in seq(1:length(hminFactor))) {
        start_time <- Sys.time()
        hmin <- hminFactor[[l]]*hminRef
        if(xy_selected_toolessdata){# only execute when enough data
        UD <- BRB(traj, D=vv, Tmax=tmax, Lmin=lmin, hmin=hmin, b=T, same4all=T, grid=gridVal, extent=extVal, filtershort=T, type="UD")
        
        hr95 <- getverticeshr(UD, percent=95, standardize=TRUE,unin = c("m"),unout = c("km2"))
        hr50 <- getverticeshr(UD, percent=50, standardize=TRUE,unin = c("m"),unout = c("km2"))
        hr99 <- getverticeshr(UD, percent=99, standardize=TRUE,unin = c("m"),unout = c("km2"))
        hr95_sf <- st_as_sf(hr95)
        hr50_sf <- st_as_sf(hr50)
        hr99_sf <- st_as_sf(hr99)
        
        par(0,0,0,0)
        step <- 200
        
        ########## plot results ##########
        hminTxt <- gsub(".", "_", as.character(hminFactor[[l]]), fixed=T)

        
        end_time <- Sys.time()
        duration <- end_time-start_time
        
        ########## export results ############
        if(xy_selected_toolessdata) {
          resultList_50 <-
            data.frame(
              focal = i,
              homerange = hr50$area,
              hrpercentage ="50%",
              n_gps = n_gps,
              n_follows = n_follows,
              Velocity_percentile99 = velocity,
              GPSuncertainty = relUn,
              hmin = hmin,
              tmax = tmax,
              lmin = lmin,
              tau = tau,
              hminfac = hminFactor[[l]],
              hminRef = hminRef,
              processingtime = duration,
              percentagedata=per
            )
          resultList_95 <-
            data.frame(
              focal = i,
              homerange = hr95$area,
              hrpercentage ="95%",
              n_gps = n_gps,
              n_follows = n_follows,
              Velocity_percentile99 = velocity,
              GPSuncertainty = relUn,
              hmin = hmin,
              tmax = tmax,
              lmin = lmin,
              tau = tau,
              hminfac = hminFactor[[l]],
              hminRef = hminRef,
              processingtime = duration,
              percentagedata=per
            )
          resultList_99 <-
            data.frame(
              focal = i,
              homerange = hr99$area,
              hrpercentage ="99%",
              n_gps = n_gps,
              n_follows = n_follows,
              Velocity_percentile99 = velocity,
              GPSuncertainty = relUn,
              hmin = hmin,
              tmax = tmax,
              lmin = lmin,
              tau = tau,
              hminfac = hminFactor[[l]],
              hminRef = hminRef,
              processingtime = duration,
              percentagedata=per
            )
        }
        if (!xy_selected_toolessdata) {
                  resultList_50 <-
            data.frame(
              focal = i,
              homerange = NA,
              hrpercentage ="50%",
              n_gps = n_gps,
              n_follows = n_follows,
              Velocity_percentile99 = velocity,
              GPSuncertainty = relUn,
              hmin = hmin,
              tmax = tmax,
              lmin = lmin,
              tau = tau,
              hminfac = hminFactor[[l]],
              hminRef = hminRef,
              processingtime = duration,
              percentagedata=per
            )
          resultList_95 <-
            data.frame(
              focal = i,
              homerange = NA,
              hrpercentage ="95%",
              n_gps = n_gps,
              n_follows = n_follows,
              Velocity_percentile99 = velocity,
              GPSuncertainty = relUn,
              hmin = hmin,
              tmax = tmax,
              lmin = lmin,
              tau = tau,
              hminfac = hminFactor[[l]],
              hminRef = hminRef,
              processingtime = duration,
              percentagedata=per
            )
          resultList_99 <-
            data.frame(
              focal = i,
              homerange = NA,
              hrpercentage ="99%",
              n_gps = n_gps,
              n_follows = n_follows,
              Velocity_percentile99 = velocity,
              GPSuncertainty = relUn,
              hmin = hmin,
              tmax = tmax,
              lmin = lmin,
              tau = tau,
              hminfac = hminFactor[[l]],
              hminRef = hminRef,
              processingtime = duration,
              percentagedata=per
            )
        }
        brb_list_bootstrap <- brb_list_bootstrap %>% rbind(resultList_50,resultList_95,resultList_99)
        
        print(paste(hminTxt, ": Calculation finished", sep=""))
      }
    }

  }
}
st_write(brb_list_bootstrap, dsn = paste(
  paste(
    imdir,
    "brb_list_bootstrap",
    sep = ""
  ),
  "csv",
  sep = "."
))
system("say alle brown brücken für unterschiedliche samples berechnet")


```

```{r BRB - HR per-researchperiods and all - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/BRB/periods/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/BRB/periods/",format(Sys.Date(), "%Y%m%d"), "/", sep="")

#### retrieve sampling intervals
mainSI <- rep(30,length(selected_females))

brb_list <- data.frame() # results
for (samplinginterval in c(1:length(period_list))){
  print(paste("brb for period: ",samplinginterval))
  ##### Determine the smoothing parameters
  selected_females <- period_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
  selected_females <- as_vector(selected_females)
  
  iterator <- 1
  indexitter <- 0
  for (i in selected_females) { # do the following for every animal in the list
    print(paste("brb for focal: ",i))
    n_follows <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
        group_by(follow) %>% summarise() %>% nrow()

    xy_selected <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
      dplyr::select(E,N,timelag_lead,manualDatetime)
    n_gps <- xy_selected %>% nrow()

    colnames(xy_selected) <- c("x","y","timelag_lead","timestamp") #adjust column names
    xy_selected <- xy_selected %>%  mutate(timelag_lead = round(timelag_lead,digits=0))
  
    # only select points where timegaps are between 20-40min --> To do pretty bad selection we would need a function which selects points automatically by subsampling if mean timegap is too high or by interpolating if too small. But how much does it really affect the result of HR --> because we mainly compare HR sizes.
    #xy_selected <- xy_selected %>% filter((timelag_lead/60)>20) #select which SI to remove (xy_selected$timelag_lead/60 < (mainSI[[iterator]]-10))
    #xy_selected <- xy_selected %>% filter((timelag_lead/60)<40)
    #xy_selected <- xy_selected[!(is.na(xy_selected$y)),]
    
    # check if enough gps points are still here otherwise make NA's
    xy_selected_toolessdata <- TRUE
    if(nrow(xy_selected)<30){xy_selected_toolessdata <- FALSE}
    
    # samplingI <- vector()
    # samplingI <- xy_selected$timelag_lead/60
    # samplingI <- samplingI[2:length(samplingI)] # First value of timelag_lead is NA
    # plot(table(samplingI), type="h", xlim=c(20,40),col=4, ylab="frequency", xlab="sampling interval [min]")
    
    xy_selected <- xy_selected[order(xy_selected$timestamp , decreasing = FALSE ),] 
    moveObj <- move(x=xy_selected$x, y=xy_selected$y, time=xy_selected$timestamp, proj=CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" ),animal=i)
    ##########################################################################################
    relUn <- 12 # relocation uncertainty [m]
    ##########################################################################################
    velocity <- unname(quantile(speed(moveObj), 0.99)) * 60 # unit: [m/min]
    hmin <- relUn + (0.5 * velocity * 30) # meters, smoothing
    iterator <- iterator+1
    
    
    
    traj <- as.ltraj(xy=xy_selected[,c("x", "y")], date=xy_selected$timestamp, id=i)
  
    ##########################################################################################
    tmax <- 70*60 # seconds, maximum time span for a segment
    lmin <- 12 # meters, minimal distance of a segment in order to count as movement
    tau <- 1*60 #seconds, approx. duration of one interpolated interval -> defines the number of intervals per step 
    extVal <- 15
    gridVal <- 2000
    hminRef <- hmin # computed through "BRB_leo_hmin computation.R"
    hminFactor <- c(0.6, 0.8, 1,1.2,1.4,  1.6, 1.8)
    ##########################################################################################
  
    vv <- BRB.D(traj, Tmax=tmax, Lmin=lmin) #estimate the diffusion coefficient
    for (l in seq(1:length(hminFactor))) {
      start_time <- Sys.time()
      hmin <- hminFactor[[l]]*hminRef
      if(xy_selected_toolessdata){# only execute when enough data
      UD <- BRB(traj, D=vv, Tmax=tmax, Lmin=lmin, hmin=hmin, b=T, same4all=T, grid=gridVal, extent=extVal, filtershort=T, type="UD")
      hr95 <- getverticeshr(UD, percent=95, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr50 <- getverticeshr(UD, percent=50, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr99 <- getverticeshr(UD, percent=99, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr95_sf <- st_as_sf(hr95)
      hr50_sf <- st_as_sf(hr50)
      hr99_sf <- st_as_sf(hr99)
      
      par(0,0,0,0)
      step <- 200
      
      ########## plot results ##########
      hminTxt <- gsub(".", "_", as.character(hminFactor[[l]]), fixed=T)
      xmin_xmax_round_steps<- c(step*(ceiling(xmin/step)):(round(xmax/step)))
      xmin_xmax_round_steps <- xmin_xmax_round_steps[2:(length(xmin_xmax_round_steps)-2)] #manual adjustment of tick breaks how many does it need
      ymin_ymax_round_steps<- c(step*(ceiling(ymin/step)):(round(ymax/step)))
      ymin_ymax_round_steps <- ymin_ymax_round_steps[2:(length(ymin_ymax_round_steps)-1)] #manual adjustment of tick breaks how many does it need    
      
  
      
      png(paste(imdir,i, "_BRB_UD_", hminTxt,"_",names(period_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(UD, xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()
      
      png(paste(imdir,i, "_BRB_VUD_", hminTxt,"_",names(period_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(getvolumeUD(UD), xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()}
      
      end_time <- Sys.time()
      duration <- end_time-start_time
      
      ########## export results ############
      if(xy_selected_toolessdata) {
        resultList_50 <-
          data.frame(
            focal = i,
            homerange = hr50$area,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr50_sf$geometry
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = hr95$area,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr95_sf$geometry
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = hr99$area,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr99_sf$geometry
          )
      }
      if (!xy_selected_toolessdata) {
                resultList_50 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
      }
      brb_list <- brb_list %>% rbind(resultList_50,resultList_95,resultList_99)
      
      print(paste(hminTxt, ": Calculation finished", sep=""))
    }
  }
}

st_write(brb_list, dsn = paste(
  paste(
    imdir,
    "brb_values_allperiods",
    sep = ""
  ),
  "shp",
  sep = "."
))
system("say alle kernel distanzen berechnet für die verschiedenen Untersuchungsperioden")


```

```{r BRB - HR per-year - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/BRB/years/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/BRB/years/",format(Sys.Date(), "%Y%m%d"), "/", sep="")

#### retrieve sampling intervals
mainSI <- rep(30,length(selected_females))

brb_list <- data.frame() # results
for (samplinginterval in c(1:length(year_list))){
  print(paste("brb for year: ",samplinginterval))
  ##### Determine the smoothing parameters
  selected_females <- year_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
  selected_females <- as_vector(selected_females)
  
  iterator <- 1
  for (i in selected_females) { # do the following for every animal in the list
    print(paste("brb for focal: ",i))
    n_follows <- year_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
        group_by(follow) %>% summarise() %>% nrow()

    xy_selected <- year_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
      dplyr::select(E,N,timelag_lead,manualDatetime)
    n_gps <- xy_selected %>% nrow()

    colnames(xy_selected) <- c("x","y","timelag_lead","timestamp") #adjust column names
    xy_selected <- xy_selected %>%  mutate(timelag_lead = round(timelag_lead,digits=0))
  
    # only select points where timegaps are between 20-40min --> To do pretty bad selection we would need a function which selects points automatically by subsampling if mean timegap is too high or by interpolating if too small. But how much does it really affect the result of HR --> because we mainly compare HR sizes.
    #xy_selected <- xy_selected %>% filter((timelag_lead/60)>20) #select which SI to remove (xy_selected$timelag_lead/60 < (mainSI[[iterator]]-10))
    #xy_selected <- xy_selected %>% filter((timelag_lead/60)<40)
    #xy_selected <- xy_selected[!(is.na(xy_selected$y)),]
    
    # check if enough gps points are still here otherwise make NA's
    xy_selected_toolessdata <- TRUE
    if(nrow(xy_selected)<30){xy_selected_toolessdata <- FALSE}
    
    # samplingI <- vector()
    # samplingI <- xy_selected$timelag_lead/60
    # samplingI <- samplingI[2:length(samplingI)] # First value of timelag_lead is NA
    # plot(table(samplingI), type="h", xlim=c(20,40),col=4, ylab="frequency", xlab="sampling interval [min]")
    
    xy_selected <- xy_selected[order(xy_selected$timestamp , decreasing = FALSE ),] 
    moveObj <- move(x=xy_selected$x, y=xy_selected$y, time=xy_selected$timestamp, proj=CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" ),animal=i)
    ##########################################################################################
    relUn <- 12 # relocation uncertainty [m]
    ##########################################################################################
    velocity <- unname(quantile(speed(moveObj), 0.99)) * 60 # unit: [m/min]
    hmin <- relUn + (0.5 * velocity * 30) # meters, smoothing
    iterator <- iterator+1
    
    
    
    traj <- as.ltraj(xy=xy_selected[,c("x", "y")], date=xy_selected$timestamp, id=i)
  
    ##########################################################################################
    tmax <- 70*60 # seconds, maximum time span for a segment
    lmin <- 12 # meters, minimal distance of a segment in order to count as movement
    tau <- 1*60 #seconds, approx. duration of one interpolated interval -> defines the number of intervals per step 
    extVal <- 15
    gridVal <- 2000
    hminRef <- hmin # computed through "BRB_leo_hmin computation.R"
    hminFactor <- c(0.6, 0.8, 1,1.2,1.4,  1.6, 1.8)
    ##########################################################################################
  
    vv <- BRB.D(traj, Tmax=tmax, Lmin=lmin) #estimate the diffusion coefficient
    for (l in seq(1:length(hminFactor))) {
      start_time <- Sys.time()
      hmin <- hminFactor[[l]]*hminRef
      if(xy_selected_toolessdata){# only execute when enough data
      UD <- BRB(traj, D=vv, Tmax=tmax, Lmin=lmin, hmin=hmin, b=T, same4all=T, grid=gridVal, extent=extVal, filtershort=T, type="UD")
      
      hr95 <- getverticeshr(UD, percent=95, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr50 <- getverticeshr(UD, percent=50, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr99 <- getverticeshr(UD, percent=99, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr95_sf <- st_as_sf(hr95)
      hr50_sf <- st_as_sf(hr50)
      hr99_sf <- st_as_sf(hr99)
      
      par(0,0,0,0)
      step <- 200
      
      ########## plot results ##########
      hminTxt <- gsub(".", "_", as.character(hminFactor[[l]]), fixed=T)
      xmin_xmax_round_steps<- c(step*(ceiling(xmin/step)):(round(xmax/step)))
      xmin_xmax_round_steps <- xmin_xmax_round_steps[2:(length(xmin_xmax_round_steps)-2)] #manual adjustment of tick breaks how many does it need
      ymin_ymax_round_steps<- c(step*(ceiling(ymin/step)):(round(ymax/step)))
      ymin_ymax_round_steps <- ymin_ymax_round_steps[2:(length(ymin_ymax_round_steps)-1)] #manual adjustment of tick breaks how many does it need    
      
  
      
      png(paste(imdir,i, "_BRB_UD_", hminTxt,"_",names(year_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(UD, xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()
      
      png(paste(imdir,i, "_BRB_VUD_", hminTxt,"_",names(year_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(getvolumeUD(UD), xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()}
      
      end_time <- Sys.time()
      duration <- end_time-start_time
      
      ########## export results ############
      if(xy_selected_toolessdata) {
        resultList_50 <-
          data.frame(
            focal = i,
            homerange = hr50$area,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(year_list)[samplinginterval],
            processingtime = duration,
            geometry=hr50_sf$geometry
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = hr95$area,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(year_list)[samplinginterval],
            processingtime = duration,
            geometry=hr95_sf$geometry
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = hr99$area,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(year_list)[samplinginterval],
            processingtime = duration,
            geometry=hr99_sf$geometry
          )
      }
      if (!xy_selected_toolessdata) {
                resultList_50 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(year_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(year_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(year_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
      }
      brb_list <- brb_list %>% rbind(resultList_50,resultList_95,resultList_99)
      
      print(paste(hminFactor[[l]],": Calculation finished", sep=""))
    }
  }
}

st_write(brb_list, dsn = paste(
  paste(
    imdir,
    "brb_values_years",
    sep = ""
  ),
  "shp",
  sep = "."
))
system("say alle kernel distanzen berechnet für die verschiedenen Untersuchungsjahre")


```

## dBBMM
```{r TESTING dBBMM - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/CTMM/periods/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/CTMM/periods/",format(Sys.Date(), "%Y%m%d"), "/", sep="")

#### retrieve sampling intervals
mainSI <- rep(30,length(selected_females))

ctmm_list <- data.frame() # results
for (samplinginterval in c(1:length(period_list))){
  print(paste("brb for period: ",samplinginterval))
  ##### Determine the smoothing parameters
  selected_females <- period_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
  selected_females <- as_vector(selected_females)
  
  iterator <- 1
  for (i in selected_females) { # do the following for every animal in the list
    print(paste("brb for focal: ",i))
    n_follows <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
        group_by(follow) %>% summarise() %>% nrow()

    xy_selected <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
      dplyr::select(E,N,timelag_lead,manualDatetime)
    n_gps <- xy_selected %>% nrow()

    colnames(xy_selected) <- c("x","y","timelag_lead","timestamp") #adjust column names
    xy_selected <- xy_selected %>%  mutate(timelag_lead = round(timelag_lead,digits=0))
  
    # only select points where timegaps are between 20-40min --> To do pretty bad selection we would need a function which selects points automatically by subsampling if mean timegap is too high or by interpolating if too small. But how much does it really affect the result of HR --> because we mainly compare HR sizes.
    #xy_selected <- xy_selected %>% filter((timelag_lead/60)>20) #select which SI to remove (xy_selected$timelag_lead/60 < (mainSI[[iterator]]-10))
    #xy_selected <- xy_selected %>% filter((timelag_lead/60)<40)
    #xy_selected <- xy_selected[!(is.na(xy_selected$y)),]
    
    # check if enough gps points are still here otherwise make NA's
    xy_selected_toolessdata <- TRUE
    if(nrow(xy_selected)<30){xy_selected_toolessdata <- FALSE}
    
    # samplingI <- vector()
    # samplingI <- xy_selected$timelag_lead/60
    # samplingI <- samplingI[2:length(samplingI)] # First value of timelag_lead is NA
    # plot(table(samplingI), type="h", xlim=c(20,40),col=4, ylab="frequency", xlab="sampling interval [min]")
    
    xy_selected <- xy_selected[order(xy_selected$timestamp , decreasing = FALSE ),] 
    moveObj <- move(x=xy_selected$x, y=xy_selected$y, time=xy_selected$timestamp, proj=CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" ),animal=i)
    ##########################################################################################
    relUn <- 12 # relocation uncertainty [m]
    ##########################################################################################
    velocity <- unname(quantile(speed(moveObj), 0.99)) * 60 # unit: [m/min]
    hmin <- relUn + (0.5 * velocity * 30) # meters, smoothing
    iterator <- iterator+1
    
    
    
    traj <- as.ltraj(xy=xy_selected[,c("x", "y")], date=xy_selected$timestamp, id=i)
  
    ##########################################################################################
    tmax <- 70*60 # seconds, maximum time span for a segment
    lmin <- 12 # meters, minimal distance of a segment in order to count as movement
    tau <- 1*60 #seconds, approx. duration of one interpolated interval -> defines the number of intervals per step 
    extVal <- 15
    gridVal <- 2000
    hminRef <- hmin # computed through "BRB_leo_hmin computation.R"
    hminFactor <- c(0.6, 0.8, 1,1.2,1.4,  1.6, 1.8)
    ##########################################################################################
  
    vv <- BRB.D(traj, Tmax=tmax, Lmin=lmin) #estimate the diffusion coefficient
    for (l in seq(1:length(hminFactor))) {
      start_time <- Sys.time()
      hmin <- hminFactor[[l]]*hminRef
      if(xy_selected_toolessdata){# only execute when enough data
      UD <- BRB(traj, D=vv, Tmax=tmax, Lmin=lmin, hmin=hmin, b=T, same4all=T, grid=gridVal, extent=extVal, filtershort=T, type="UD")
      
      hr95 <- getverticeshr(UD, percent=95, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr50 <- getverticeshr(UD, percent=50, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr99 <- getverticeshr(UD, percent=99, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr95_sf <- st_as_sf(hr95)
      hr50_sf <- st_as_sf(hr50)
      hr99_sf <- st_as_sf(hr99)
      
      par(0,0,0,0)
      step <- 200
      
      ########## plot results ##########
      hminTxt <- gsub(".", "_", as.character(hminFactor[[l]]), fixed=T)
      xmin_xmax_round_steps<- c(step*(ceiling(xmin/step)):(round(xmax/step)))
      xmin_xmax_round_steps <- xmin_xmax_round_steps[2:(length(xmin_xmax_round_steps)-2)] #manual adjustment of tick breaks how many does it need
      ymin_ymax_round_steps<- c(step*(ceiling(ymin/step)):(round(ymax/step)))
      ymin_ymax_round_steps <- ymin_ymax_round_steps[2:(length(ymin_ymax_round_steps)-1)] #manual adjustment of tick breaks how many does it need    
      
  
      
      png(paste(imdir,i, "_BRB_UD_", hminTxt,"_",names(period_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(UD, xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()
      
      png(paste(imdir,i, "_BRB_VUD_", hminTxt,"_",names(period_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(getvolumeUD(UD), xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()}
      
      end_time <- Sys.time()
      duration <- end_time-start_time
      
      ########## export results ############
      if(xy_selected_toolessdata) {
        resultList_50 <-
          data.frame(
            focal = i,
            homerange = hr50$area,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr50_sf$geometry
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = hr95$area,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr95_sf$geometry
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = hr99$area,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr99_sf$geometry
          )
      }
      if (!xy_selected_toolessdata) {
                resultList_50 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
      }
      ctmm_list <- ctmm_list %>% rbind(resultList_50,resultList_95,resultList_99)
      
      print(paste(hminTxt, ": Calculation finished", sep=""))
    }
  }
}

st_write(ctmm_list, dsn = paste(
  paste(
    imdir,
    "ctmm_values_allperiods",
    sep = ""
  ),
  "shp",
  sep = "."
))
system("say alle kernel distanzen berechnet für die verschiedenen Untersuchungsperioden")


```

```{r CTMM - HR size online,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}

##### Save Dataset for CTMM online f.e.
SUAQ_rangepoints_11all20_loc_females12_wgs <- SUAQ_rangepoints_11all20_loc_wgs[order(SUAQ_rangepoints_11all20_loc_wgs$focal,SUAQ_rangepoints_11all20_loc_wgs$manualDatetime , decreasing = TRUE ),] %>% rename(
    c("event-id"="identification","timestamp"="manualDatetime","location-long"="long","location-lat"="lat","Animal ID"="focal","Tag No"="follow")
  ) %>% filter(`Animal ID`%in% c("lisa","friska","ellie","cissy","lilly","yulia","raffi","sarabi","trident","alice","tiara","cinnamon")) %>% mutate("sensor-type"="gps")
write_csv(SUAQ_rangepoints_11all20_loc_wgs, path = paste(
  paste(
    "/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/",
    "SUAQ_rangepoints_11all20_loc_females12_wgs",
    sep = ""
  ),
  "csv",
  sep = "."
))
SUAQ_rangepoints_11all20_loc_females6_wgs <- SUAQ_rangepoints_11all20_loc_wgs[order(SUAQ_rangepoints_11all20_loc_wgs$focal,SUAQ_rangepoints_11all20_loc_wgs$manualDatetime , decreasing = TRUE ),] %>% rename(
    c("event-id"="identification","timestamp"="manualDatetime","location-long"="long","location-lat"="lat","Animal ID"="focal","Tag No"="follow")
  ) %>% filter(`Animal ID`%in% c("lisa","friska","ellie","cissy","lilly","yulia")) %>% mutate("sensor-type"="gps")
write_csv(SUAQ_rangepoints_11all20_loc_females6_wgs, path = paste(
  paste(
    "/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/",
    "SUAQ_rangepoints_11all20_loc_females6_wgs",
    sep = ""
  ),
  "csv",
  sep = "."
))


SUAQ_rangepoints_11all20_loc_females_10_wgs <- SUAQ_rangepoints_11all20_loc_wgs[order(SUAQ_rangepoints_11all20_loc_wgs$focal,SUAQ_rangepoints_11all20_loc_wgs$manualDatetime , decreasing = TRUE ),] %>% rename(
    c("event-id"="identification","timestamp"="manualDatetime","location-long"="long","location-lat"="lat","Animal ID"="focal","Tag No"="follow")
  ) %>% filter(`Animal ID`%in% c("lisa",
      "friska",
      "ellie",
      "cissy",
      "lilly",
      "yulia",
      "raffi",
      "sarabi",
      "trident",
      "tiara")) %>% mutate("sensor-type"="gps")
write_csv(SUAQ_rangepoints_11all20_loc_females_10_wgs, path = paste(
  paste(
    "/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/",
    "SUAQ_rangepoints_11all20_loc_females_10_wgs",
    sep = ""
  ),
  "csv",
  sep = "."
))

```


# Trees per Homeranges and full dataset
```{r Homeranges - Create full dataset ,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
## Import datasets
### read in ctmm results
#### pooled
shplist_pooled<- list.files("/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/output/Homeranges/CTMM/periods/pooled/20210325/10_most_ctmm_4errors_12uere_pooled_HR_shp/", pattern="*.shp", full.names=TRUE)

start_time <- Sys.time()
SUAQ_ctmm_homeranges_pooled<- shplist_pooled%>% 
  map_df(~st_read_homeranges(.))
end_time <- Sys.time()
print(start_time-end_time)
SUAQ_ctmm_homeranges_pooled <- SUAQ_ctmm_homeranges_pooled %>% mutate(modelbase="pooled",algorithm="ctmm")
SUAQ_ctmm_homeranges_pooled$area <- as.numeric(st_area(SUAQ_ctmm_homeranges_pooled))/1000000

#### unpooled
shplist_unpooled<- list.files("/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/output/Homeranges/CTMM/periods/unpooled/20210325/hrshape/", pattern="*.shp", full.names=TRUE)

start_time <- Sys.time()
SUAQ_ctmm_homeranges_unpooled<- shplist_unpooled%>% 
  map_df(~st_read_homeranges(.))
end_time <- Sys.time()
print(start_time-end_time)
SUAQ_ctmm_homeranges_unpooled <- SUAQ_ctmm_homeranges_unpooled %>% mutate(modelbase="unpooled",algorithm="ctmm")
SUAQ_ctmm_homeranges_unpooled$area <- as.numeric(st_area(SUAQ_ctmm_homeranges_unpooled))/1000000

#### total
ctmm_allperiods_sf <- SUAQ_ctmm_homeranges_unpooled %>% rbind(SUAQ_ctmm_homeranges_pooled) %>% dplyr::select(-statisticalBoundary) %>% mutate(period="total research period")

#### Homerange sizes overview
kde_allperiods_sf<- st_read(dsn="output/Homeranges/KDE/periods/20210315/kde_values_allperiods.shp") %>% 
  mutate(algorithm="kde") %>% st_set_crs(23847)
brb_allperiods_sf<- st_read(dsn="output/Homeranges/BRB/periods/20210315/brb_values_allperiods.shp") %>% 
  mutate(algorithm="brb") %>% rename(area=homerng) %>% st_set_crs(23847)
mcp_allperiods_sf<- st_read(dsn="output/Homeranges/MCP/periods/20210316/mcp_values_allperiods.shp") %>% 
  mutate(algorithm="mcp") %>% rename(area=homerng) %>% st_set_crs(23847)
all_homeranges_sf <- kde_allperiods_sf %>% bind_rows(brb_allperiods_sf) %>% bind_rows(mcp_allperiods_sf) %>% bind_rows(ctmm_allperiods_sf)
all_homeranges_sf$period <- factor(all_homeranges_sf$period, levels = c("before 2012", "2012 until 2016", "left", "2016 until August 2018", "Since August 2018","total research period"))
all_homeranges <- all_homeranges_sf %>% st_drop_geometry()

### Feeding trees
SUAQ_waypoints_11all20_loc_tree <- SUAQ_waypoints_11all20_loc %>% filter(PtType=="tree")
SUAQ_waypoints_11all20_loc_tree_sf <-  st_as_sf(SUAQ_waypoints_11all20_loc_tree, 
                          coords = c("E","N"), 
                          crs = 23847) # can also do that in the new script
SUAQ_waypoints_11all20_loc_sf <- SUAQ_waypoints_11all20_loc
SUAQ_waypoints_11all20_loc_sf <-  st_as_sf(SUAQ_waypoints_11all20_loc_sf, 
                          coords = c("E","N"), 
                          crs = 23847) # can also do that in the new script




```

```{r Homeranges - Number of trees in HR ,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
##### Initialize progressbar
pb <-  txtProgressBar(min = 0, max = NROW(all_homeranges_sf), initial = 0) 
stepi <-0
all_homeranges_sf <- all_homeranges_sf %>% mutate(number_of_trees=NA_integer_)
##### Adding number of fruittrees
for (i in c(1:NROW(all_homeranges_sf))){
  number_of_pnts_within<- st_intersection(SUAQ_waypoints_11all20_loc_tree_sf$geometry,all_homeranges_sf[i,]) %>% NROW()
  all_homeranges_sf$number_of_trees[i] <- number_of_pnts_within
  setTxtProgressBar(pb,stepi) # change progress bar
  stepi <-  stepi + 1
}
pb <-  txtProgressBar(min = 0, max = NROW(all_homeranges_sf), initial = 0) 
stepi <-0
all_homeranges_sf <- all_homeranges_sf %>% mutate(researcheffort_ngps=NA_integer_)
for (i in c(1:NROW(all_homeranges_sf))){
  number_of_pnts_within<- st_intersection(SUAQ_waypoints_11all20_loc_sf$geometry,all_homeranges_sf[i,]) %>% NROW()
  all_homeranges_sf$researcheffort_ngps[i] <- number_of_pnts_within
  setTxtProgressBar(pb,stepi) # change progress bar
  stepi <-  stepi + 1
}
all_homeranges_sf <- all_homeranges_sf %>% mutate(fruittree_normalized=number_of_trees/researcheffort_ngps) %>% mutate(fruittree_norm_per_sqkm=fruittree_normalized/area) %>% mutate(fruittree_per_sqkm=number_of_trees/area)

st_write(all_homeranges_sf, dsn = paste(
  paste(
    "/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/",
    "all_homeranges_sf",
    sep = ""
  ),
  "shp",  
  sep = "."
))
write.csv(colnames(all_homeranges_sf),"/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/all_homeranges_sf_colnames.csv")

### Create homeranges dataframe with HR levels in each row
all_homeranges_dense50 <- all_homeranges %>% filter(hrprcnt=="50%") 
all_homeranges_dense95 <- all_homeranges %>% filter(hrprcnt=="95%")  %>% dplyr::select(area,hrprcnt,focal,bndwthm,period,hmin)
all_homeranges_dense99 <- all_homeranges %>% filter(hrprcnt=="99%")  %>% dplyr::select(area,hrprcnt,focal,bndwthm,period,hmin)
all_homeranges_dense100 <- all_homeranges %>% filter(hrprcnt=="100%")  %>% dplyr::select(area,hrprcnt,focal,bndwthm,period,hmin)
all_homeranges_dense <- all_homeranges_dense50 %>% 
  left_join(all_homeranges_dense95,by=c("focal"="focal","bndwthm"="bndwthm","period"="period","hmin"="hmin")) %>% 
  left_join(all_homeranges_dense99,by=c("focal"="focal","bndwthm"="bndwthm","period"="period","hmin"="hmin")) %>% 
  left_join(all_homeranges_dense100,by=c("focal"="focal","bndwthm"="bndwthm","period"="period","hmin"="hmin")) %>%   rename(area_50=area.x,hrprcnt_50=hrprcnt.x,area_95=area.y,hrprcnt_95=hrprcnt.y,area_99=area.x.x,hrprcnt_99=hrprcnt.x.x,area_100=area.y.y,hrprcnt_100=hrprcnt.y.y)
all_homeranges_dense <- all_homeranges_dense %>% mutate(corearea_ratio=area_99/area_50) 


all_homeranges_dense %>% filter(hminfac%in%c(1,NA)) %>% filter(bndwthm%in%c("HREF",NA)) %>%  ggplot(aes(corearea_ratio,focal)) +geom_boxplot()+geom_point(aes(color=algorithm))


```

# Overlaps and Number of Trees in Overlap
## Overlaps calculated from 95 and 50% HR
```{r Homeranges overlaps 50 and 95 (Simple, Jaccard, more??) - Create full dataset ,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# reduce the HR amount select the one which we wanna analyse
all_homeranges_sf <- st_read(dsn="/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/all_homeranges_sf.shp")
colnames(all_homeranges_sf)<- c("bndwthm","hrprcnt","area","n_gps","n_fllws","focal","period","algorithm","Vlct_99","GPSncrt","hmin","tmax","lmin","tau","hminfac","hminRef","model","modltype","modelbase","number_of_trees","researcheffort_ngps","fruittree_normalized","fruittree_norm_per_sqkm","fruittree_per_sqkm","geometry")
all_homeranges_sf_selected <- all_homeranges_sf %>% filter(bndwthm %in% c(NA,"HBCV","HSCV")) %>% filter(hrprcnt%in%c("95%","50%")) %>% filter(hminfac %in% c(NA,0.6,1,1.6)) %>%  dplyr::select(-Vlct_99,-lmin,-fruittree_normalized,-researcheffort_ngps,-number_of_trees,-fruittree_per_sqkm,-fruittree_norm_per_sqkm,-GPSncrt,-tau,-hmin)

vecalgo<- levels(factor(as.factor(all_homeranges_sf_selected$algorithm)))
periods<- levels(factor(as.factor(all_homeranges_sf_selected$period)))

all_overlaps_jaccard <- data.frame()
starttest <- TRUE
for(algo in vecalgo){
  print(algo)
  vechomreange <- levels(factor(as.factor(all_homeranges_sf_selected$hrprcnt)))
  for (hr in vechomreange){
    print(hr)
    if(algo=="ctmm"){methods <- levels(factor(as.factor(all_homeranges_sf_selected$modelbase)))}
    if(algo=="kde"){methods <- levels(factor(as.factor(all_homeranges_sf_selected$bndwthm)))}
    if(algo=="brb"){methods <- levels(factor(as.factor(all_homeranges_sf_selected$hminfac)))}
    if(algo=="mcp"){methods <- "mcptotal"}
    for (peri in periods){
      print(peri)
      for(method in methods){
      print(method)
      if(algo=="kde"){algo_data <- all_homeranges_sf_selected %>% filter(algorithm==algo) %>% filter(hrprcnt==hr) %>% filter(bndwthm==method) %>% filter(period==peri)}
      if(algo=="ctmm"){algo_data <- all_homeranges_sf_selected %>% filter(algorithm==algo) %>% filter(hrprcnt==hr) %>% filter(modelbase==method) %>% filter(period==peri)}
      if(algo=="brb"){algo_data <- all_homeranges_sf_selected %>% filter(algorithm==algo) %>% filter(hrprcnt==hr) %>% filter(hminfac==method)%>% filter(period==peri)}
      if(algo=="mcp"){algo_data <- all_homeranges_sf_selected %>% filter(algorithm==algo) %>% filter(hrprcnt==hr)%>% filter(period==peri)}
      focals<- levels(as.factor(algo_data$focal))
      focalcombination<- expand.grid(focals,focals)
      indx <- !duplicated(t(apply(focalcombination, 1, sort))) # finds non - duplicates in sorted rows
      focalcombination<- focalcombination[indx, ]
      focalcombination<- focalcombination %>% filter(Var1!=Var2)
      for(overlap in c(1:NROW(focalcombination))){
        #print(paste0("overlap calc of ", focalcombination[overlap,1]," : ",focalcombination[overlap,2]))
        sf_focal1 <- algo_data %>% filter(focal==focalcombination[overlap,1])
        sf_focal2 <- algo_data %>% filter(focal==focalcombination[overlap,2])
        result_overlap<- st_intersection(sf_focal1,sf_focal2)
        jaccard_overlap <- ifelse(as.logical(st_intersects(sf_focal2,sf_focal1, sparse = FALSE)),as.numeric(((st_area(result_overlap)/1000000)/(st_area(st_union(sf_focal1,sf_focal2))/1000000))),0)
        result_overlap<- bind_cols(result_overlap,data.frame(overlap=as.numeric(st_area(result_overlap)/1000000)))
        result_overlap<- bind_cols(result_overlap,data.frame(jaccard_overlap=jaccard_overlap)) %>% ungroup()
        ifelse(starttest,
               all_overlaps_jaccard <- result_overlap ,
               all_overlaps_jaccard <- all_overlaps_jaccard %>% rbind(result_overlap))
        starttest <- FALSE

      }
      }
    }
  }
}

### Adding number of trees in overlaps --> use the above initialized datasets
##### Initialize progressbar
pb <-  txtProgressBar(min = 0, max = NROW(all_overlaps_jaccard), initial = 0) 
stepi <-0
all_overlaps_jaccard <- all_overlaps_jaccard %>% mutate(number_of_trees=NA_integer_)


##### Adding number of fruittrees
for (i in c(1:NROW(all_overlaps_jaccard))){
  number_of_pnts_within<- st_intersection(SUAQ_waypoints_11all20_loc_tree_sf$geometry,all_overlaps_jaccard[i,]) %>% NROW()
  all_overlaps_jaccard$number_of_trees[i] <- number_of_pnts_within
  setTxtProgressBar(pb,stepi) # change progress bar
  stepi <-  stepi + 1
}
pb <-  txtProgressBar(min = 0, max = NROW(all_overlaps_jaccard), initial = 0) 
stepi <-0
all_overlaps_jaccard <- all_overlaps_jaccard %>% mutate(researcheffort_ngps=NA_integer_)
for (i in c(1:NROW(all_overlaps_jaccard))){
  number_of_pnts_within<- st_intersection(SUAQ_waypoints_11all20_loc_sf$geometry,all_overlaps_jaccard[i,]) %>% NROW()
  all_overlaps_jaccard$researcheffort_ngps[i] <- number_of_pnts_within
  setTxtProgressBar(pb,stepi) # change progress bar
  stepi <-  stepi + 1
}
all_overlaps_jaccard <- all_overlaps_jaccard %>% mutate(fruittree_normalized=number_of_trees/researcheffort_ngps) %>% mutate(fruittree_norm_per_sqkm=fruittree_normalized/area) %>% mutate(fruittree_per_sqkm=number_of_trees/area)


st_write(all_overlaps_jaccard, dsn = paste(
  paste(
    "/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/",
    "all_overlaps_jaccard",
    sep = ""
  ),
  "shp",
  sep = "."
))
write.csv(colnames(all_overlaps_jaccard),"/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/all_overlaps_jaccard.csv")

### Adding additional infos
explanatoryvars <- SUAQ_rangepoints_11all20_loc_females_10[,c("focal","SUAQ_orangutans.matriline", "SUAQ_orangutans.agecategory","SUAQ_orangutans.dominancecat")] %>% group_by(focal,SUAQ_orangutans.matriline,SUAQ_orangutans.agecategory,SUAQ_orangutans.dominancecat) %>% summarise()
colnames(explanatoryvars) <- c("focal","matriline","agecat","dominance")

all_overlaps_50 <- all_overlaps_jaccard %>% left_join(explanatoryvars, by=c("focal"="focal")) %>% filter(hrprcnt=="50%")
all_overlaps_95 <- all_overlaps_jaccard %>% left_join(explanatoryvars, by=c("focal"="focal")) %>% filter(hrprcnt=="95%")
colnames(explanatoryvars) <- c("focal1","matriline1","agecat1","dominance1")
all_overlaps_50 <- all_overlaps_50 %>% left_join(explanatoryvars, by=c("focal.1"="focal1")) %>% filter(hrprcnt=="50%")
all_overlaps_95 <- all_overlaps_95 %>% left_join(explanatoryvars, by=c("focal.1"="focal1")) %>% filter(hrprcnt=="95%")
all_overlaps_50 <- all_overlaps_50 %>% mutate(dominancediff=abs(dominance1-dominance),related=ifelse(matriline1==matriline,TRUE,FALSE),agediff=abs(agecat1-agecat))
all_overlaps_95 <- all_overlaps_95 %>% mutate(dominancediff=abs(dominance1-dominance),related=ifelse(matriline1==matriline,TRUE,FALSE),agediff=abs(agecat1-agecat))


all_overlaps_95 %>% 
  #filter(period=="total research period") %>%  
  ggplot(aes(jaccard_overlap,agediff,group=agediff)) + 
  geom_boxplot(aes(jaccard_overlap ,agediff))+
  geom_point(aes(color=as.character(paste0(focal," : ",focal.1))),size=7)+
  facet_wrap(~ algorithm)

all_overlaps_95 %>% 
  #filter(period=="total research period") %>%  
  ggplot(aes(jaccard_overlap,dominancediff,group=dominancediff)) + 
  geom_boxplot(aes(jaccard_overlap ,dominancediff))+
  geom_point(aes(color=as.character(paste0(focal," : ",focal.1))),size=7)+
  facet_wrap(~ algorithm)

all_overlaps_95 %>% 
  #filter(period=="total research period") %>%  
  ggplot(aes(jaccard_overlap,related)) + 
  geom_boxplot(aes(jaccard_overlap ,related))+
  geom_point(aes(color=as.character(paste0(focal," : ",focal.1))),size=7)+
  facet_wrap(~ algorithm)

ggplot() + 
  geom_point(data=all_overlaps_jaccard,aes(focal,fruittree_normalized,color=hrprcnt))+
  geom_jitter(data=all_homeranges_sf_selected,aes(focal,fruittree_normalized))
```

```{r Homeranges simple overlaps - load created full datasets ,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
### Load datasets if already available and adding right colnames because they are automatically cut probably by st_write function
all_overlaps<- st_read(dsn="/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/all_overlaps_jaccard.shp")
colnames<- read.csv("/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/all_overlaps_jaccard.csv")
colnames<- colnames$x
colnames(all_overlaps) <- colnames
all_homeranges_sf <- st_read(dsn="/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/all_homeranges_sf.shp")
colnames<- read.csv("/Users/stefgr/Nextcloud/UZH/12_Semester/20200313_Masterthesis/R/data/Processed_data/all_homeranges_sf_colnames.csv")
colnames<- colnames$x
colnames(all_homeranges_sf) <- colnames



```



## Overlaps calculated from UD of algorithms
```{r Homeranges overlaps recalculate UD from BRB - HR per-researchperiods and all - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# Filter research periods if only one of the periods should be calculated
# period_list <- period_list[1] # keep only the total research period

# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/BRB/periods/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/BRB/periods/",format(Sys.Date(), "%Y%m%d"), "/", sep="")

#### retrieve sampling intervals
mainSI <- rep(30,length(selected_females))

brb_list <- data.frame() # results
brb_overlap_UD <- list() # only for tital research period activate filter above
for (samplinginterval in c(1:length(period_list))){
  print(paste("brb for period: ",samplinginterval))
  ##### Determine the smoothing parameters
  selected_females <- period_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
  selected_females <- as_vector(selected_females)
  
  iterator <- 1
  indexitter <- 0
  for (i in selected_females) { # do the following for every animal in the list
    print(paste("brb for focal: ",i))
    n_follows <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
        group_by(follow) %>% summarise() %>% nrow()

    xy_selected <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>%
      dplyr::select(E,N,timelag_lead,manualDatetime)
    n_gps <- xy_selected %>% nrow()

    colnames(xy_selected) <- c("x","y","timelag_lead","timestamp") #adjust column names
    xy_selected <- xy_selected %>%  mutate(timelag_lead = round(timelag_lead,digits=0))
  
    # only select points where timegaps are between 20-40min --> To do pretty bad selection we would need a function which selects points automatically by subsampling if mean timegap is too high or by interpolating if too small. But how much does it really affect the result of HR --> because we mainly compare HR sizes.
    #xy_selected <- xy_selected %>% filter((timelag_lead/60)>20) #select which SI to remove (xy_selected$timelag_lead/60 < (mainSI[[iterator]]-10))
    #xy_selected <- xy_selected %>% filter((timelag_lead/60)<40)
    #xy_selected <- xy_selected[!(is.na(xy_selected$y)),]
    
    # check if enough gps points are still here otherwise make NA's
    xy_selected_toolessdata <- TRUE
    if(nrow(xy_selected)<30){xy_selected_toolessdata <- FALSE}
    
    # samplingI <- vector()
    # samplingI <- xy_selected$timelag_lead/60
    # samplingI <- samplingI[2:length(samplingI)] # First value of timelag_lead is NA
    # plot(table(samplingI), type="h", xlim=c(20,40),col=4, ylab="frequency", xlab="sampling interval [min]")
    
    xy_selected <- xy_selected[order(xy_selected$timestamp , decreasing = FALSE ),] 
    moveObj <- move(x=xy_selected$x, y=xy_selected$y, time=xy_selected$timestamp, proj=CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" ),animal=i)
    ##########################################################################################
    relUn <- 12 # relocation uncertainty [m]
    ##########################################################################################
    velocity <- unname(quantile(speed(moveObj), 0.99)) * 60 # unit: [m/min]
    hmin <- relUn + (0.5 * velocity * 30) # meters, smoothing
    iterator <- iterator+1
    
    
    
    traj <- as.ltraj(xy=xy_selected[,c("x", "y")], date=xy_selected$timestamp, id=i)
  
    ##########################################################################################
    tmax <- 70*60 # seconds, maximum time span for a segment
    lmin <- 12 # meters, minimal distance of a segment in order to count as movement
    tau <- 1*60 #seconds, approx. duration of one interpolated interval -> defines the number of intervals per step 
    extVal <- 15
    gridVal <- 2000
    hminRef <- hmin # computed through "BRB_leo_hmin computation.R"
    hminFactor <- c(0.6, 0.8, 1,1.2,1.4,  1.6, 1.8)
    ##########################################################################################
  
    vv <- BRB.D(traj, Tmax=tmax, Lmin=lmin) #estimate the diffusion coefficient
    for (l in seq(1:length(hminFactor))) {
      start_time <- Sys.time()
      hmin <- hminFactor[[l]]*hminRef
      if(xy_selected_toolessdata){# only execute when enough data
      UD <- BRB(traj, D=vv, Tmax=tmax, Lmin=lmin, hmin=hmin, b=T, same4all=T, grid=gridVal, extent=extVal, filtershort=T, type="UD")
      hr95 <- getverticeshr(UD, percent=95, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr50 <- getverticeshr(UD, percent=50, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr99 <- getverticeshr(UD, percent=99, standardize=TRUE,unin = c("m"),unout = c("km2"))
      hr95_sf <- st_as_sf(hr95)
      hr50_sf <- st_as_sf(hr50)
      hr99_sf <- st_as_sf(hr99)
      
      par(0,0,0,0)
      step <- 200
      
      ########## save UD ##########
      brb_overlap_UD[[1+indexitter]] <- list(i,UD,"brb",hminFactor[[l]],n_gps,n_follows)
      indexitter <- indexitter+1
      
      ########## plot results ##########
      hminTxt <- gsub(".", "_", as.character(hminFactor[[l]]), fixed=T)
      xmin_xmax_round_steps<- c(step*(ceiling(xmin/step)):(round(xmax/step)))
      xmin_xmax_round_steps <- xmin_xmax_round_steps[2:(length(xmin_xmax_round_steps)-2)] #manual adjustment of tick breaks how many does it need
      ymin_ymax_round_steps<- c(step*(ceiling(ymin/step)):(round(ymax/step)))
      ymin_ymax_round_steps <- ymin_ymax_round_steps[2:(length(ymin_ymax_round_steps)-1)] #manual adjustment of tick breaks how many does it need    
      
  
      
      png(paste(imdir,i, "_BRB_UD_", hminTxt,"_",names(period_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(UD, xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()
      
      png(paste(imdir,i, "_BRB_VUD_", hminTxt,"_",names(period_list)[samplinginterval], ".png", sep=""), width=1800, height=1400, res=240, units="px")
      image(getvolumeUD(UD), xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
      plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
      plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
      plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
      plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
      plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
      plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
      plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
      axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
      axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
      title(paste(i, ": BRB (", hminFactor[[l]], "*hmin)", sep=""), line=-1.5)
      legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
      dev.off()}
      
      end_time <- Sys.time()
      duration <- end_time-start_time
      
      ########## export results ############
      if(xy_selected_toolessdata) {
        resultList_50 <-
          data.frame(
            focal = i,
            homerange = hr50$area,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr50_sf$geometry
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = hr95$area,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr95_sf$geometry
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = hr99$area,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=hr99_sf$geometry
          )
      }
      if (!xy_selected_toolessdata) {
                resultList_50 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="50%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_95 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="95%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
        resultList_99 <-
          data.frame(
            focal = i,
            homerange = NA,
            hrpercentage ="99%",
            n_gps = n_gps,
            n_follows = n_follows,
            Velocity_percentile99 = velocity,
            GPSuncertainty = relUn,
            hmin = hmin,
            tmax = tmax,
            lmin = lmin,
            tau = tau,
            hminfac = hminFactor[[l]],
            hminRef = hminRef,
            period = names(period_list)[samplinginterval],
            processingtime = duration,
            geometry=NA
          )
      }
      brb_list <- brb_list %>% rbind(resultList_50,resultList_95,resultList_99)
      
      print(paste(hminTxt, ": Calculation finished", sep=""))
    }
  }
}

st_write(brb_list, dsn = paste(
  paste(
    imdir,
    "brb_values_allperiods",
    sep = ""
  ),
  "shp",
  sep = "."
))
system("say alle kernel distanzen berechnet für die verschiedenen Untersuchungsperioden")


```

```{r Homeranges overlaps recalculate UD from KDE - HR per-researchperiods and all - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# Filter research periods if only one of the periods should be calculated
period_list <- period_list[1] # keep only the total research period

# createa directory for kde periods calculation (named by the current date)
dir.create(paste(
    "output/Homeranges/KDE/periods/",
    format(Sys.Date(), "%Y%m%d"),
    sep = ""
  ))
imdir <- paste("output/Homeranges/KDE/periods/",format(Sys.Date(), "%Y%m%d"), "/", sep="")

kde_results_perperiod <- data.frame() # results
kde_overlap_UD <- list() # only for tital research period activate filter above

for (samplinginterval in c(1:length(period_list))){
print(paste("kdes for period: ",samplinginterval))
##### Determine the smoothing parameters
selected_females <- period_list[[samplinginterval]] %>% group_by(focal) %>% summarise()
selected_females <- as_vector(selected_females)

bandwiths <-  data.frame(focal=as_vector(selected_females),
                 hpi=vector(mode="numeric", length=length(selected_females)),hbcv=vector(mode="numeric", length=length(selected_females)),hscv=vector(mode="numeric", length=length(selected_females)),
                 stringsAsFactors=FALSE)
iterator <- 1
indexitter <- 0
for (i in selected_females){
  print(paste("kdes for focal: ",i))
  xy_selected <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  colnames(xy_selected) <- c("x","y") #adjust column names
  hpi <- sqrt(prod(sqrt(diag(Hpi.diag(xy_selected[,1:2], nstage=2))))) # using diagonal bandwith not full bandwith? why?
  hbcv <- sqrt(prod(sqrt(diag(Hbcv.diag(xy_selected[,1:2], whichbcv=1)))))
  hscv <- sqrt(prod(sqrt(diag(Hscv.diag(xy_selected[,1:2], nstage=2)))))
  bandwiths[iterator,"focal"] <- i
  bandwiths[iterator,"hpi"] <- as.numeric(hpi)
  bandwiths[iterator,"hbcv"] <- as.numeric(hbcv)
  bandwiths[iterator,"hscv"] <- as.numeric(hscv)

  iterator <- iterator+1
}
rownames(bandwiths) <- c(1:length(selected_females))

### KDE: Calculate UD, HR
iterator <- 1
results_focal <- data.frame()
for (i in selected_females){ # compute KDE home range for every individual
  n_follows <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>% group_by(follow) %>% summarise() %>% nrow()
  xy_selected <- period_list[[samplinginterval]] %>% filter(focal%in%c(i)) %>% dplyr::select(E,N)
  n_gps <- xy_selected %>% nrow()
  
  colnames(xy_selected) <- c("x","y") #adjust column names
  resultList <- list()
  methListTitle <- list("REF", "PI", "BCV", "SCV")
  methList <- list("HREF", "HPI", "HBCV", "HSCV")
  
  xy_selected <- SpatialPoints(xy_selected,proj4string = CRS( "+proj=utm +zone=47 +a=6378160 +b=6356774.50408554 +towgs84=-24,-15,5,0,0,0,0 +units=m +no_defs" )) # Converts SpatialPointDataFrame to SpatialPoint (only x & y)
  
  extVal <- 15 #0.3 How much bigger the UD should be calculated than the research are (minimum bounding box). The higher the more Computation
  gridVal <- 2000 #700 How much cells or pixels should be considered for UD calculation. The higher the more Computation

  for (j in seq(1, length(methList))) { # compute one home range for each bandwidth method and individual
    if (j == 1) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h="href"))
    }
    else if (j == 2) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hpi"]))
    }
    else if (j == 3) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hbcv"]))
    }
    else if (j == 4) {
      UD <- (kernelUD(xy_selected, grid=gridVal, extent=extVal, h=bandwiths[iterator,"hscv"])) 
    }
    else {
      print("Error")
    }
    hr99 <- getverticeshr(UD, percent=99,unin = c("m"),unout = c("km2"))
    hr95 <- getverticeshr(UD, percent=95,unin = c("m"),unout = c("km2"))
    hr50 <- getverticeshr(UD, percent=50,unin = c("m"),unout = c("km2"))
    
    hr99_sf <- st_as_sf(hr99)
    hr95_sf <- st_as_sf(hr95)
    hr50_sf <- st_as_sf(hr50)
    
    ########## save UD ##########
      kde_overlap_UD[[1+indexitter]] <- list(i,UD,"kde",methList[[j]],n_gps,n_follows)
      indexitter <- indexitter+1
    
    ########## plot results ##########
    png(paste(imdir, selected_females[[iterator]],"_",gsub(" ", "", names(period_list)[samplinginterval], fixed = TRUE), "_KDE_UD_", methList[[j]], ".png", sep=""), width=1800, height=1400, res=240, units="px")
    par(0,0,0,0)
    image(UD, xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1))) 
    plot(hr50, lty=3, lwd=1.5, border="darkseagreen3", add=T,axes=F)
    plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
    plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
    plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
    plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
    plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
    axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
    axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
    #text(xmin-550, (ymin+ymax)/2, "y [m]", adj=c(NA,-4), srt=90)
    title(paste(selected_females[[iterator]], ": KDE ", "(", methListTitle[[j]], ")", sep=""), line=-1.5)
    legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
    dev.off()

    png(paste(imdir, selected_females[[iterator]],"_",gsub(" ", "", names(period_list)[samplinginterval], fixed = TRUE), "_KDE_VUD_", methList[[j]], ".png", sep=""), width=1800, height=1400, res=240, units="px")
    image(getvolumeUD(UD), xlab="x [m]",ylab="y [m]",xlim=c(xmin_xmax_round_steps[1],tail(xmin_xmax_round_steps,n=1)),ylim=c(ymin_ymax_round_steps[1],tail(ymin_ymax_round_steps,n=1)))
    plot(hr50, lty=3, lwd=1.5, border="azure4", add=T,axes=F)
    plot(hr95, lty=1, lwd=1, border="cadetblue2", add=T, axes=F)
    plot(hr99, lty=4, lwd=1, border="darkturquoise", add=T, axes=F)
    plot(SUAQ_contour_sf_loc,lty=3, lwd=0.2, col=alpha("black", 1), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 21,cex=1.5, col=alpha("black", 0.7), add=T, axes=F)
    plot(SUAQ_station_sf_loc,pch = 16,cex=1, col=alpha("darkred", 0.7),lwd=0.3, bg="darkred", add=T, axes=F)
    plot(SUAQ_river_sf_loc,lty=1, lwd=0.3, col=alpha("cornflowerblue", 0.7), add=T, axes=F)
    plot(SUAQ_pathnetwork_sf_loc, lty=1, lwd=0.3, col=alpha("darkolivegreen", 0.7), add=T, axes=F)
    axis(1, at=xmin_xmax_round_steps,pos=ymin+axisshifty) # x axis
    axis(2, at=ymin_ymax_round_steps, pos=xmin+axisshiftx)
    #text(xmin-500, (ymin+ymax)/2, "y [m]", adj=c(NA,-4), srt=90)
    title(paste(selected_females[[iterator]], ": KDE ", "(", methListTitle[[j]], ")", sep=""), line=-1.5)
    legend("topright", c("HR 50%", "HR 95%", "HR 100%","Trail","River"), col=c("coral1", "cadetblue2", "darkturquoise","darkred",alpha("cornflowerblue", 0.7)), lwd=c(2.5,2,2,0.3,4), lty=c(3,1,4,1,1), inset=c(legendoffsety,legendoffsety))
    dev.off()
    
    results_selectedmethod <- data.frame(bandwithmethod=c(methList[[j]],methList[[j]],methList[[j]]),
            hrpercentage=c("50%","95%","99%"),
            area=c(hr50_sf$area,hr95_sf$area,hr99_sf$area),
            n_gps=c(n_gps,n_gps,n_gps),
            n_follows=c(n_follows,n_follows,n_follows),
            geometry=c(hr50_sf$geometry,hr95_sf$geometry,hr99_sf$geometry))

    results_selectedfocal<- results_selectedmethod %>% mutate(focal=selected_females[[iterator]])
    
    results_focal <- rbind(results_focal,results_selectedfocal)
  }

  

  print(paste(selected_females[[iterator]], ": Calculation finished", sep=""))
  iterator <- iterator+1
}
results_focal <- results_focal %>% mutate(period=names(period_list)[samplinginterval])
kde_results_perperiod <- kde_results_perperiod %>% rbind(results_focal)


}
st_write(kde_results_perperiod, dsn = paste(
  paste(imdir,
    "kde_values_allperiods",
    sep = ""
  ),
  "shp",
  sep = "."
))
system("say alle kernel distanzen berechnet für die verschiedenen Untersuchungsperioden")


```

# Visualizations
```{r Homeranges - VISUALS ANALYSIS fruittrees ,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
### Finding: In core areas the number of trees increase linear with area in 95% area it increases faster but then flattens
all_homeranges_sf  %>%  ggplot(aes(area,number_of_trees,color=algorithm)) +geom_point()
all_homeranges_sf %>% ggplot(aes(area,fruittree_normalized,color=hrprcnt)) + geom_point()

### Fruittreeindex
all_homeranges_sf %>%filter(algorithm=="brb") %>% 
  filter(hminfac==1) %>%  
  ggplot(aes(focal,fruittree_normalized))+
  geom_point(aes(color=hrprcnt),size=3)+
  geom_smooth()+
  labs(x="focal",y=expression("normalized trees [tree/tot gps effort]"),color="HR [%]") +
  scale_fill_viridis()

```

```{r Overlaps - VISUALS ANALYSIS fruittrees ,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
### Finding: In core areas the number of trees increase linear with area in 95% area it increases faster but then flattens
all_overlaps_jaccard  %>%  ggplot(aes(jaccard_overlap,number_of_trees,color=algorithm)) +geom_point()
all_homeranges_sf %>% ggplot(aes(area,fruittree_normalized,color=hrprcnt)) + geom_point()

### Fruittreeindex
all_overlaps_jaccard %>%
  #filter(algorithm=="brb") %>% 
  #filter(hminfac==1) %>%  
  ggplot(aes(overlap,fruittree_normalized))+
  geom_point(aes(color=hrprcnt),size=3)+
  geom_point(data=all_homeranges_sf,aes(area,fruittree_normalized),alpha=0.2)+
  labs(x="overlap or area",y=expression("normalized trees [tree/tot gps effort]"),color="HR [%]") +
  scale_fill_viridis()+
  scale_x_log10()

```

```{r Homerange - VISUALS - STABILITY BOOTSTRAPPING sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# Homerange stability: BOOTSTRAPPING METHODS
### KDE temporal
kde_temporal_bootstrap_monthly<- read_csv("output/Homeranges/KDE/temporalbootstrap/20210315/kde_temporal_bootstrap_monthly.csv")

### make less breaks x axis
period_vec<-kde_temporal_bootstrap_monthly %>% group_by(period) %>%summarise() %>%  pull(period) 
breaks_vec<-period_vec[seq(1, length(period_vec), 6)]

kde_temporal_bootstrap_monthly_ptsize <- kde_temporal_bootstrap_monthly %>% group_by(focal,n_gps) %>% summarise() %>% arrange(desc(n_gps)) %>% ungroup()
kde_temporal_bootstrap_monthly_ptsize<- kde_temporal_bootstrap_monthly_ptsize%>%  group_by(focal) %>%
  top_n(n = 1)%>% rename(n_gps_tot=n_gps)
kde_temporal_bootstrap_monthly <- kde_temporal_bootstrap_monthly  %>% 
  left_join(kde_temporal_bootstrap_monthly_ptsize, by=c("focal"="focal")) %>% 
  mutate(relativeamount=n_gps/n_gps_tot)

plot_kde_temporal_bootstrap_monthly<- kde_temporal_bootstrap_monthly %>% 
  ggplot(aes(period,area,colour=hrpercentage)) + geom_point(aes(size=relativeamount))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 7))+
  facet_grid(focal ~  bandwithmethod)+
  scale_x_discrete(breaks=breaks_vec)+
  scale_colour_manual(values = c("50%" = "#1b9e77", "95%" = "#d95f02", "99%" = "#7570b3"))+
  scale_size_continuous(range = c(0.1, 1.5))

plot_kde_temporal_bootstrap_monthly %>% ggsave(filename = paste(
  paste(
    "/Users/stefgr/Dropbox/Apps/Overleaf/Msc_Thesis/figures/",
    format(Sys.Date(), "%Y%m%d"),"_",
    "plot_kde_temporal_bootstrap_monthly_withamountofdata",
    sep = ""
  ),
  "png",
  sep = "."
),dpi=600,width = 21,height = 29.7,units="cm")

### BRB temporal
brb_temporal_bootstrap_monthly<- read_csv("output/Homeranges/BRB/bootstrap/sampling/20210315/brb_list_bootstrap.csv")

### make less breaks x axis
period_vec<-brb_temporal_bootstrap_monthly %>% group_by(percentagedata) %>%summarise() %>%  pull(percentagedata) 
breaks_vec<-period_vec[seq(1, length(period_vec), 6)]

brb_temporal_bootstrap_monthly_ptsize <- brb_temporal_bootstrap_monthly %>% group_by(focal,n_gps) %>% summarise() %>% arrange(desc(n_gps)) %>% ungroup()
brb_temporal_bootstrap_monthly_ptsize<- brb_temporal_bootstrap_monthly_ptsize%>%  group_by(focal) %>%
  top_n(n = 1)%>% rename(n_gps_tot=n_gps)
brb_temporal_bootstrap_monthly <- brb_temporal_bootstrap_monthly  %>% 
  left_join(brb_temporal_bootstrap_monthly_ptsize, by=c("focal"="focal")) %>% 
  mutate(relativeamount=n_gps/n_gps_tot)

# quick explanation the case that sometimes the percentage of data subsampled is higher but the amount of data is lesser than another with a lower subsampling percentage, is due to the fact that I subsampled percentage of follows from the WHOLE dataset. Not only from one individual. Therefore sometimes a lot of follows of one individual are included an sometimes less. Whats better? depends because subsampling the whole dataset would more adjusting for sampling differences or so.
plot_brb_temporal_bootstrap_monthly<- brb_temporal_bootstrap_monthly %>% 
  filter(hrpercentage=="95%") %>%  
  arrange(desc(relativeamount)) %>% 
  ggplot(aes(relativeamount,homerange,colour=focal)) + geom_jitter(aes(size=n_gps_tot,shape=as.factor(hminfac)))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 10))+
  scale_fill_viridis()+
  #scale_x_discrete(breaks=breaks_vec)+
  #scale_colour_manual(values = c("50%" = "#1b9e77", "95%" = "#d95f02", "99%" = "#7570b3"))+
  scale_size_continuous(range = c(0.6, 7))+
  labs(y=expression("area [km"^2*"]"),title = "BRB homeranges per subset of dataset", x= "percentage of data",color="Orangutan",size="Total GPS per focal",shape="Variation of hmin parametrization")
  #geom_path()
plot_brb_temporal_bootstrap_monthly
plot_brb_temporal_bootstrap_monthly %>% ggsave(filename = paste(
  paste(
    "/Users/stefgr/Dropbox/Apps/Overleaf/Msc_Thesis/figures/",
    format(Sys.Date(), "%Y%m%d"),"_",
    "plot_brb_sampling_bootstrap_monthly_95all_hminparametrization",
    sep = ""
  ),
  "png",
  sep = "."
),dpi=600,width = 29.7,height = 21,units="cm")


```

```{r Homerange - VISUALS - HR sizes,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
#### Homerange sizes overview
plot_all50_area_perfemale<- all_homeranges_sf %>% filter(hrprcnt=="50%") %>% 
  ggplot( aes(x=focal, y=area, fill=algorithm)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.4, alpha=0.3) +
    theme_ipsum() +
    theme(
      #legend.position="none",
      plot.title = element_text(size=18),
      axis.title.x = element_blank()
    ) +
    ggtitle(expression("Core home range areas: KDE"^"50%"*"")) +
    ylab(expression("area [km"^2*"]"))
plot_all50_area_perfemale


plot_all50_area_perfemale %>% ggsave(filename = paste(
  paste(
    "/Users/stefgr/Dropbox/Apps/Overleaf/Msc_Thesis/figures/",
    format(Sys.Date(), "%Y%m%d"),"_",
    "plot_50range_BRB_KDE_MCP_area_perfemale",
    sep = ""
  ),
  "png",
  sep = "."
),dpi=150,width = 29.7,height = 21,units="cm")



#### Analyze ??: distribution of weather?
plot_anypoints <- SUAQ_waypoints_11all20_loc[order(SUAQ_waypoints_11all20_loc$manualDatetime , decreasing = FALSE ),] %>%
  filter(PtType=="nightnest") %>% 
  filter(SUAQ_weather.temperature_min_evening>0) %>% 
  group_by(follow) %>%
  ggplot()+
  geom_point(aes(E,N,color=log10(SUAQ_weather.temperature_min_evening)),size=2.5)+
  #geom_path(aes(E,N), alpha = 0.7,size=0.5)+
  geom_sf(data = SUAQ_pathnetwork_sf_loc,
                                               aes(colour = "black", alpha = 1),
                                               fill = NA,colour = "black")
  
ggplotly(plot_anypoints)

#### Analyze edges: Missing Homeranges?
SUAQ_waypoints_11all20_loc_lost <- SUAQ_waypoints_11all20_loc %>% filter(PtType%in%c("lost"))
SUAQ_waypoints_11all20_loc_lost

plot_lostpoints<-SUAQ_waypoints_11all20_loc_lost[order(SUAQ_waypoints_11all20_loc_lost$manualDatetime , decreasing = FALSE ),] %>%
  group_by(follow) %>%
  ggplot()+
  geom_point(data=SUAQ_waypoints_11all20_loc[order(SUAQ_waypoints_11all20_loc$manualDatetime , decreasing = FALSE ),], aes(E,N),size=0.2,fill="azure4",color="azure4",alpha=0.3)+
  geom_point(aes(E,N,color=focal),size=0.8)+
  geom_text(aes(E,N+40,label=focal,color=focal),size=3.5)+
  #geom_path(aes(E,N), alpha = 0.7,size=0.5)+
  geom_sf(data = SUAQ_pathnetwork_sf_loc,
                                               aes(colour = "black", alpha = 1),
                                               fill = NA,colour = "black")
  
ggplotly(plot_lostpoints)

#### Analyze weather: distribution of weather?
plot_weatherpoints<-SUAQ_waypoints_11all20_loc[order(SUAQ_waypoints_11all20_loc$manualDatetime , decreasing = FALSE ),] %>%
  filter(SUAQ_weather.rainfall_ml_evening!=0) %>% 
  #filter(PtType=="mornnest") %>% 
  group_by(follow) %>%
  ggplot()+
  geom_point(aes(E,N,color=log10(SUAQ_weather.rainfall_ml_evening)),size=1)+
  #geom_path(aes(E,N), alpha = 0.7,size=0.5)+
  geom_sf(data = SUAQ_pathnetwork_sf_loc,
                                               aes(colour = "black", alpha = 1),
                                               fill = NA,colour = "black")
  
ggplotly(plot_weatherpoints)

#### BRB: Find most recent folder
mostrecent <- list.dirs(path = "output/Homeranges/KDE/periods/", full.names = FALSE, recursive = FALSE)
mostrecent <- as.data.frame(mostrecent)
colnames(mostrecent) <- c("date")
mostrecent <- mostrecent %>% filter(date!="_archive") %>%  arrange(desc(date)) %>% top_n(1,date)
imdir <-paste("output/Homeranges/KDE/periods/",mostrecent$date[1], "/", sep="")
kde_values_allperiods <- read_csv(paste0(imdir,"kde_values_allperiods.csv"))

#### BRB: Find most recent folder
mostrecent <- list.dirs(path = "output/Homeranges/BRB/periods/", full.names = FALSE, recursive = FALSE)
mostrecent <- as.data.frame(mostrecent)
colnames(mostrecent) <- c("date")
mostrecent <- mostrecent %>% filter(date!="_archive") %>%  arrange(desc(date)) %>% top_n(1,date)
imdir <-paste("output/Homeranges/BRB/periods/",mostrecent$date[1], "/", sep="")

#### Load homerange results
brb_list <- read_csv(paste0(imdir,"brb_values_allperiods.csv"))

test<- SUAQ_rangepoints_11all20_loc %>% filter(focal%in%c("friska")) %>% filter(year(manualDate)<2012)
year(SUAQ_waypoints_11all20_loc$manualDate[1])

brb_selected_periods<- brb_list %>% 
  #filter(focal %in%c("ellie","lisa","friska","cissy"))%>% 
  mutate(period=fct_relevel(period, 
            "before 2012" ,"2012 until 2016","2016 until August 2018","Since August 2018","total research period"))
  #filter(hminfac==1)

brb_list_selected_err<-  brb_selected_periods %>% 
  mutate(hminfac=as.numeric(hminfac)) %>%  
  filter(hminfac %in% c(0.8,1.2))  %>% 
  spread(hminfac,hr95) %>%
  group_by(focal,nGPS,period) %>%
  summarise_each(funs(first(.[!is.na(.)]))) %>% 
  rename(hminfac_0_8=`0.8`,hminfac_1_2=`1.2`)



png(file=paste(
  paste(
    "/Users/stefgr/Dropbox/Apps/Overleaf/Msc_Thesis/figures/",
    format(Sys.Date(), "%Y%m%d"),"_",
    "brb_selected_periods",
    sep = ""
  ),
  "png",
  sep = "."
),width=900, height=700, res=120, units="px") # or other device
brb_selected_periods %>% filter(hminfac==1) %>% ggplot() + 
  geom_errorbar(data=brb_list_selected_err,aes(x=period,ymin=hminfac_0_8, ymax=hminfac_1_2,color=focal),width=0.2)+
  geom_point(aes(period,hr95,group=interaction(focal),color=focal),size=3)+
  geom_line(aes(period,hr95,group=interaction(focal),color=focal),size=0.5)+
  scale_color_brewer(palette="Set1") +
  theme_minimal()+
  geom_text(aes(x=period,y=hr95,label=nGPS),check_overlap = TRUE,hjust=1.5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(axis.title.x=element_blank())+
  labs(y=expression("Homerange 95% [km"^2*"]"))
dev.off()

```

# Statistics
```{r Homerange - Statistics, Explanatory variables,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
# KDE Homeranges periods and all
explanatoryvars <- SUAQ_rangepoints_11all20_loc_females_10[,c("focal","SUAQ_orangutans.matriline", "SUAQ_orangutans.agecategory","SUAQ_orangutans.dominancecat")] %>% group_by(focal,SUAQ_orangutans.matriline,SUAQ_orangutans.agecategory,SUAQ_orangutans.dominancecat) %>% summarise()
colnames(explanatoryvars) <- c("focal","matriline","agecat","dominance")
all_homeranges_50 <- all_homeranges_sf %>% st_drop_geometry() %>%  left_join(explanatoryvars, by=c("focal"="focal")) %>% filter(hrprcnt=="50%")
all_homeranges_95 <- all_homeranges_sf %>% st_drop_geometry() %>% left_join(explanatoryvars, by=c("focal"="focal")) %>% filter(hrprcnt=="95%")

### Matriline
all_homeranges_50_naomit <- all_homeranges_50 %>% dplyr::select(area,matriline,period,focal,n_gps,n_fllws,algorithm) %>%  na.omit()
all_homeranges_95_naomit <- all_homeranges_95 %>% dplyr::select(area,matriline,period,focal,n_gps,n_fllws,algorithm) %>% na.omit()
hr_all_nullmodell_95 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_95_naomit,REML=FALSE)
hr_all_nullmodell_50 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_50_naomit,REML=FALSE)

all_homeranges_50_naomit %>% ggplot(aes(as.factor(matriline),area,colour=period))+
  geom_boxplot() +
  geom_jitter(alpha=0.5)
# which sample indicator is more correlated with area
all_homeranges_50 %>% ggplot(aes(n_gps,area,colour=period))+
  geom_boxplot() +
  geom_jitter(alpha=0.5)
test_samplesize <- lm(area ~ n_gps+n_fllws,all_homeranges_50_naomit,REML=FALSE)
anova(test_samplesize) # n-gps has higher explanatory power

#### core area
lmer_numOfPoints50 <- lmer(area ~ matriline+(1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_50_naomit,REML=FALSE)
anova(lmer_numOfPoints50,hr_all_nullmodell_50)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints50)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs
#### core area
lmer_numOfPoints95 <- lmer(area ~ matriline+(1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_95_naomit,REML=FALSE)
anova(lmer_numOfPoints95,hr_all_nullmodell_95)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints95)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs



### Age
all_homeranges_50_naomit <- all_homeranges_50 %>% dplyr::select(area,agecat,period,focal,n_gps,n_fllws,algorithm) %>%  na.omit() 
all_homeranges_95_naomit <- all_homeranges_95 %>% dplyr::select(area,agecat,period,focal,n_gps,n_fllws,algorithm) %>% na.omit()
hr_all_nullmodell_95 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_95_naomit,REML=FALSE)
hr_all_nullmodell_50 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_50_naomit,REML=FALSE)

all_homeranges_50_naomit %>% ggplot(aes(as.factor(agecat),area,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()
all_homeranges_95_naomit %>% ggplot(aes(as.factor(agecat),area,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()

#### core area
lmer_numOfPoints50 <- lmer(area ~ agecat+(1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_50_naomit,REML=FALSE)
anova(lmer_numOfPoints50,hr_all_nullmodell_50)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints50)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs
#### core area
lmer_numOfPoints95 <- lmer(area ~ agecat+(1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_95_naomit,REML=FALSE)
anova(lmer_numOfPoints95,hr_all_nullmodell_95)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints95)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs


### Dominance
all_homeranges_50_naomit <- all_homeranges_50 %>% dplyr::select(area,dominance,period,focal,n_gps,n_fllws,algorithm) %>%  na.omit() 
all_homeranges_95_naomit <- all_homeranges_95 %>% dplyr::select(area,dominance,period,focal,n_gps,n_fllws,algorithm) %>% na.omit()
hr_all_nullmodell_95 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_95_naomit,REML=FALSE)
hr_all_nullmodell_50 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_50_naomit,REML=FALSE)

all_homeranges_50_naomit %>% ggplot(aes(as.factor(dominance),area,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()
all_homeranges_95_naomit %>% ggplot(aes(as.factor(dominance),area,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()

#### core area
lmer_numOfPoints50 <- lmer(area ~ dominance+(1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_50_naomit,REML=FALSE)
anova(lmer_numOfPoints50,hr_all_nullmodell_50)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints50)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs
#### core area
lmer_numOfPoints95 <- lmer(area ~ dominance+(1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_95_naomit,REML=FALSE)
anova(lmer_numOfPoints95,hr_all_nullmodell_95)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints95)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs

# TREES
### Dominance
all_tree_naomit_50 <- all_homeranges_50 %>% dplyr::select(fruittree_normalized,area,dominance,period,focal,n_gps,n_fllws,algorithm) %>%  na.omit() 
all_tree_naomit_95 <- all_homeranges_95 %>% dplyr::select(fruittree_normalized,area,dominance,period,focal,n_gps,n_fllws,algorithm) %>% na.omit()
hr_all_nullmodell_95 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_tree_naomit_95,REML=FALSE)
hr_all_nullmodell_50 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_tree_naomit_50,REML=FALSE)

all_tree_naomit_50 %>% ggplot(aes(as.factor(dominance),fruittree_normalized,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()
all_tree_naomit_95 %>% ggplot(aes(as.factor(dominance),fruittree_normalized,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()

#### core area
lmer_numOfPoints50 <- lmer(area ~ dominance+(1 | focal)+(1| n_gps)+(1| algorithm),all_tree_naomit_50,REML=FALSE)
anova(lmer_numOfPoints50,hr_all_nullmodell_50)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints50)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs
#### core area
lmer_numOfPoints95 <- lmer(area ~ dominance+(1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_95_naomit,REML=FALSE)
anova(lmer_numOfPoints95,hr_all_nullmodell_95)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints95)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs


### Dominance
all_homeranges_50_naomit <- all_homeranges_50 %>% dplyr::select(area,SUAQ_orangutans.dominancecat,period,focal,n_gps,n_fllws,algorithm) %>%  na.omit() 
all_homeranges_95_naomit <- all_homeranges_95 %>% dplyr::select(area,SUAQ_orangutans.dominancecat,period,focal,n_gps,n_fllws,algorithm) %>% na.omit()
hr_all_nullmodell_95 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_95_naomit,REML=FALSE)
hr_all_nullmodell_50 <- lmer(area ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_50_naomit,REML=FALSE)

all_homeranges_50_naomit %>% ggplot(aes(as.factor(SUAQ_orangutans.dominancecat),area,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()
all_homeranges_95_naomit %>% ggplot(aes(as.factor(SUAQ_orangutans.dominancecat),area,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()

#### core area
lmer_numOfPoints50 <- lmer(area ~ SUAQ_orangutans.dominancecat+(1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_50_naomit,REML=FALSE)
anova(lmer_numOfPoints50,hr_all_nullmodell_50)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints50)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs
#### core area
lmer_numOfPoints95 <- lmer(area ~ SUAQ_orangutans.dominancecat+(1 | focal)+(1| n_gps)+(1| algorithm),all_homeranges_95_naomit,REML=FALSE)
anova(lmer_numOfPoints95,hr_all_nullmodell_95)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints95)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs


```

```{r Homerange overlaps - Statistics, Explanatory variables,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
### Explain jaccard overlap
all_overlaps_50_naomit <- all_overlaps_50 %>% dplyr::select(area,n_gps,overlap,jaccard_overlap,related,period,focal,focal.1,algorithm) %>% na.omit() %>% st_drop_geometry() %>% ungroup()
all_overlaps_95_naomit <- all_overlaps_95 %>% dplyr::select(area,n_gps,overlap,jaccard_overlap,related,period,focal,focal.1,algorithm) %>% na.omit()%>% st_drop_geometry() %>% ungroup()
lmer_overlap_all_nullmodell_95 <- lmer(jaccard_overlap ~ (1 | focal)+(1| n_gps)+(1|algorithm),all_overlaps_95_naomit,REML=FALSE)
lmer_overlap_all_nullmodell_50 <- lmer(jaccard_overlap ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_50_naomit,REML=FALSE)

#### core area, jaccard overlap and matriline
lmer_jaccard_50 <- lmer(jaccard_overlap ~ related+(1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_50_naomit,REML=FALSE)
anova(lmer_jaccard_50,lmer_overlap_all_nullmodell_50)
anova(lmer_jaccard_50)

##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_jaccard_50)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs
#### core area
lmer_numOfPoints95 <- lmer(jaccard_overlap ~ related+(1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_95_naomit,REML=FALSE)
anova(lmer_numOfPoints95,lmer_overlap_all_nullmodell_95)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints95)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs



### Age
all_overlaps_50_naomitage <- all_overlaps_50 %>% dplyr::select(area,n_gps,overlap,jaccard_overlap,agediff,period,focal,focal.1,algorithm) %>%  na.omit() 
all_overlaps_95_naomitage <- all_overlaps_95 %>% dplyr::select(area,n_gps,overlap,jaccard_overlap,agediff,period,focal,focal.1,algorithm) %>% na.omit()
lmer_overlap_nullmodell_95 <- lmer(jaccard_overlap ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_95_naomitage,REML=FALSE)
lmer_overlap_nullmodell_50 <- lmer(jaccard_overlap ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_50_naomitage,REML=FALSE)

all_overlaps_50_naomitage %>% ggplot(aes(as.factor(agediff),jaccard_overlap,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()
all_overlaps_95_naomitage %>% ggplot(aes(as.factor(agediff),jaccard_overlap,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()

#### core area
lmer_jaccard_50_agediff <- lmer(jaccard_overlap ~ agediff+(1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_50_naomitage,REML=FALSE)
anova(lmer_jaccard_50_agediff,lmer_overlap_nullmodell_50)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints50)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs
#### core area
lmer_jaccard_95_agediff <- lmer(jaccard_overlap ~ agediff+(1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_95_naomitage,REML=FALSE)
anova(lmer_jaccard_95_agediff,lmer_overlap_nullmodell_95)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints95)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs


### Dominance
all_overlaps_50_naomit_dominance <- all_overlaps_50 %>% dplyr::select(area,n_gps,overlap,jaccard_overlap,dominancediff,period,focal,focal.1,algorithm) %>%  na.omit() 
all_overlaps_95_naomit_dominance <- all_overlaps_95 %>% dplyr::select(area,n_gps,overlap,jaccard_overlap,dominancediff,period,focal,focal.1,algorithm) %>% na.omit()
lmer_overlap_nullmodell_95 <- lmer(jaccard_overlap ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_95_naomit_dominance,REML=FALSE)
lmer_overlap_nullmodell_50 <- lmer(jaccard_overlap ~ (1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_50_naomit_dominance,REML=FALSE)

all_overlaps_50_naomit_dominance %>% ggplot(aes(as.factor(dominancediff),jaccard_overlap,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()
all_overlaps_95_naomit_dominance %>% ggplot(aes(as.factor(dominancediff),jaccard_overlap,colour=n_gps))+
  geom_boxplot() +
  geom_jitter()

#### core area
lmer_jaccard_50_dominance <- lmer(jaccard_overlap ~ dominancediff+(1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_50_naomit_dominance,REML=FALSE)
anova(lmer_jaccard_50_dominance,lmer_overlap_nullmodell_50)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_jaccard_50_dominance)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs
#### core area
lmer_jaccard_95_dominance <- lmer(jaccard_overlap ~ dominancediff+(1 | focal)+(1| n_gps)+(1| algorithm),all_overlaps_95_naomit_dominance,REML=FALSE)
anova(lmer_jaccard_95_dominance,lmer_overlap_nullmodell_95)
##### Create P values for lm
coefs <- data.frame(coef(summary(lmer_numOfPoints95)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs
```

```{r DELETE Homerange - Normalized fruit tree points,echo=FALSE, message=FALSE,comment=FALSE, warning=FALSE}
SUAQ_waypoints_11all20_loc_tree <- SUAQ_waypoints_11all20_loc %>% filter(PtType=="tree")
SUAQ_waypoints_11all20_loc_tree_sf <-  st_as_sf(SUAQ_waypoints_11all20_loc_tree, 
                          coords = c("E","N"), 
                          crs = 23847) # can also do that in the new sricpt
SUAQ_waypoints_11all20_loc_tree_sf_xy <- SUAQ_waypoints_11all20_loc_tree_sf %>% dplyr::select(geometry)
#### Create a agggregated tree dataset. Distance of trees minimum 20m to be aggregated
data(meuse)
pts <- st_as_sf(meuse, coords = c("x", "y"),remove=FALSE)

# This will give lots of warnings since there are non-numeric columns
pts_agg <- aggregate.sf(pts, pts, mean, 
     join = function(x, y) st_is_within_distance(x, y, dist = 100))


#### Create a raster over research area with resolution in meter
resolution <- 25
r <- raster(xmn=xmin, ymn=ymin, xmx=xmax, ymx=ymax, res=resolution)
r[] <- 0

# note that you should use (lon, lat), in that order!
SUAQ_waypoints_11all20_loc_tree_xy <- SUAQ_waypoints_11all20_loc_tree %>% dplyr::select(E,N)
r_trees <- rasterize(SUAQ_waypoints_11all20_loc_tree_xy, r, fun="count")
plot(r_trees)

# note that you should use (lon, lat), in that order!
SUAQ_waypoints_11all20_loc_xy <- SUAQ_waypoints_11all20_loc %>% dplyr::select(E,N)
r_allGPS <- rasterize(SUAQ_waypoints_11all20_loc_xy, r, fun="count")
plot(r_allGPS)


normalized_tree <- overlay(r_trees, r_allGPS, fun="/")
plot(normalized_tree)



#### Visuals
#### Analyze ??: distribution of weather?
test<- ggplot()+
  #geom_path(aes(E,N), alpha = 0.7,size=0.5)+
  geom_sf(data = pts_agg,
                                               aes(colour = "green"),
                                               colour = "green")+
  geom_sf(data = pts,
                                               aes(colour = "black"),
                                               colour = "black",size=0.5)
  
ggplotly(test)


```

