---
title: "Projekt"
author: "Stefan Graf"
date: "28 3 2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Markings:
#### Quality check: --> are quality checks of chunks executed before
########### DELETE ############--> are chunks to delete
## Title --> are titles for sections
### explanations of the following chunk
##### SUAQ TRACKS 2018 & 2019 ##### --> Code sections to understand code better

```{r import, echo= 'F',results='hide'}
# CLASSICAL PACKAGE IMPORTER (Stefan Graf & Tobias Frey)
packages_checker <- function(packages_list){
  for (package in packages_list){
    if (!require(package, character.only = T)) {
      install.packages(package)
    }
    library(package, character.only = T)
  }
}
packages_checker(
  c('tmaptools',# read_gpx function
    'plotKML',# read GPX to dataframe
    'lubridate',
    'zoo',
    'data.tree',
    'hR',# making realtionships of orangutans
    'dbplyr',
    'plyr',
    'dplyr',
    'tidyr',
    'tidyverse',
    'sf', # Simple feature --> super spatial data handling structure
    'raster', # For raster data
    'ggspatial',
    'rgdal',
    'data.table',
    'RColorBrewer', # color schemas from color brewer web
    'ggraph', # visualization
    'igraph',
    'tmap',
    'scales',
    'trajr',
    'plotly', # zoomable, pannable ggplot
    'mapview',
    'ggmap', # easier interactive ggmaps 
    'leaflet', # Javascript library (originally) for plotting on a maptile interactive map
    'tmap',
    'stringr',
    'remotes', # geom_convexhull function
    'magrittr',
    'ggridges',
    'forcats',
    'move',# Brownian Bridge Movement Model
    'adehabitatHR',
    'ggpubr',
    'htmlwidgets',
    'gridExtra',# additional to ggplot, can add a density plot on the side of a histogram "marginal plot"
    'grid',
    'ggExtra'
    ))
rm(list = ls())
getwd()


### CUSTOM FUNCTIONS ###

# Function for displaying scale labels in ggplot not as f.e. 4e+05
fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, scientific = TRUE)
     # quote the part before the exponent to keep all the digits
     l <- gsub("^(.*)e", "'\\1'e", l)
     # turn the 'e+' into plotmath format
     l <- gsub("e", "%*%10^", l)
     # return this as an expression
     parse(text=l)
}


clockS = function(t){hour(t)*3600+minute(t)*60+second(t)}
st_kde <- function(points,cellsize, bandwith, extent = NULL){
  require(MASS)
  require(raster)
  require(sf)
  if(is.null(extent)){
    extent_vec <- st_bbox(points)[c(1,3,2,4)]
  } else{
    extent_vec <- st_bbox(extent)[c(1,3,2,4)]
  }
  
  n_y <- ceiling((extent_vec[4]-extent_vec[3])/cellsize)
  n_x <- ceiling((extent_vec[2]-extent_vec[1])/cellsize)
  
  extent_vec[2] <- extent_vec[1]+(n_x*cellsize)-cellsize
  extent_vec[4] <- extent_vec[3]+(n_y*cellsize)-cellsize
  
  coords <- st_coordinates(points)
  matrix <- kde2d(coords[,1],coords[,2],h = bandwith,n = c(n_x,n_y),lims = extent_vec)
  raster(matrix)
}




#### Custom functions for analysis
clockS = function(t){hour(t)*3600+minute(t)*60+second(t)}


euclid <- function(x1,y1,x2,y2){
  return(sqrt((x1-x2)^2+(y1-y2)^2))
}

turning_angle <- function(x,y,lead_lag = 1){ 
  if(length(x) < 3){return(NA)}
  if(length(x) != length(y)){stop("x and y must be of the same length")}
  p1x <- lag(x,lead_lag)
  p1y <- lag(y,lead_lag)
  p2x <- x
  p2y <- y
  p3x <- lead(x,lead_lag)
  p3y <- lead(y,lead_lag)
  p12 <- euclid(p1x,p1y,p2x,p2y)
  p13 <- euclid(p1x,p1y,p3x,p3y)
  p23 <- euclid(p2x,p2y,p3x,p3y)
  rad <- acos((p12^2+p23^2-p13^2)/(2*p12*p23))
  grad <- (rad*180)/pi
  grad[p12 == 0 | p23 == 0] <- NA
  d <-  (p3x-p1x)*(p2y-p1y)-(p3y-p1y)*(p2x-p1x)
  d <- ifelse(d == 0,1,d)
  d[d>0] <- 1
  d[d<0] <- -1
  d[d==0] <- 1
  turning <- grad*d*-1+180
  return(turning)
}

# Source https://exploratory.io/note/kanaugust/1701090969905358
calculate_mode <- function(x) {
  uniqx <- unique(na.omit(x))
  uniqx[which.max(tabulate(match(x, uniqx)))]
}

```

#### SUAQ research site
## Preprocessing
1. Getting data from excel/csv files
2. Getting data from gpx files, clean and order it
3. Merge Files
```{r GPS Loader}
### Load Project variables ### 
##### waypoints ##### 
SUAQ_waypoints_11all20_loc <- read_csv("data/Processed_data/SUAQ_waypoints_11all20.csv")

##### Weather and FAI #####
SUAQ_weather <- read_csv("data/Processed_data/SUAQ_weather.csv")
SUAQ_weather <- SUAQ_weather %>% rowwise() %>% 
  mutate(avg_temperature_tot=mean(c(SUAQ_weather.min_temp_average,SUAQ_weather.max_temp_average),na.rm=TRUE),
         rainfall_avg=mean(c(as.numeric(SUAQ_weather.rainfall_ml_morning),as.numeric(SUAQ_weather.rainfall_ml_evening)),na.rm=TRUE))
SUAQ_weather_monthly<-SUAQ_weather %>% 
  group_by(SUAQ_weather.year,SUAQ_weather.month) %>% 
  summarise_all(funs(mean))
SUAQ_weather_monthly<- SUAQ_weather_monthly%>% dplyr::select(-SUAQ_weather.weather_id,-SUAQ_weather.day,-SUAQ_weather.daymonthyear)





SUAQ_fai <- read_csv("data/Processed_data/SUAQ_fai.csv")

##### Joining info based on date #####
SUAQ_waypoints_11all20_loc <- SUAQ_waypoints_11all20_loc %>% left_join(SUAQ_weather,by=c("manualDate"="SUAQ_weather.daymonthyear"))

autoplot(glm_rainfall)

glm_rainfall<- lm(manual_distTolast~rainfall_avg,SUAQ_waypoints_11all20_loc)
summary(lm)


##### Selection only pointtypes which are taken on the orangutan way? is this true for all? ##### 
SUAQ_rangepoints_11all20_loc <- SUAQ_waypoints_11all20_loc %>% filter(SUAQ_waypoints_11all20_loc$PtType %in% c("range","foun","lost","mornnest","nightnest","daynest"))



##### Project attributes to calculate DJL ##### 
#### Create timelag to last waypoints in not spatial dataframe #### 
SUAQ_rangepoints_11all20_loc<- SUAQ_rangepoints_11all20_loc %>% 
  group_by(follow)%>% arrange(manualTime,.by_group = TRUE) %>% 
  mutate(timelag=as.numeric(difftime(as.POSIXlt(manualTime),
                                     as.POSIXlt(lag(manualTime)),units 
                                     ="secs")),
         # calculate distance with approach number 1 --> st_distance but to do that we need to calculate lead first because the lead function doesnt work within st distance (not same datatype (df))
         timelag_lead=as.numeric(difftime(as.POSIXlt(lead(manualTime)),
                                          as.POSIXlt(manualTime),units ="secs")),
         manual_distTolast=sqrt((E-lag(E))^2+(N-lag(N))^2),
         manual_distToNext=sqrt((E-lead(E))^2+(N-lead(N))^2),
         speed_last = manual_distTolast/timelag,
         speed_next = manual_distToNext/timelag_lead,
         turningAngle=turning_angle(E,N)
         )
SUAQ_waypoints_11all20_loc<- SUAQ_waypoints_11all20_loc %>% 
  group_by(follow)%>% arrange(manualTime,.by_group = TRUE) %>% 
  mutate(timelag=as.numeric(difftime(as.POSIXlt(manualTime),
                                     as.POSIXlt(lag(manualTime)),units 
                                     ="secs")),
         # calculate distance with approach number 1 --> st_distance but to do that we need to calculate lead first because the lead function doesnt work within st distance (not same datatype (df))
         timelag_lead=as.numeric(difftime(as.POSIXlt(lead(manualTime)),
                                          as.POSIXlt(manualTime),units ="secs")),
         manual_distTolast=sqrt((E-lag(E))^2+(N-lag(N))^2),
         manual_distToNext=sqrt((E-lead(E))^2+(N-lead(N))^2),
         speed_last = manual_distTolast/timelag,
         speed_next = manual_distToNext/timelag_lead,
         turningAngle=turning_angle(E,N)
         )




##### Selection only females ##### 
SUAQ_waypoints_11all20_loc_females <-
  SUAQ_waypoints_11all20_loc %>%   filter(!(SUAQ_orangutans.Sex == "male"))
SUAQ_rangepoints_11all20_loc_females <-
  SUAQ_rangepoints_11all20_loc %>%   filter(!(SUAQ_orangutans.Sex == "male"))
SUAQ_rangepoints_11all20_loc_females_10 <-
  SUAQ_waypoints_11all20_loc %>%   filter(!(SUAQ_orangutans.Sex == "male")) %>% filter(
    focal %in% c(
      "lisa",
      "friska",
      "ellie",
      "cissy",
      "lilly",
      "yulia",
      "raffi",
      "sarabi",
      "trident",
      "alice",
      "mocca",
      "cinnamon"
    )
  )
SUAQ_waypoints_11all20_loc_females_10 <-
  SUAQ_waypoints_11all20_loc %>%   filter(!(SUAQ_orangutans.Sex == "male")) %>% filter(
    focal %in% c(
      "lisa",
      "friska",
      "ellie",
      "cissy",
      "lilly",
      "yulia",
      "raffi",
      "sarabi",
      "trident",
      "alice",
      "mocca",
      "cinnamon"
    )
  )


```


### Analysis
## First visuals SUAQ
```{r AnalysisDJL}
##### DJL ######
DJL_follows<- SUAQ_waypoints_11all20_loc_females_10 %>% 
  filter(followtype=="NN") %>% 
  group_by(follow,focal,SUAQ_followlog.ClassFocal) %>% 
  summarise(DJL=sum(as.numeric(manual_distTolast), na.rm = TRUE),count=n()) %>% 
  arrange(SUAQ_followlog.ClassFocal) %>% ungroup() %>% 
  mutate(focal = fct_reorder(focal, SUAQ_followlog.ClassFocal))

plot1<- DJL_follows %>% 
  ggplot( aes(x=focal, y=DJL)) +
    geom_jitter(color="coral4", size=0.7, alpha=0.9, shape = 4)+ # add  aes(x=focal, y=DJL) to see which follow it is on a plotly although colors are wrong then... little hack
    geom_boxplot(aes(x=focal, y=DJL, fill=focal),outlier.fill = focal,outlier.stroke= focal) +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    ggtitle("Day journey length for Nest-to-Nest follows in Suaq grouped by orangutans") +
    xlab("")+
    ylab("DJL [m]")+
    coord_cartesian(ylim=c(0, 5000))+
    theme(legend.position = "none")+
    theme(axis.text.x = element_text(angle = 90))
ggplotly(plot1)
view(SUAQ_waypoints_11all20_loc_females_10[SUAQ_waypoints_11all20_loc_females_10$follow==1822,])


##### Edge lengths ######
plot_distToLats_over_200<-SUAQ_rangepoints_11all20_loc %>% 
  filter(followtype=="NN") %>% 
  filter((manual_distToNext>200)&( manual_distTolast >200)) %>% 
  group_by(focal,manualDate,follow) %>%
  ggplot(aes(E,N))+
  geom_point(aes(color=PtType),size=1)+
  geom_path(aes(color=as.factor(follow)), alpha = 0.8,size=0.6)
ggplotly(plot_distToLats_over_200)


##### DJL & Weather ######
plot_weather_avgRainfall<-SUAQ_weather_monthly %>% 
  mutate(SUAQ_weather.month = fct_relevel(SUAQ_weather.month, 
            "January", "February", "March", 
            "April", "May", "June", 
            "July", "August", "September", "October","November", "December")) %>% 
  ggplot()+
  geom_tile(aes(SUAQ_weather.month, as.factor(SUAQ_weather.year), fill= rainfall_avg))+
  scale_fill_distiller() +
  theme_ipsum() +
  scale_fill_gradient(low="mintcream", high="steelblue",na.value = NA) +
  labs(title = "Average rainfall total",x="",y="") + labs(fill='rainfall [ml]')+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.7))


plot_weather_moRainfall<-SUAQ_weather_monthly %>% 
  mutate(SUAQ_weather.month = fct_relevel(SUAQ_weather.month, 
            "January", "February", "March", 
            "April", "May", "June", 
            "July", "August", "September", "October","November", "December")) %>% 
  ggplot()+
  geom_tile(aes(SUAQ_weather.month, as.factor(SUAQ_weather.year), fill= SUAQ_weather.rainfall_ml_morning))+
  scale_fill_distiller() +
  theme_ipsum() +
  scale_fill_gradient(low="mintcream", high="steelblue",na.value = NA) +
  labs(title = "Average rainfall morning",x="",y="") + labs(fill='rainfall [ml]')+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.7))


plot_weather_eveRainfall<-SUAQ_weather_monthly %>% 
  mutate(SUAQ_weather.month = fct_relevel(SUAQ_weather.month, 
            "January", "February", "March", 
            "April", "May", "June", 
            "July", "August", "September", "October","November", "December")) %>% 
  ggplot()+
  geom_tile(aes(SUAQ_weather.month, as.factor(SUAQ_weather.year), fill= SUAQ_weather_monthly$SUAQ_weather.rainfall_ml_evening))+
  scale_fill_distiller() +
  theme_ipsum() +
  scale_fill_gradient(low="mintcream", high="steelblue",na.value = NA) +
  labs(title = "Average rainfall evening",x="",y="") + labs(fill='rainfall [ml]')+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.7))


plot_weather_moWaterLevel<-SUAQ_weather_monthly %>% 
  mutate(SUAQ_weather.month = fct_relevel(SUAQ_weather.month, 
            "January", "February", "March", 
            "April", "May", "June", 
            "July", "August", "September", "October","November", "December")) %>% 
  ggplot()+
  geom_tile(aes(SUAQ_weather.month, as.factor(SUAQ_weather.year), fill= SUAQ_weather_monthly$SUAQ_weather.river_waterlevel_morning))+
  scale_fill_distiller() +
  theme_ipsum() +
  scale_fill_gradient(low="mintcream", high="steelblue",na.value = NA) +
  labs(title = "Waterlevel morning",x="",y="") + labs(fill='waterlevel [cm?]')+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.7))


plot_weather_eveWaterLevel<-SUAQ_weather_monthly %>% 
  mutate(SUAQ_weather.month = fct_relevel(SUAQ_weather.month, 
            "January", "February", "March", 
            "April", "May", "June", 
            "July", "August", "September", "October","November", "December")) %>% 
  ggplot()+
  geom_tile(aes(SUAQ_weather.month, as.factor(SUAQ_weather.year), fill= SUAQ_weather_monthly$SUAQ_weather.river_waterlevel_evening))+
  scale_fill_distiller() +
  theme_ipsum() +
  scale_fill_gradient(low="mintcream", high="steelblue",na.value = NA) +
  labs(title = "Waterlevel evening",x="",y="") + labs(fill='waterlevel [cm?]')+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.7))


plot_weather_Avgtemp<-SUAQ_weather_monthly %>% 
  mutate(SUAQ_weather.month = fct_relevel(SUAQ_weather.month, 
            "January", "February", "March", 
            "April", "May", "June", 
            "July", "August", "September", "October","November", "December")) %>%
  ggplot()+
  geom_tile(aes(SUAQ_weather.month, as.factor(SUAQ_weather.year), fill= avg_temperature_tot))+
  scale_fill_distiller() +
  theme_ipsum() +
  scale_fill_gradient(low="mintcream", high="deeppink3",na.value = NA) +
  labs(title = "Average temperature",x="",y="") + labs(fill='temperature [C°]')+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.7))

plot_weather_AvgMaxtemp<-SUAQ_weather_monthly %>% 
  mutate(SUAQ_weather.month = fct_relevel(SUAQ_weather.month, 
            "January", "February", "March", 
            "April", "May", "June", 
            "July", "August", "September", "October","November", "December")) %>%
  ggplot()+
  geom_tile(aes(SUAQ_weather.month, as.factor(SUAQ_weather.year), fill= SUAQ_weather.max_temp_average))+
  scale_fill_distiller() +
  theme_ipsum() +
  scale_fill_gradient(low="mintcream", high="deeppink3",na.value = NA) +
  labs(title = "Average maximum temperature",x="",y="") + labs(fill='temperature [C°]')+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.7))


plot_weather_AvgMintemp<-SUAQ_weather_monthly %>% 
  mutate(SUAQ_weather.month = fct_relevel(SUAQ_weather.month, 
            "January", "February", "March", 
            "April", "May", "June", 
            "July", "August", "September", "October","November", "December")) %>% 
  ggplot()+
  geom_tile(aes(SUAQ_weather.month, as.factor(SUAQ_weather.year), fill= SUAQ_weather.min_temp_average))+
  scale_fill_distiller() +
  theme_ipsum() +
  scale_fill_gradient(low="mintcream", high="deeppink3",na.value = NA) +
  labs(title = "Average minimum temperature",x="",y="") + labs(fill='temperature [C°]')+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.7))


plot_weather_avgRainfall
plot_weather_moRainfall
plot_weather_eveRainfall
plot_weather_moWaterLevel
plot_weather_eveWaterLevel
plot_weather_AvgMaxtemp
plot_weather_AvgMintemp
plot_weather_Avgtemp



```



